---
title: "Figure 4 Analysis"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
  
}
```


# Libraries

Load necessary libraries for analysis.

```{r Necessary Libraries, message=FALSE, warning=FALSE, echo=TRUE}
library(tidyverse)
library(readxl)
library(readr)
library(clusterProfiler)
library(org.Hs.eg.db)
```

Define my custom plot theme

```{r Custom plot theme, echo=TRUE, message=FALSE, warning=FALSE}

# Define the custom theme
# plot_theme_custom <- function() {
#   theme_minimal() +
#     theme(
#       #line for x and y axis
#       axis.line = element_line(linewidth = 1,
#                                color = "black"),
# 
#       #axis ticks only on x and y, length standard
#       axis.ticks.x = element_line(color = "black",
#                                 linewidth = 1),
#       axis.ticks.y = element_line(color = "black",
#                                 linewidth = 1),
#       axis.ticks.length = unit(0.05, "in"),
# 
#       #text and font
#       axis.text = element_text(color = "black",
#                                family = "Arial",
#                                size = 7),
#       axis.title = element_text(color = "black",
#                                 family = "Arial",
#                                 size = 9),
#       legend.text = element_text(color = "black",
#                                  family = "Arial",
#                                  size = 7),
#       legend.title = element_text(color = "black",
#                                   family = "Arial",
#                                   size = 9),
#       plot.title = element_text(color = "black",
#                                 family = "Arial",
#                                 size = 10),
# 
#       #blank background and border
#       panel.background = element_blank(),
#       panel.border = element_blank(),
# 
#       #gridlines for alignment
#       panel.grid.major = element_line(color = "grey80", linewidth = 0.5),  #grey major grid for align in illus
#       panel.grid.minor = element_line(color = "grey90", linewidth = 0.5) #grey minor grid for align in illus
#     )
# }

# saveRDS(plot_theme_custom, "data/plot_theme_custom.RDS")

# theme_custom <- readRDS("data/plot_theme_custom.RDS")
theme_custom <- readRDS("data/plot_theme_custom_update.RDS")
```


Define a function to save plots as .pdf and .png

```{r pdf png saving function, echo=TRUE, message=FALSE, warning=FALSE}

save_plot <- function(plot, filename, 
                      folder = ".", 
                      width = 8, 
                      height = 6, 
                      units = "in", 
                      dpi = 300, 
                      add_date = TRUE) {
  if (missing(filename)) stop("Please provide a filename (without extension) for the plot.")

  date_str <- if (add_date) paste0("_", format(Sys.Date(), "%y%m%d")) else ""
  pdf_file <- file.path(folder, paste0(filename, date_str, ".pdf"))
  png_file <- file.path(folder, paste0(filename, date_str, ".png"))
  ggsave(filename = pdf_file, plot = plot, device = cairo_pdf, width = width, height = height, units = units, bg = "transparent")
  ggsave(filename = png_file, plot = plot, device = "png", width = width, height = height, units = units, dpi = dpi, bg = "transparent")
  message("Saved plot as Cairo PDF: ", pdf_file)
  message("Saved plot as PNG: ", png_file)
}

output_folder <- "C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/Fig pdfs"


#save plot function created
#to use: just define the plot name, filename_base, width, height

#here is a function to save complex heatmaps to pdf and png

save_complex_heatmap <- function(plot, filename, folder, width = 6, height = 8, add_date = TRUE) {
  if (!dir.exists(folder)) dir.create(folder, recursive = TRUE)
  
  #add in the date to the end of the filename
  #this way I can easily track when each plot was created/saved
  date_str <- if (add_date) paste0("_", 
                                   format(Sys.Date(), 
                                          "%y%m%d")) else ""
  
  #add the date to the filename here
  pdf_path <- file.path(folder, paste0(filename, date_str, ".pdf"))
  png_path <- file.path(folder, paste0(filename, date_str, ".png"))
  
  #save to a pdf file for illustrator
  pdf(pdf_path, width = width, height = height)
  draw(plot, heatmap_legend_side = "right")
  dev.off()
  
  #now save the png file for easy visualization
  png(png_path, width = width, height = height, 
      units = "in", res = 300)
  draw(plot, heatmap_legend_side = "right")
  dev.off()
  
  message("Saved heatmap PDF to: ", pdf_path)
  message("Saved heatmap PNG: ", png_path)
}

#use this function in the same way as the above for all other plots
```

# Compare Response Clusters to Known TOP2i Treatments

## Read in TOP2i DEGs from Matthews et al., Paul et al.

Matthews ER, Johnson OD, Horn KJ, Gutiérrez JA, Powell SR, et al. (2024) Anthracyclines induce cardiotoxicity through a shared gene expression response signature. PLOS Genetics 20(2): e1011164. https://doi.org/10.1371/journal.pgen.1011164


Topoisomerase II inhibitors CX-5461 and Doxorubicin differ in their cardiotoxicity profiles
Sayan Paul, José A. Gutiérrez, Alyssa R. Bogar, Michelle C. Ward. bioRxiv 2025.07.20.665626; doi: https://doi.org/10.1101/2025.07.20.665626


```{r TOP2i DEG Lists, message=FALSE, warning=FALSE}

#These gene lists were obtained from:
#Matthews ER, Johnson OD, Horn KJ, Gutiérrez JA, Powell SR, et al. (2024) Anthracyclines induce cardiotoxicity through a shared gene expression response signature. PLOS Genetics 20(2): e1011164. https://doi.org/10.1371/journal.pgen.1011164

DEGs_RM_list <- readRDS("data/Fig4/DEGs_TOP2i_RM.RDS")

#only keep the conditions that I want to compare to
keep_conditions <- c("sigVDA24", "sigVDX24", "sigVEP24", "sigVMT24")

DEGs_RM_24 <- DEGs_RM_list[keep_conditions]

#change the first column to Entrez_ID for all sets
DEGs_RM_24 <- lapply(DEGs_RM_24, function(df) {
  # Rename the first column to Entrez_ID, regardless of current name
  colnames(df)[1] <- "Entrez_ID"
  return(df)
})

#DOX
DOX_RM_DEGs <- DEGs_RM_24$sigVDX24 %>% 
  as.data.frame()

dim(DOX_RM_DEGs)
#6645 genes

#DNR
DNR_RM_DEGs <- DEGs_RM_24$sigVDA24 %>% 
  as.data.frame()

dim(DNR_RM_DEGs)
#7017 genes

#EPI
EPI_RM_DEGs <- DEGs_RM_24$sigVEP24 %>% 
  as.data.frame()

dim(EPI_RM_DEGs)
#6328 genes

#MTX
MTX_RM_DEGs <- DEGs_RM_24$sigVMT24 %>% 
  as.data.frame()

dim(MTX_RM_DEGs)
#1115 genes

#now combine these all into one dataframe for later

DEGs_RM_24_df <- dplyr::bind_rows(
  # DOX_RM_DEGs %>% mutate(Drug = "DOX_RM"),
  DNR_RM_DEGs %>% mutate(Drug = "DAUN_RM"),
  EPI_RM_DEGs %>% mutate(Drug = "EPI_RM"),
  MTX_RM_DEGs %>% mutate(Drug = "MITO_RM"),
)


#These gene lists were obtained from: 
#Topoisomerase II inhibitors CX-5461 and Doxorubicin differ in their cardiotoxicity profiles
#Sayan Paul, José A. Gutiérrez, Alyssa R. Bogar, Michelle C. Ward
#bioRxiv 2025.07.20.665626; doi: https://doi.org/10.1101/2025.07.20.665626


#CX 0.5 24hr
CX_SP_24 <- read.csv("data/Fig4/Toptable_CX_0.5_24.csv")

dim(CX_SP_24)
#14279 genes, make sure to now filter this for significance

#DOX 0.5 24hr
DOX_SP_24 <- read.csv("data/Fig4/Toptable_DOX_0.5_24.csv")
dim(DOX_SP_24)
#14279 genes, filter for significance

#filter both above for adj.P.Val < 0.05

DEGs_SP_sig <- list(
  "CX_SP" = CX_SP_24$Entrez_ID[CX_SP_24$adj.P.Val < 0.05],
  "DOX_SP" = DOX_SP_24$Entrez_ID[DOX_SP_24$adj.P.Val < 0.05]
)

#now make dataframes and place them with RM's data to complete my Top2i comparative dataframe

CX_SP_DEGs <- CX_SP_24 %>%
  filter(adj.P.Val < 0.05) %>%
  as.data.frame() 

dim(CX_SP_DEGs)
#278 genes

DOX_SP_DEGs <- DOX_SP_24 %>%
  filter(adj.P.Val < 0.05) %>%
  as.data.frame()

dim(DOX_SP_DEGs)
#9828 genes

#since SP's data has Entrez_ID as an integer rather than a character, coerce these to character in order for bind_rows to work properly

# CX_SP_DEGs$Entrez_ID <- as.character(CX_SP_DEGs$Entrez_ID)
# DOX_SP_DEGs$Entrez_ID <- as.character(DOX_SP_DEGs$Entrez_ID)

#these two also don't have a SYMBOL column so let's add that in

#CX df
# CX_SP_DEGs <- CX_SP_DEGs %>%
#   mutate(
#     SYMBOL = mapIds(
#       org.Hs.eg.db,
#       keys = Entrez_ID,
#       column = "SYMBOL",
#       keytype = "ENTREZID",
#       multiVals = "first"
#     )
#   ) %>%
#   dplyr::select(Entrez_ID, SYMBOL, everything())  # move SYMBOL to second column

# Same for df6
# DOX_SP_DEGs <- DOX_SP_DEGs %>%
#   mutate(
#     SYMBOL = mapIds(
#       org.Hs.eg.db,
#       keys = Entrez_ID,
#       column = "SYMBOL",
#       keytype = "ENTREZID",
#       multiVals = "first"
#     )
#   ) %>%
#   dplyr::select(Entrez_ID, SYMBOL, everything())


# DEGs_Top2i_24_df <- dplyr::bind_rows(
#   DOX_RM_DEGs %>% mutate(Treatment = "DOX_RM"),
#   DNR_RM_DEGs %>% mutate(Treatment = "DAUN_RM"),
#   EPI_RM_DEGs %>% mutate(Treatment = "EPI_RM"),
#   MTX_RM_DEGs %>% mutate(Treatment = "MITO_RM"),
#   CX_SP_DEGs %>% mutate(Treatment = "CX_SP"),
#   DOX_SP_DEGs %>% mutate(Treatment = "DOX_SP")
# ) 
  
#now this table has all columns consistently named and in the same formats

# saveRDS(DEGs_Top2i_24_df, "data/Fig4/DEGlist_Top2i_SP_RM.RDS")

DEGs_Top2i_24_df <- readRDS("data/Fig4/DEGlist_Top2i_SP_RM.RDS")

#also make a list version just in case

# DEGs_Top2i_24_list <- list(
#   "DOX_RM" = DOX_RM_DEGs, #6645 DEGs
#   "DNR_RM" = DNR_RM_DEGs, #7017 DEGs
#   "EPI_RM" = EPI_RM_DEGs, #6328 DEGs
#   "MTX_RM" = MTX_RM_DEGs, #1115 DEGs
#   "CX_SP" = CX_SP_DEGs, #278 DEGs
#   "DOX_SP" = DOX_SP_DEGs #9828 DEGs
# )

#total 10,106 DEGs from SP
#total 21,105 DEGs from RM

# saveRDS(DEGs_Top2i_24_list, "data/Fig4/DEGs_Top2i_RM_SP_list.RDS")

DEGs_Top2i_24_list <- readRDS("data/Fig4/DEGs_Top2i_RM_SP_list.RDS")

```


# Figure 4A - Percentage of TOP2i DEGs in tx+0 DEGs

Proportion barplots with permutation testing to ascertain whether DEGs shared between DOX and other TOP2i treatments are more likely than chance to be in both groups.

```{r Fig4 Barplots Permutation Testing, message=FALSE, warning=FALSE}

#read in my DEGs
DOX_24T_R <- read.csv("data/DE/Toptable_RUV_24T_final.csv") %>% dplyr::select(-X)
DOX_24R_R <- read.csv("data/DE/Toptable_RUV_24R_final.csv") %>% dplyr::select(-X)
DOX_144R_R <- read.csv("data/DE/Toptable_RUV_144R_final.csv") %>% dplyr::select(-X)

deg_list <- list(
  "t0"   = DOX_24T_R$Entrez_ID[DOX_24T_R$adj.P.Val < 0.05],
  "t24"   = DOX_24R_R$Entrez_ID[DOX_24R_R$adj.P.Val < 0.05],
  "t144"  = DOX_144R_R$Entrez_ID[DOX_144R_R$adj.P.Val < 0.05]
)

#reference set - t0
t0_set <- unique(na.omit(as.character(deg_list$"t0")))
cat("Total t0 DEGs:", length(t0_set), "\n")

#background set - all genes tested
all_genes <- readRDS("data/DE/all_genes.RDS")
cat("Total background genes:", length(all_genes), "\n")


#define TOP2i treatment DEG sets from literature
comparison_DEGs <- list(
  DOX_RM = as.character(DOX_RM_DEGs$Entrez_ID),
  DNR_RM = as.character(DNR_RM_DEGs$Entrez_ID),
  EPI_RM = as.character(EPI_RM_DEGs$Entrez_ID),
  MTX_RM = as.character(MTX_RM_DEGs$Entrez_ID),
  CX_SP  = as.character(CX_SP_DEGs$Entrez_ID),
  DOX_SP = as.character(DOX_SP_DEGs$Entrez_ID)
)

#build data for a stacked barplot with percentages

plot_df <- purrr::map_dfr(names(comparison_DEGs), function(name) {
  comp_DEGs <- unique(comparison_DEGs[[name]])
  
  overlap <- sum(t0_set %in% comp_DEGs)
  non_overlap <- length(t0_set) - overlap
  
  tibble(
    Treatment = name,
    Category = c("DEG", "nonDEG"),
    Count = c(overlap, non_overlap),
    Total = length(t0_set),
    Fraction = c(overlap, non_overlap)/length(t0_set)
  )
})

plot_df <- plot_df %>%
  mutate(
    Treatment = factor(Treatment, levels = names(comparison_DEGs)),
    Category = factor(Category, levels = c("DEG", "nonDEG"))
  )

#legend labels (with n counts)
legend_labels <- plot_df %>%
  filter(Category == "DEG Status") %>%
  mutate(Label = paste0(Category, " (n=", Count, ")")) %>%
  dplyr::select(Treatment, Label) %>%
  deframe()

#perform permutation testing for enrichment
perm_test_overlap <- function(gene_set, t0_genes, gene_universe, n_perm = 10000) {
  observed <- length(intersect(t0_genes, gene_set))
  
  perm_overlaps <- replicate(n_perm, {
    random_genes <- sample(gene_universe, length(t0_genes))
    length(intersect(random_genes, gene_set))
  })
  
  p_val <- mean(perm_overlaps >= observed)
  enrichment <- observed / mean(perm_overlaps)
  
  return(list(observed = observed,
              mean_perm = mean(perm_overlaps),
              enrichment = enrichment,
              p_value = p_val))
}

perm_results <- lapply(comparison_DEGs, 
                       perm_test_overlap,
                       t0_genes = t0_set,
                       gene_universe = all_genes,
                       n_perm = 10000)

#summary stats table
perm_summary_df <- map_dfr(names(perm_results), function(name) {
  res <- perm_results[[name]]
  tibble(
    Treatment = name,
    Observed = res$observed,
    MeanPerm = res$mean_perm,
    Enrichment = res$enrichment,
    PValue = res$p_value
  )
}) %>%
  mutate(FDR = p.adjust(PValue, method = "BH"),
         Significant = case_when(
           FDR < 0.001 ~ "***",
           FDR < 0.01  ~ "**",
           FDR < 0.05  ~ "*",
           TRUE ~ "ns"
         ))

#show summary stats table
print(perm_summary_df)

#label positions
#inside overlap bars: percentages
overlap_labels <- plot_df %>%
  filter(Category == "Overlap") %>%
  left_join(perm_summary_df %>% dplyr::select(Treatment, Significant), by = "Treatment") %>%
  mutate(
    Label = paste0(round(Fraction*100, 1), ""),
    y_pos = 0.5   #fixed position at 50%
  )

#above bars: significance stars
label_positions <- plot_df %>%
  group_by(Treatment) %>%
  summarise(y_pos = 1.0, .groups = "drop") %>%
  left_join(perm_summary_df %>% dplyr::select(Treatment, Significant), by = "Treatment")


#plot final barplot
TOP2i_Combined <- ggplot(plot_df, aes(x = Treatment, y = Fraction, fill = Category)) +
  geom_bar(stat = "identity", position = "stack", width = 0.6) +
  
  #percentages inside overlap bars (fixed position)
  geom_text(
    data = overlap_labels,
    aes(x = Treatment, y = y_pos, label = Label),
    inherit.aes = FALSE,
    color = "white",
    fontface = "bold",
    size = 4
  ) +
  
  #significance stars above bars
  geom_text(
    data = label_positions,
    aes(x = Treatment, y = y_pos, label = Significant),
    inherit.aes = FALSE,
    size = 4,
    fontface = "bold",
    color = "black"
  ) +
  
  scale_y_continuous(limits = c(0, 100)) +
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8"),
                    labels = legend_labels) +
  labs(
    title = "Proportion of TOP2i DEGs in tx+0 DEGs",
    x = "TOP2i Treatment",
    y = "Percentage of DOX tx+0 DEGs (%)",
    fill = "TOP2i DEGs"
  ) +
  theme_custom()

# save_plot(
#   plot = TOP2i_Combined,
#   folder = output_folder,
#   filename = "TOP2i_Barplot_Combined_Fig4_EMP",
#   height = 6,
#   width = 8
# )

```


# Figure S10 - TOP2i DEGs Acute vs Chronic Clusters

```{r Supp Fig 10 Barplots, message=FALSE, warning=FALSE}

#read in my chronic and acute cluster gene lists
final_genes_1_RUV <- readRDS("data/Fig3/final_genes_1_RUV.RDS")
final_genes_2_RUV <- readRDS("data/Fig3/final_genes_2_RUV.RDS")

acute_set <- unique(na.omit(final_genes_1_RUV))
chronic_set <- unique(na.omit(final_genes_2_RUV))

all_genes

comparison_DEGs <- list(
  DOX_RM = as.character(DOX_RM_DEGs$Entrez_ID),
  DNR_RM = as.character(DNR_RM_DEGs$Entrez_ID),
  EPI_RM = as.character(EPI_RM_DEGs$Entrez_ID),
  MTX_RM = as.character(MTX_RM_DEGs$Entrez_ID),
  CX_SP  = as.character(CX_SP_DEGs$Entrez_ID),
  DOX_SP = as.character(DOX_SP_DEGs$Entrez_ID)
)

#fisher's exact test per treatment
fisher_test_overlap <- function(acute, chronic, degs) {
  a <- sum(acute %in% degs)
  b <- length(acute) - a
  c <- sum(chronic %in% degs)
  d <- length(chronic) - c
  
  contingency <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  fisher_res <- fisher.test(contingency)
  
  list(
    contingency = contingency,
    p_value = fisher_res$p.value,
    significant = case_when(
      fisher_res$p.value < 0.001 ~ "***",
      fisher_res$p.value < 0.01  ~ "**",
      fisher_res$p.value < 0.05  ~ "*",
      TRUE ~ ""
    )
  )
}

# Prepare summary table for plotting
plot_summary <- map_dfr(names(comparison_DEGs), function(treat) {
  degs <- comparison_DEGs[[treat]]
  
  # Fisher test
  fisher_res <- fisher_test_overlap(acute_set, chronic_set, degs)
  
  # Counts for each motif
  acute_overlap <- sum(acute_set %in% degs)
  acute_nonoverlap <- length(acute_set) - acute_overlap
  chronic_overlap <- sum(chronic_set %in% degs)
  chronic_nonoverlap <- length(chronic_set) - chronic_overlap
  
  tibble(
    Treatment = treat,
    Motif = rep(c("Acute", "Chronic"), each = 2),
    Category = rep(c("DEG", "nonDEG"), 2),
    Count = c(acute_overlap, acute_nonoverlap, chronic_overlap, chronic_nonoverlap),
    Fraction = c(acute_overlap, acute_nonoverlap, chronic_overlap, chronic_nonoverlap) /
               c(length(acute_set), length(acute_set), length(chronic_set), length(chronic_set)),
    Significant = rep(c(fisher_res$significant, ""), 2)  #only on DEG bars
  )
})

#ensure correct order
plot_summary$Category <- factor(plot_summary$Category, levels = c("DEG", "nonDEG"))

#position stars at y = 1
label_positions <- plot_summary %>%
  filter(Category == "DEG") %>%
  mutate(y_pos = 1)

#plot
top2i_barplot_motifs_supp7 <- ggplot(plot_summary, aes(x = Treatment, y = Fraction, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = label_positions %>% filter(Significant != ""),
            aes(x = Treatment, y = y_pos, label = Significant),
            inherit.aes = FALSE, size = 6, fontface = "bold", color = "black") +
  geom_text(aes(label = paste0(round(Fraction*100,1), "")),
            position = position_stack(vjust = 0.5), size = 5, color = "white") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("DEG" = "blue", "nonDEG" = "red")) +
  facet_wrap(~Motif) +
  labs(
    title = "Percentage of TOP2i DEGs (Matthews et al., Paul et al.) in Acute and Chronic response clusters",
    x = "TOP2i Treatment",
    y = "Overlap Percentage (%)",
    fill = ""
  ) +
  theme_custom()

print(top2i_barplot_motifs_supp7)

# save_plot(
#   plot = top2i_barplot_motifs_supp7,
#   filename = "Supp7_TOP2i_Barplots_EMP",
#   folder = output_folder,
#   height = 5,
#   width = 8.5
# )

#print percentages and p-values

for(treat in names(comparison_DEGs)) {
  degs <- comparison_DEGs[[treat]]
  acute_overlap <- sum(acute_set %in% degs)
  chronic_overlap <- sum(chronic_set %in% degs)
  prop_ac <- acute_overlap / length(acute_set)
  prop_ch <- chronic_overlap / length(chronic_set)
  fisher_res <- fisher_test_overlap(acute_set, chronic_set, degs)
  
  cat(sprintf("%s - Acute: %.2f%% (%d/%d), Chronic: %.2f%% (%d/%d), Fisher p = %.5f\n",
              treat, 100*prop_ac, acute_overlap, length(acute_set),
              100*prop_ch, chronic_overlap, length(chronic_set),
              fisher_res$p_value))
}

#significance stars are for likelihood of being in the Chronic cluster over the Acute cluster - stars are the same on both sides but only really count for Chronic

```

# Figure 4B - Enrichment of TOP2i DEGs in Chronic cluster relative to Acute

```{r Fig 4b, message=FALSE, warning=FALSE}

#use the workflow above to now plot forest plots of odds ratios for these comparisons
#I am comparing sets of DEGs from other TOP2i against my motifs here
#I want to use my fisher's exact test data

#same inputs as before to preserve them
# final_genes_1_RUV
# final_genes_2_RUV

# acute_set 
# chronic_set

#remove DOX and preserve order DNR EPI MTX CX
comparison_DEGs_noDOX <- list(
  DNR_RM = as.character(DNR_RM_DEGs$Entrez_ID),
  EPI_RM = as.character(EPI_RM_DEGs$Entrez_ID),
  MTX_RM = as.character(MTX_RM_DEGs$Entrez_ID),
  CX_SP  = as.character(CX_SP_DEGs$Entrez_ID)
)

treatment_order <- c("DNR_RM", "EPI_RM", "MTX_RM", "CX_SP")

fisher_test_overlap <- function(acute, chronic, degs) {
  a <- sum(acute %in% degs)
  b <- length(acute) - a
  c <- sum(chronic %in% degs)
  d <- length(chronic) - c
  
  contingency <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  fisher_res <- fisher.test(contingency)
  
  tibble(
    OR = unname(fisher_res$estimate),
    CI_low = fisher_res$conf.int[1],
    CI_high = fisher_res$conf.int[2],
    p_value = fisher_res$p.value,
    Significant = case_when(
      fisher_res$p.value < 0.001 ~ "***",
      fisher_res$p.value < 0.01 ~ "**",
      fisher_res$p.value < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )
}


#build a summary table for the results of fishers exact for the forest plots
forest_summary <- map_dfr(names(comparison_DEGs_noDOX), function(treat) {
  res <- fisher_test_overlap(chronic_set, acute_set, comparison_DEGs_noDOX[[treat]])
  res %>% mutate(Treatment = treat)
}) %>% 
  mutate(Treatment = factor(Treatment, levels = rev(treatment_order)), #define order
         Group = ifelse(grepl("_RM$", Treatment), "RM", "SP"))

#create my combined plot
forest_TOP2i <- ggplot(forest_summary, 
                       aes(y = Treatment, x = OR, 
                           xmin = CI_low, xmax = CI_high)) + 
  geom_point(size = 4) +
  geom_errorbarh(height = 0.2, linewidth = 0.5, color = "black") +
  geom_text(aes(label = Significant), position = position_nudge(y = 0.1), 
            color = "black", fontface = "bold", size = 5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  scale_x_log10(limits = c(0.1, 1000),
                breaks = c(0.1, 1, 10, 100, 1000)) +
  scale_color_manual(values = c("_" = "grey", "_" = "grey")) + #remove the groups this way even though it's not the best method
  labs(
    title = "Enrichment of TOP2i DEGs in Chronic response cluster relative to Acute response cluster",
    x = "log odds ratio (95% confidence interval)",
    y = ""
  ) +
  theme_custom()

print(forest_TOP2i)

#make a summary table

forest_summary_TOP2i <- forest_summary

View(forest_summary_TOP2i)



```


# DOX Proteomics Comparisons

For this comparison, we utilized Supplementary Table 2 from Johnson, O. D., Paul, S., Gutiérrez, J. A., Russell, W. K., & Ward, M. C. (2025). DNA-damage-associated protein co-expression network in cardiomyocytes informs on tolerance to genetic variation and disease. iScience, 28(5). https://doi.org/10.1016/j.isci.2025.112474 

I have taken the differentially abundant proteins (DAPs) identified after DOX treatment for 24 hours, as defined by an adj. p value < 0.05

We would like to assess whether these DAPs are preferentially encoded by Acute or Chronic genes. On the flip side, we would like to assess whether the nonDAPs in the set are also enriched in one cluster vs the other.

```{r read in DAPs and nonDAPs from ODJ SP proteomics paper, warning=FALSE, message=FALSE}
#I have the excel sheet downloaded from the paper mentioned above, which I will read in here and then filter the proteins for DAPs vs nonDAPs

doxprot_supp2 <- read_excel("C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/proteomics_supptable2_ODJ_SP_MCW_downloaded_260408.xlsx", 
    range = "B2:AA4180")

#now go ahead and make two lists, one with DAPs and one with nonDAPs
dox_daps <- doxprot_supp2 %>% 
  filter(P.Value.adj < 0.05) %>% 
  pull(Protein)
#according to the paper, should be 319 DAPs
length(dox_daps)

dox_nondaps <- doxprot_supp2 %>%
  filter(P.Value.adj >= 0.05) %>% 
  pull(Protein)
#according to the paper, should be 3,859 DAPs
length(dox_nondaps)

#now combine these into one list
doxprot_list <- list(
  DAPs = dox_daps,
  nonDAPs = dox_nondaps
)

#now I'll have to convert between these protein names (UniProt) and my gene clusters (Entrez_ID) - translate these DAPs into Entrez_ID
#I will use bitr from the clusterProfiler package as a first attempt to translate between ids

```


## Identify DAPs and nonDAPs encoded by genes in Acute vs Chronic clusters

```{r translate daps to genes, warning=FALSE, message=FALSE}

#1. translate DAPs and nonDAPs into Entrez_ID
dox_daps_tr <- bitr(dox_daps,
                    fromType = "UNIPROT",
                    toType = "ENTREZID",
                    OrgDb = org.Hs.eg.db) %>% 
  pull(ENTREZID)

dox_nondaps_tr <- bitr(dox_nondaps,
                       fromType = "UNIPROT",
                       toType = "ENTREZID",
                       OrgDb = org.Hs.eg.db) %>% 
  pull(ENTREZID)

#print the number of Entrez_ID entries found per set
cat("DAPs:   ", length(dox_daps),    "| mapped to Entrez_ID:", length(dox_daps_tr), "\n")
cat("nonDAPs: ", length(dox_nondaps),  "| mapped to Entrez_ID:", length(dox_nondaps_tr), "\n")

#2. define gene sets of interest (Acute and Chronic clusters)
comparison_prot <- list(
  DAP = dox_daps_tr,
  nonDAP = dox_nondaps_tr
)

comparison_genes <- list(
  Acute = acute_set,
  Chronic = chronic_set
)

```

```{r daps vs nondaps acute vs chronic sets}
#bars = Acute and Chronic clusters
#fill = proportion of genes in each cluster that are DAP vs nonDAP
#Fisher's test = are DAPs over-represented in Acute vs Chronic?
#iterate over the two clusters, and for each one ask
#        "how many of these genes encode a DAP vs a nonDAP?"

#make a function to overlap genes in the fisher's exact test
fisher_test_overlap_genes <- function(acute, chronic, genes) {
  a <- sum(acute %in% genes)
  b <- length(acute) - a
  c <- sum(chronic %in% genes)
  d <- length(chronic) - c

  contingency <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  fisher_res  <- fisher.test(contingency)

  list(
    contingency = contingency,
    p_value     = fisher_res$p.value,
    significant = case_when(
      fisher_res$p.value < 0.001 ~ "***",
      fisher_res$p.value < 0.01  ~ "**",
      fisher_res$p.value < 0.05  ~ "*",
      TRUE                       ~ ""
    )
  )
}

plot_summary_plot1 <- map_dfr(c("Acute", "Chronic"), function(cluster) {
  
  genes <- if (cluster == "Acute") acute_set else chronic_set
  
  #how many genes in this cluster encode a DAP vs nonDAP
  dap_overlap    <- sum(genes %in% dox_daps_tr)
  nondap_overlap <- length(genes) - dap_overlap  # everything else is nonDAP
  
  fisher_res <- fisher_test_overlap_genes(acute_set, chronic_set, dox_daps_tr)
  
  tibble(
    Cluster     = cluster,
    Category    = c("DAP", "nonDAP"),
    Count       = c(dap_overlap, nondap_overlap),
    Fraction    = c(dap_overlap, nondap_overlap) / length(genes),
    Significant = c(fisher_res$significant, "")
  )
})

plot_summary_plot1$Category <- factor(plot_summary_plot1$Category, 
                                      levels = c("DAP", "nonDAP"))

label_positions_plot1 <- plot_summary_plot1 %>%
  filter(Category == "DAP") %>%
  mutate(y_pos = 1)

dap_barplot_plot1 <- ggplot(plot_summary_plot1, 
                            aes(x = Cluster, y = Fraction, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = label_positions_plot1 %>% filter(Significant != ""),
            aes(x = Cluster, y = y_pos, label = Significant),
            inherit.aes = FALSE, size = 6, fontface = "bold", color = "black") +
  geom_text(aes(label = paste0(round(Fraction * 100, 1), "")),
            position = position_stack(vjust = 0.5), size = 5, color = "white") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("DAP" = "#0f9eff", "nonDAP" = "#19aa99")) +
  labs(
    title = "Proportion of Acute and Chronic cluster genes encoding DAPs vs nonDAPs",
    x     = "Gene Cluster",
    y     = "Proportion of Cluster Genes",
    fill  = ""
  ) +
  theme_custom()

print(dap_barplot_plot1)

# save_plot(dap_barplot_plot1,
#           filename = "DAPvnonDAP_inChronicvAcute_Barplot_Proportion_EMP",
#           folder = output_folder)

#print summary statistics for plot 1
fisher_res <- fisher_test_overlap_genes(acute_set, chronic_set, dox_daps_tr)

for (cluster in c("Acute", "Chronic")) {
  genes       <- if (cluster == "Acute") acute_set else chronic_set
  dap_overlap <- sum(genes %in% dox_daps_tr)
  prop_dap    <- dap_overlap / length(genes)

  cat(sprintf("%s - DAP: %.2f%% (%d/%d), 
              nonDAP: %.2f%% (%d/%d), 
              Fisher p = %.5f\n",
              cluster,
              100 * prop_dap, dap_overlap,
              length(genes),
              100 * (1 - prop_dap), 
              length(genes) - dap_overlap, 
              length(genes),
              fisher_res$p_value))
}
```

```{r comparing proportion of daps and nondaps, message=FALSE, warning=FALSE}
#bars = DAPs and nonDAPs
#fill = proportion of each protein set whose genes fall in Acute vs Chronic
#fisher's test = are Acute genes over-represented among DAPs vs nonDAPs?
#logic: iterate over the two PROTEIN SETS, and for each one ask
#        "how many of these proteins come from Acute vs Chronic genes?"

plot_summary_plot2 <- map_dfr(c("DAP", "nonDAP"), function(prot_set) {
  
  genes <- if (prot_set == "DAP") dox_daps_tr else dox_nondaps_tr
  
  acute_overlap   <- sum(genes %in% acute_set)
  chronic_overlap <- sum(genes %in% chronic_set)
  #denominator is only genes that map to a cluster
  total_mapped    <- acute_overlap + chronic_overlap
  
  fisher_res <- fisher_test_overlap_genes(dox_daps_tr, dox_nondaps_tr, acute_set)
  
  tibble(
    ProteinSet  = prot_set,
    Category    = c("Acute", "Chronic"),
    Count       = c(acute_overlap, chronic_overlap),
    Fraction    = c(acute_overlap, chronic_overlap) / total_mapped,
    Significant = c(fisher_res$significant, "")
  )
})

plot_summary_plot2$Category <- factor(plot_summary_plot2$Category, 
                                      levels = c("Chronic", "Acute"))

label_positions_plot2 <- plot_summary_plot2 %>%
  filter(Category == "Acute") %>%
  mutate(y_pos = 1)

dap_barplot_plot2 <- ggplot(plot_summary_plot2, 
                            aes(x = ProteinSet, y = Fraction,
                                fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = label_positions_plot2 %>% filter(Significant != ""),
            aes(x = ProteinSet, y = y_pos, label = Significant),
            inherit.aes = FALSE, size = 6, fontface = "bold", color = "black") +
  geom_text(aes(label = paste0(round(Fraction * 100, 1), "")),
            position = position_stack(vjust = 0.5), size = 5, color = "white") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("Acute" = "#83C67C", "Chronic" = "#C499C6")) +
  labs(
    title = "Proportion of cluster-mapping DAPs and nonDAPs in Acute vs Chronic",
    x     = "Protein Set",
    y     = "Proportion of Cluster-Mapped Proteins",
    fill  = ""
  ) +
  theme_custom()

print(dap_barplot_plot2)

# save_plot(dap_barplot_plot2,
#           filename = "AcutevChronic_inDAPvnonDAP_Barplot_Proportion_EMP",
#           folder = output_folder)

#summary table
fisher_res <- fisher_test_overlap_genes(dox_daps_tr, dox_nondaps_tr, acute_set)

for (prot_set in c("DAP", "nonDAP")) {
  genes           <- if (prot_set == "DAP") dox_daps_tr else dox_nondaps_tr
  acute_overlap   <- sum(genes %in% acute_set)
  chronic_overlap <- sum(genes %in% chronic_set)
  total_mapped    <- acute_overlap + chronic_overlap

  cat(sprintf("%s - Acute: %.2f%% (%d/%d), 
              Chronic: %.2f%% (%d/%d), 
              Fisher p = %.5f\n",
              prot_set,
              100 * acute_overlap / total_mapped,   
              acute_overlap,   total_mapped,
              100 * chronic_overlap / total_mapped, 
              chronic_overlap, total_mapped,
              fisher_res$p_value))
}
```

### Add a category for DAPs/nonDAPs not present in either motif

```{r revised DAPs non DAPs chronic v acute plots, warning=FALSE, message=FALSE}
#bars = DAPs and nonDAPs
#fill = proportion of each protein set whose genes fall in Acute vs Chronic (vs Neither)
#fisher's test = are Acute genes over-represented among DAPs vs nonDAPs?
#logic: iterate over the two PROTEIN SETS, and for each one ask
#        "how many of these proteins come from Acute vs Chronic genes?"

#pre-calculate counts for the subtitle
dap_counts <- list(
  A = sum(dox_daps_tr %in% acute_set),
  C = sum(dox_daps_tr %in% chronic_set),
  N = length(dox_daps_tr) - (sum(dox_daps_tr %in% acute_set) + sum(dox_daps_tr %in% chronic_set))
)

nondap_counts <- list(
  A = sum(dox_nondaps_tr %in% acute_set),
  C = sum(dox_nondaps_tr %in% chronic_set),
  N = length(dox_nondaps_tr) - (sum(dox_nondaps_tr %in% acute_set) + sum(dox_nondaps_tr %in% chronic_set))
)

dynamic_subtitle <- paste0(
  "DAP Counts: | ACute ", dap_counts$A, " | Chronic: ", dap_counts$C, " | Neither: ", dap_counts$N,
  " \nnonDAP Counts: | Acute ", nondap_counts$A, " | Chronic: ", nondap_counts$C, " | Neither: ", nondap_counts$N
)

plot_summary_plot2 <- map_dfr(c("DAP", "nonDAP"), function(prot_set) {
  
  genes <- if (prot_set == "DAP") dox_daps_tr else dox_nondaps_tr
  
  acute_overlap   <- sum(genes %in% acute_set)
  chronic_overlap <- sum(genes %in% chronic_set)
  # Calculate genes that are in the protein set but neither motif
  neither_overlap <- length(genes) - (acute_overlap + chronic_overlap)
  #denominator is now the full set to account for "Neither"
  total_genes     <- length(genes)
  
  fisher_res <- fisher_test_overlap_genes(dox_daps_tr, dox_nondaps_tr, acute_set)
  
  tibble(
    ProteinSet  = prot_set,
    Category    = c("Acute", "Chronic", "Neither"),
    Count       = c(acute_overlap, chronic_overlap, neither_overlap),
    Fraction    = c(acute_overlap, chronic_overlap, neither_overlap) / total_genes,
    Significant = c(fisher_res$significant, "", "")
  )
})

plot_summary_plot2$Category <- factor(plot_summary_plot2$Category, 
                                      levels = c("Chronic", "Acute", "Neither"))

label_positions_plot2 <- plot_summary_plot2 %>%
  filter(Category == "Acute") %>%
  mutate(y_pos = 1)

dap_barplot_plot2 <- ggplot(plot_summary_plot2, 
                            aes(x = ProteinSet, y = Fraction,
                                fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = label_positions_plot2 %>% filter(Significant != ""),
            aes(x = ProteinSet, y = y_pos, label = Significant),
            inherit.aes = FALSE, size = 6, fontface = "bold", color = "black") +
  geom_text(aes(label = paste0(round(Fraction * 100, 1), "")),
            position = position_stack(vjust = 0.5), size = 5, color = "black") +
  scale_y_continuous(limits = c(0, 1.1), breaks = seq(0, 1, 0.2)) +
  scale_fill_manual(values = c("Acute" = "#83C67C", "Chronic" = "#C499C6", "Neither" = "#D3D3D3")) +
  labs(
    title = "Proportion of cluster-mapping DAPs and nonDAPs in Acute vs Chronic",
    subtitle = dynamic_subtitle,
    x     = "Protein Set",
    y     = "Proportion of Total Proteins",
    fill  = ""
  ) +
  theme_custom()

print(dap_barplot_plot2)

# save_plot(dap_barplot_plot2,
#            filename = "AcutevChronic_inDAPvnonDAP_Barplot_Proportion_v2_EMP",
#            folder = output_folder)

#summary table
fisher_res <- fisher_test_overlap_genes(dox_daps_tr, dox_nondaps_tr, acute_set)

for (prot_set in c("DAP", "nonDAP")) {
  genes           <- if (prot_set == "DAP") dox_daps_tr else dox_nondaps_tr
  acute_overlap   <- sum(genes %in% acute_set)
  chronic_overlap <- sum(genes %in% chronic_set)
  neither_overlap <- length(genes) - (acute_overlap + chronic_overlap)
  total_genes     <- length(genes)

  cat(sprintf("%s - Acute: %.2f%% (%d/%d), 
              Chronic: %.2f%% (%d/%d), 
              Neither: %.2f%% (%d/%d),
              Fisher p = %.5f\n",
              prot_set,
              100 * acute_overlap / total_genes,   
              acute_overlap,   total_genes,
              100 * chronic_overlap / total_genes,  
              chronic_overlap, total_genes,
              100 * neither_overlap / total_genes,
              neither_overlap, total_genes,
              fisher_res$p_value))
}
```



```{r DAPs vs DEGs comparison}

#load in DEGs for all timepoints 
DOX_24T_R <- read.csv("data/DE/Toptable_RUV_24T_final.csv") %>% 
  dplyr::select(-(X))
DOX_24R_R <- read.csv("data/DE/Toptable_RUV_24R_final.csv") %>% 
  dplyr::select(-(X))
DOX_144R_R <- read.csv("data/DE/Toptable_RUV_144R_final.csv") %>% 
  dplyr::select(-(X))

#extract Significant DEGs from these lists
DEGs_24T_R <- DOX_24T_R$Entrez_ID[DOX_24T_R$adj.P.Val < 0.05]
DEGs_24R_R <- DOX_24R_R$Entrez_ID[DOX_24R_R$adj.P.Val < 0.05]
DEGs_144R_R <- DOX_144R_R$Entrez_ID[DOX_144R_R$adj.P.Val < 0.05]

timepoint_DEGs <- list(
  "tx+0"   = DEGs_24T_R,
  "tx+24"   = DEGs_24R_R,
  "tx+144"  = DEGs_144R_R
)

#read in the DAPs as well

dap_list <- dox_daps_tr

#now I want to compare how many DAPs are DEGs at each respective timepoint
dap_overlap_summary <- purrr::map_dfr(names(timepoint_DEGs), function(tp) {
  
  deg_list <- unique(timepoint_DEGs[[tp]])
  
  overlap_count <- sum(deg_list %in% dap_list)
  deg_count     <- length(deg_list)
  dap_count     <- length(dap_list)
  
  tibble(
    Timepoint = tp,
    DAP_in_DEG = overlap_count,
    Total_DEGs    = deg_count,
    Total_DAPs    = dap_count
    )
})

print(dap_overlap_summary)

```

