---
title: "Figure 1 Analysis"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
  
}
```


# Libraries

Load necessary libraries for analysis.

```{r Necessary Libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(readr)
library(drc)
library(scales)
library(RImageJROI) #to read in ROI from ImageJ/Fiji
library(stringr)
library(purrr)
library(rstatix)
library(ggpubr)
library(kableExtra)
```

Define my custom plot theme

```{r Custom plot theme, echo=TRUE, message=FALSE, warning=FALSE}

# Define the custom theme
# plot_theme_custom <- function() {
#   theme_minimal() +
#     theme(
#       #line for x and y axis
#       axis.line = element_line(linewidth = 1,
#                                color = "black"),
# 
#       #axis ticks only on x and y, length standard
#       axis.ticks.x = element_line(color = "black",
#                                 linewidth = 1),
#       axis.ticks.y = element_line(color = "black",
#                                 linewidth = 1),
#       axis.ticks.length = unit(0.05, "in"),
# 
#       #text and font
#       axis.text = element_text(color = "black",
#                                family = "Arial",
#                                size = 7),
#       axis.title = element_text(color = "black",
#                                 family = "Arial",
#                                 size = 9),
#       legend.text = element_text(color = "black",
#                                  family = "Arial",
#                                  size = 7),
#       legend.title = element_text(color = "black",
#                                   family = "Arial",
#                                   size = 9),
#       plot.title = element_text(color = "black",
#                                 family = "Arial",
#                                 size = 10),
# 
#       #blank background and border
#       panel.background = element_blank(),
#       panel.border = element_blank(),
# 
#       #gridlines for alignment
#       panel.grid.major = element_line(color = "grey80", linewidth = 0.5),  #grey major grid for align in illus
#       panel.grid.minor = element_line(color = "grey90", linewidth = 0.5) #grey minor grid for align in illus
#     )
# }

# saveRDS(plot_theme_custom, "data/plot_theme_custom.RDS")

# theme_custom <- readRDS("data/plot_theme_custom.RDS")
theme_custom <- readRDS("data/plot_theme_custom_update.RDS")
```


Define a function to save plots as .pdf and .png

```{r pdf png saving function, echo=TRUE, message=FALSE, warning=FALSE}

save_plot <- function(plot, filename, 
                      folder = ".", 
                      width = 8, 
                      height = 6, 
                      units = "in", 
                      dpi = 300, 
                      add_date = TRUE) {
  if (missing(filename)) stop("Please provide a filename (without extension) for the plot.")

  date_str <- if (add_date) paste0("_", format(Sys.Date(), "%y%m%d")) else ""
  pdf_file <- file.path(folder, paste0(filename, date_str, ".pdf"))
  png_file <- file.path(folder, paste0(filename, date_str, ".png"))
  ggsave(filename = pdf_file, plot = plot, device = cairo_pdf, width = width, height = height, units = units, bg = "transparent")
  ggsave(filename = png_file, plot = plot, device = "png", width = width, height = height, units = units, dpi = dpi, bg = "transparent")
  message("Saved plot as Cairo PDF: ", pdf_file)
  message("Saved plot as PNG: ", png_file)
}

output_folder <- "C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/Fig pdfs"


#save plot function created
#to use: just define the plot name, filename_base, width, height
```

# Figure 1A - Experimental Schematic (no code)

This schematic did not require analysis.

# Figure 1B - Flow Cytometry Barplot

Data to create Figure 1B.

```{r Flow Cytometry Barplot, fig.width=3, fig.height=3, warning=FALSE, message=FALSE}
#Individuals for Flow Cytometry
#1 = 87-1 (F)
#2 = 78-1 (F)
#3 = 75-1 (F)
#4 = 84-1 (M)
#5 = 17-3 (M)
#6 = 90-1 (M)
#6R = 90-1R (M)

ind_col <- list(
  "1" = "#FF9F64",
  "2" = "#78EBDA",
  "3" = "#ADCD77",
  "4" = "#E6CC50",
  "5" = "#A68AC0",
  "6" = "#F16F90",
  "7" = "#C61E4E"
)

#read in csv with flow cytometry information
flow_dxr <- read_csv("C:/Users/emmap/OneDrive/Desktop/Ward Lab/Experiments/DXR Project/Flow/Recovery_flow_purity.csv",
                     show_col_types = FALSE)

# View(flow_dxr)

flow <- flow_dxr %>% column_to_rownames("Line")

#make a barplot

flow_dxr_df <- as.data.frame(flow_dxr)

#Plot
flow_plot_tnnt2 <- ggplot(flow_dxr_df, aes(x = Ind, y = Purity)) +
  geom_bar(stat = "identity", 
           fill = ind_col,
           width = 0.3 / (0.3 + 0.02)) +
  scale_y_continuous(limits = c(0, 100), 
                     expand = c(0, 0)) +
  scale_x_discrete(expand = expansion(mult = c(0.02, 0.02)))+
  labs(title = "iPSC-CM Purity",
       y = "TNNT2 Expression (%)", 
       x = "Individual") +
  theme_custom() +
  theme(strip.text.y = element_text(color = "white"))

print(flow_plot_tnnt2)

# save_plot(
#   plot = flow_plot_tnnt2,
#   filename = "FlowCytometry_TNNT2_withrep_EMP",
#   folder = output_folder,
#   width = 2.8,
#   height = 5.25
# )

```

# Figure 1C - Dose Response Curves

Data for Figure 1C.

```{r Full DRC dataset, results=FALSE, message=FALSE, warning=FALSE}
AllData <- read_excel("C:/Users/emmap/OneDrive/Desktop/Ward Lab/Experiments/DXR Project/DRCs/5FUProj_DRCTime_78-1_75-1_17-3_EMP_240226.xlsx",
     sheet = "Fin_Sheet_Analysis")
# View(AllData)

Ind2_78_data <- read_excel("C:/Users/emmap/OneDrive/Desktop/Ward Lab/Experiments/DXR Project/DRCs/5FUProj_DRCTime_78-1_75-1_17-3_EMP_240226.xlsx",
     sheet = "Fin_78_Sheet")
# View(Ind2_78_data)

#this is now Individual 2 after reordering individuals earlier - 78-1 F

Ind3_75_data <- read_excel("C:/Users/emmap/OneDrive/Desktop/Ward Lab/Experiments/DXR Project/DRCs/5FUProj_DRCTime_78-1_75-1_17-3_EMP_240226.xlsx",
     sheet = "Fin_75_Sheet")
# View(Ind3_75_data)

#this is now Ind3 since I reordered individuals - 75-1 F

Ind5_17_data <- read_excel("C:/Users/emmap/OneDrive/Desktop/Ward Lab/Experiments/DXR Project/DRCs/5FUProj_DRCTime_78-1_75-1_17-3_EMP_240226.xlsx",
     sheet = "Fin_17_Sheet")
# View(Ind5_17_data)

#this is still ind 5 according to my RNAseq - 17-3 M

```

## Filter for Time and Treatment Groups

```{r Filter Data, echo=FALSE, message=FALSE, warning=FALSE}
####all ind####
dox_all <- AllData %>% filter(Drug =="DOX")
dmso_all <- AllData %>% filter(Drug =="DMSO")

####Ind2 78-1####
dox_78 <- Ind2_78_data %>% filter(Drug=="DOX")
dmso_78 <- Ind2_78_data %>% filter(Drug=="DMSO")

####Ind3 75-1####
dox_75 <- Ind3_75_data %>% filter(Drug=="DOX")
dmso_75 <- Ind3_75_data %>% filter(Drug=="DMSO")

####Ind5 17-3####
dox_17 <- Ind5_17_data %>% filter(Drug=="DOX")
dmso_17 <- Ind5_17_data %>% filter(Drug=="DMSO")
```

```{r Filter Data Timepoint, echo = FALSE, message=FALSE, warning=FALSE}

####DOX Timepoints All####
dox24hrall <- dox_all %>% filter(FinTime == "24hr")

####DMSO VEH Timepoints All####
dmso24hrall <- dmso_all %>% filter(FinTime == "24hr")

#Filter these again by individual

####DOX Timepoints 75-1####
dox24hr75 <- dox_75 %>% filter(FinTime == "24hr")

####DMSO Timepoints 75-1####
dmso24hr75 <- dmso_75 %>% filter(FinTime == "24hr")

####DOX Timepoints 78-1####
dox24hr78 <- dox_78 %>% filter(FinTime == "24hr")

####DMSO Timepoints 78-1####
dmso24hr78 <- dmso_78 %>% filter(FinTime == "24hr")

####DOX Timepoints 17-3####
dox24hr17 <- dox_17 %>% filter(FinTime == "24hr")

####DMSO Timepoints 17-3####
dmso24hr17 <- dmso_17 %>% filter(FinTime == "24hr")

```

## Analysis of Raw Data

```{r Boxplots Raw Data 24hr, message=FALSE, warning=FALSE}

#####DOX Samples#####
#dox 78-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"RFLU by Concentration DOX - 78-1 24hr",cex = 2,font = 2)

boxplot(dox24hr78$Ind2_78_In ~ dox24hr78$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 78-1 Raw Data"
)

boxplot(dox24hr78$Ind2_78_Fin ~ dox24hr78$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 78-1 Raw Data"
)

#dox 75-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"RFLU by Concentration DOX - 75-1 24hr",cex = 2,font = 2)

boxplot(dox24hr75$Ind1_75_In ~ dox24hr75$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 75-1 Raw Data"
)

boxplot(dox24hr75$Ind1_75_Fin ~ dox24hr75$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 75-1 Raw Data"
)

#dox 17-3 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"RFLU by Concentration DOX - 17-3 24hr",cex = 2,font = 2)

boxplot(dox24hr17$Ind3_17_In ~ dox24hr17$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 17-3 Raw Data"
)

boxplot(dox24hr17$Ind3_17_Fin ~ dox24hr17$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 17-3 Raw Data"
)

#####DMSO samples#####
#DMSO 78-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"RFLU by Concentration DMSO - 78-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr78$Ind2_78_In ~ dmso24hr78$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 78-1"
)

boxplot(dmso24hr78$Ind2_78_Fin ~ dmso24hr78$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 78-1"
)

#DMSO 75-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"RFLU by Concentration DMSO - 75-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr75$Ind1_75_In ~ dmso24hr75$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 75-1"
)

boxplot(dmso24hr75$Ind1_75_Fin ~ dmso24hr75$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 75-1"
)

#DMSO 17-3 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"RFLU by Concentration DMSO - 17-3 24hr",cex = 2,font = 2)

boxplot(dmso24hr17$Ind3_17_In ~ dmso24hr17$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 17-3"
)

boxplot(dmso24hr17$Ind3_17_Fin ~ dmso24hr17$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 17-3"
)

```

Now that I've taken a look at the raw data, let's look at the analyzed data step by step

The following steps were taken in analysis:

1. Import raw fluorescence values (RFLU) from platereader, baseline 0hr and final timepoint

2. Subtract background wells (prestoblue only wells) by plate

3. Subtract average of empty wells (galactose only, with 50k cells/well) by plate

4. Divide all values by matched vehicle average (ie DOX 0.01 / DMSO 0.01)

5. Divide initial 0hr value from final timepoint (24, 48, 72) value 

## Remove Background Wells


```{r Boxplots BG removed 24hr, message=FALSE, warning=FALSE}
#####DOX Samples#####
#dox 78-1 no BG 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG - 78-1 24hr",cex = 2,font = 2)

boxplot(dox24hr78$Ind2_78_In_noBG ~ dox24hr78$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 78-1 no BG"
)

boxplot(dox24hr78$Ind2_78_Fin_noBG ~ dox24hr78$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 78-1 no BG"
)

#dox 75-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG - 75-1 24hr",cex = 2,font = 2)

boxplot(dox24hr75$Ind1_75_In_noBG ~ dox24hr75$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 75-1 no BG"
)

boxplot(dox24hr75$Ind1_75_Fin_noBG ~ dox24hr75$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 75-1 no BG"
)

#dox 17-3 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG - 17-3 24hr",cex = 2,font = 2)

boxplot(dox24hr17$Ind3_17_In_noBG ~ dox24hr17$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 17-3 no BG"
)

boxplot(dox24hr17$Ind3_17_Fin_noBG ~ dox24hr17$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 17-3 no BG"
)

#####DMSO samples#####
#DMSO 78-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG - 78-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr78$Ind2_78_In_noBG ~ dmso24hr78$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 78-1 no BG"
)

boxplot(dmso24hr78$Ind2_78_Fin_noBG ~ dmso24hr78$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 78-1 no BG"
)

#DMSO 75-1 no BG 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG - 75-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr75$Ind1_75_In_noBG ~ dmso24hr75$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 75-1 no BG"
)

boxplot(dmso24hr75$Ind1_75_Fin_noBG ~ dmso24hr75$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 75-1 no BG"
)

#DMSO 17-3 no BG 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG - 17-3 24hr",cex = 2,font = 2)

boxplot(dmso24hr17$Ind3_17_In_noBG ~ dmso24hr17$Conc, 
        ylim = c(0, 80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 17-3 no BG"
)

boxplot(dmso24hr17$Ind3_17_Fin_noBG ~ dmso24hr17$Conc, 
        ylim = c(0,80000), 
        ylab = "Cell Viability (RFLU)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 17-3 no BG"
)

```

Now let's plot the next step: subtracting the empty wells

## Subtract Empty Wells

```{r Boxplots Empty Wells Removed 24hr, message=FALSE, warning=FALSE}
#####DOX Samples#####
#dox 78-1 Empty Wells 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG Empty - 78-1 24hr",cex = 2,font = 2)

boxplot(dox24hr78$Ind2_78_In_noBG_Low ~ dox24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 78-1 no BG Empty"
)

boxplot(dox24hr78$Ind2_78_Fin_noBG_Low ~ dox24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 78-1 no BG"
)

#dox 75-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG Empty - 75-1 24hr",cex = 2,font = 2)

boxplot(dox24hr75$Ind1_75_In_noBG_Low ~ dox24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 75-1 no BG Empty"
)

boxplot(dox24hr75$Ind1_75_Fin_noBG_Low ~ dox24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 75-1 no BG Empty"
)

#dox 17-3 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG Empty - 17-3 24hr",cex = 2,font = 2)

boxplot(dox24hr17$Ind3_17_In_noBG_Low ~ dox24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 17-3 no BG Empty"
)

boxplot(dox24hr17$Ind3_17_Fin_noBG_Low ~ dox24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 17-3 no BG Empty"
)

#####DMSO samples#####
#DMSO 78-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG Empty - 78-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr78$Ind2_78_In_noBG_Low ~ dmso24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 78-1 no BG Empty"
)

boxplot(dmso24hr78$Ind2_78_Fin_noBG_Low ~ dmso24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 78-1 no BG Empty"
)

#DMSO 75-1 no BG Empty 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG Empty - 75-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr75$Ind1_75_In_noBG_Low ~ dmso24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 75-1 no BG Empty"
)

boxplot(dmso24hr75$Ind1_75_Fin_noBG_Low ~ dmso24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 75-1 no BG Empty"
)

#DMSO 17-3 no BG Empty 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG Empty - 17-3 24hr",cex = 2,font = 2)

boxplot(dmso24hr17$Ind3_17_In_noBG_Low ~ dmso24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 17-3 no BG Empty"
)

boxplot(dmso24hr17$Ind3_17_Fin_noBG_Low ~ dmso24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 17-3 no BG Empty"
)

```

Now let's plot the third step: dividing the matched vehicle values

## Divide by Matched Vehicle

```{r Boxplots Matched Vehicle Divided 24hr, message=FALSE, warning=FALSE}
#####DOX Samples#####
#dox 78-1 Veh 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG Empty Veh - 78-1 24hr",cex = 2,font = 2)

boxplot(dox24hr78$Ind2_78_In_noBG_Low_Veh ~ dox24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 78-1 no BG Empty Veh"
)

boxplot(dox24hr78$Ind2_78_Fin_noBG_Low_Veh ~ dox24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 78-1 no BG Veh"
)

#dox 75-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG Empty Veh - 75-1 24hr",cex = 2,font = 2)

boxplot(dox24hr75$Ind1_75_In_noBG_Low_Veh ~ dox24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 75-1 no BG Empty Veh"
)

boxplot(dox24hr75$Ind1_75_Fin_noBG_Low_Veh ~ dox24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 75-1 no BG Empty Veh"
)

#dox 17-3 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOX no BG Empty Veh - 17-3 24hr",cex = 2,font = 2)

boxplot(dox24hr17$Ind3_17_In_noBG_Low_Veh ~ dox24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 0hr 17-3 no BG Empty Veh"
)

boxplot(dox24hr17$Ind3_17_Fin_noBG_Low_Veh ~ dox24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DOXORUBICIN 24hr 17-3 no BG Empty Veh"
)

#####DMSO samples#####
#DMSO 78-1 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG Empty Veh - 78-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr78$Ind2_78_In_noBG_Low_Veh ~ dmso24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 78-1 no BG Empty Veh"
)

boxplot(dmso24hr78$Ind2_78_Fin_noBG_Low_Veh ~ dmso24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 78-1 no BG Empty Veh"
)

#DMSO 75-1 no BG Empty 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG Empty Veh - 75-1 24hr",cex = 2,font = 2)

boxplot(dmso24hr75$Ind1_75_In_noBG_Low_Veh ~ dmso24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 75-1 no BG Empty Veh"
)

boxplot(dmso24hr75$Ind1_75_Fin_noBG_Low_Veh ~ dmso24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 75-1 no BG Empty Veh"
)

#DMSO 17-3 no BG Empty 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO no BG Empty Veh - 17-3 24hr",cex = 2,font = 2)

boxplot(dmso24hr17$Ind3_17_In_noBG_Low_Veh ~ dmso24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 0hr 17-3 no BG Empty Veh"
)

boxplot(dmso24hr17$Ind3_17_Fin_noBG_Low_Veh ~ dmso24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "DMSO VEHICLE 24hr 17-3 no BG Empty Veh"
)

```

Now for the final step, dividing the final timepoint by the initial baseline reading at 0hr

## Divide Final Timepoint by Baseline

```{r Boxplots Final Data 24hr, message=FALSE, warning=FALSE}
#####DOX Samples#####
#dox All 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3, 1, 4), nrow = 2, ncol = 3), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DOXORUBICIN - 24hr",cex = 2,font = 2)

boxplot(dox24hr78$Ind2_78_Div ~ dox24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "78-1 Div"
)

boxplot(dox24hr75$Ind1_75_Div ~ dox24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "75-1 Div"
)

boxplot(dox24hr17$Ind3_17_Div ~ dox24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "17-3 Div"
)

#####DMSO samples#####
#DMSO All 24HR
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3, 1, 4), nrow = 2, ncol = 3), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5,0.5,"DMSO VEHICLE - 24hr",cex = 2,font = 2)

boxplot(dmso24hr78$Ind2_78_Div ~ dmso24hr78$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "78-1 Div"
)

boxplot(dmso24hr75$Ind1_75_Div ~ dmso24hr75$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "75-1 Div"
)

boxplot(dmso24hr17$Ind3_17_Div ~ dmso24hr17$Conc, 
        ylim = c(0, 1.5), 
        ylab = "Cell Viability (%)", 
        xlab = "Concentration (uM)", 
        main = "17-3 Div"
)
```

## Fit Models for DRC

```{r Data for Models DRCs, message=FALSE, warning=FALSE}
#these sheets are not numbered properly by individual because they're in order of experiments
dox24tox_75 <- dox24hr75 %>% mutate(response = Ind1_75_Div*100)
summary(dox24tox_75$response)

dox24tox_78 <- dox24hr78 %>% mutate(response = Ind2_78_Div*100)
summary(dox24tox_78$response)

dox24tox_17 <- dox24hr17 %>% mutate(response = Ind3_17_Div*100)
summary(dox24tox_17$response)

#Fit Models
doxtox24_78_model <- drm(response ~ Conc, data = dox24tox_78, fct = LL.4(fixed = c(NA, 0, 100, NA)))

doxtox24_75_model <- drm(response ~ Conc, data = dox24tox_75, fct = LL.4(fixed = c(NA, 0, 100, NA)))

doxtox24_17_model <- drm(response ~ Conc, data = dox24tox_17, fct = LL.4(fixed = c(NA, 0, 100, NA)))

#Add in labels
conc_labels <- c(0.01, 0.1, 0.5, 1, 10, 50, 100, 1000)
pct_labels <- c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

#extract LD50
ed75 <- ED(doxtox24_78_model, 50, interval = "delta")[1]
ed78 <- ED(doxtox24_75_model, 50, interval = "delta")[1]
ed17 <- ED(doxtox24_17_model, 50, interval = "delta")[1]

#find the median ld50 for plotting
median_ld50 <- median(c(ed78, ed75, ed17))

```

## Plot DRC

```{r Final plot with exports, message=FALSE, warning=FALSE}
# Define output folder and base filename
# output_folder
# base_name <- "viability_curve_scaled"

# Full file paths
# png_path <- file.path(output_folder, paste0(base_name, ".png"))
# svg_path <- file.path(output_folder, paste0(base_name, ".svg"))
# pdf_path <- file.path(output_folder, paste0(base_name, ".pdf"))

# New scaled dimensions (2x)
# plot_height <- 7.3212
# plot_width<- 3.8604
# plot_width <- 5.7906

#export as a png
# png(filename = png_path,
#     width = plot_width, height = plot_height,
#     units = "in", res = 300)

#plot code as above in finalized version
plot(doxtox24_78_model, type = "none", broken = TRUE,
     ylim = c(0,100), col = "#9ACCDE", lwd = 3,
     ylab = "Percent Viability (%)", xlab = "Concentration (uM)",
     main = "24hr Treatment", log = "x", axes = FALSE)
plot(doxtox24_75_model, add = TRUE, col = "#499FBD", lwd = 3, type = "none")
plot(doxtox24_17_model, add = TRUE, col = "#254F5E", lwd = 3, type = "none")

axis(1, at = conc_labels, labels = conc_labels, las = 2, cex.axis = 1)
axis(2, at = pct_labels, labels = pct_labels, las = 2, cex.axis = 0.9)

abline(v = 0.5, col = "grey50", lty = 2, lwd = 2)
abline(v = median_ld50, col = "red", lty = 3, lwd = 2)
text(x = median_ld50, y = 99,
     labels = paste0("Median LD50"),
     col = "red", pos = 4, cex = 1, font = 2)

legend("bottomleft",
       legend = c(
         paste0("Ind3 (LD50 = ", round(ed78, 2), " µM)"),
         paste0("Ind4 (LD50 = ", round(ed75, 2), " µM)"),
         paste0("Ind5 (LD50 = ", round(ed17, 2), " µM)")
       ),
       col = c("#9ACCDE", "#499FBD", "#254F5E"),
       lwd = 3, lty = c(1,1,1), cex = 1,
       title = "Individuals")

# dev.off()

#export as svg
# svg(filename = svg_path,
#     width = plot_width, height = plot_height)

#repeat plotting code with same adjustments
# plot(doxtox24_78_model, type = "none", broken = TRUE,
#      ylim = c(0,100), col = "#9ACCDE", lwd = 3,
#      ylab = "Percent Viability (%)", xlab = "Concentration (uM)",
#      main = "24hr Treatment", log = "x", axes = FALSE)
# plot(doxtox24_75_model, add = TRUE, col = "#499FBD", lwd = 3, type = "none")
# plot(doxtox24_17_model, add = TRUE, col = "#254F5E", lwd = 3, type = "none")
# axis(1, at = conc_labels, labels = conc_labels, las = 2, cex.axis = 1)
# axis(2, at = pct_labels, labels = pct_labels, las = 2, cex.axis = 0.9)
# abline(v = 0.5, col = "grey50", lty = 2, lwd = 2)
# abline(v = median_ld50, col = "red", lty = 3, lwd = 2)
# text(x = median_ld50, y = 99,
#      labels = paste0("Median LD50"),
#      col = "red", pos = 4, cex = 1, font = 2)
# legend("bottomleft",
#        legend = c(
#          paste0("Ind3 (LD50 = ", round(ed78, 2), " µM)"),
#          paste0("Ind4 (LD50 = ", round(ed75, 2), " µM)"),
#          paste0("Ind5 (LD50 = ", round(ed17, 2), " µM)")
#        ),
#        col = c("#9ACCDE", "#499FBD", "#254F5E"),
#        lwd = 3, lty = c(1,1,1), cex = 1,
#        title = "Individuals")

# dev.off()

#export as pdf
# pdf(file = pdf_path,
#     width = plot_width, height = plot_height)

#repeat plotting code again to print
# plot(doxtox24_78_model, type = "none", broken = TRUE,
#      ylim = c(0,100), col = "#9ACCDE", lwd = 3,
#      ylab = "Percent Viability (%)", xlab = "Concentration (uM)",
#      main = "24hr Treatment", log = "x", axes = FALSE)
# plot(doxtox24_75_model, add = TRUE, col = "#499FBD", lwd = 3, type = "none")
# plot(doxtox24_17_model, add = TRUE, col = "#254F5E", lwd = 3, type = "none")
# axis(1, at = conc_labels, labels = conc_labels, las = 2, cex.axis = 1)
# axis(2, at = pct_labels, labels = pct_labels, las = 2, cex.axis = 0.9)
# abline(v = 0.5, col = "grey50", lty = 2, lwd = 2)
# abline(v = median_ld50, col = "red", lty = 3, lwd = 2)
# text(x = median_ld50, y = 99,
#      labels = paste0("Median LD50"),
#      col = "red", pos = 4, cex = 1, font = 2)
# legend("bottomleft",
#        legend = c(
#          paste0("Ind3 (LD50 = ", round(ed78, 2), " µM)"),
#          paste0("Ind4 (LD50 = ", round(ed75, 2), " µM)"),
#          paste0("Ind5 (LD50 = ", round(ed17, 2), " µM)")
#        ),
#        col = c("#9ACCDE", "#499FBD", "#254F5E"),
#        lwd = 3, lty = c(1,1,1), cex = 1,
#        title = "Individuals")

# dev.off()

#saved code for saving the plots for posterity!

```

# Supplementary Figure 10 - Timepoint-specific iPSC-CM Viability Assay

```{r AB JG Viability Assay Boxplots, warning=FALSE, message=FALSE}
#My colleagues performed a cell viability assay following the recovery timepoints
#tx+0, tx+24, tx+144

#import the data
dxr_viability <- read_excel("C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/DXR_DRC_Time_Combined_75JG_87AB_EMP_251114.xlsx", 
    sheet = "EMP_noBG_normVEH")
# View(dxr_viability)

#separate out the data
dox_via <- dxr_viability %>% filter(Drug=="DOX")
veh_via <- dxr_viability %>% filter(Drug=="VEH")
untx_via <- dxr_viability %>% filter(Drug=="CMM")

#now separate by individual
dox_ind1 <- dox_via %>% filter(Individual == "Ind1") #87-1
veh_ind1 <- veh_via %>% filter(Individual == "Ind1")
untx_ind1 <- untx_via %>% filter(Individual == "Ind1")

dox_ind3 <- dox_via %>% filter(Individual == "Ind3") #75-1
veh_ind3 <- veh_via %>% filter(Individual == "Ind3")
untx_ind3 <- untx_via %>% filter(Individual == "Ind3")

#enforce factor order for plotting 
dox_ind1$Time <- factor(dox_ind1$Time, levels = c("tx+0", "tx+24", "tx+144"))
dox_ind3$Time <- factor(dox_ind3$Time, levels = c("tx+0", "tx+24", "tx+144"))
veh_ind1$Time <- factor(veh_ind1$Time, levels = c("tx+0", "tx+24", "tx+144"))
veh_ind3$Time <- factor(veh_ind3$Time, levels = c("tx+0", "tx+24", "tx+144"))
untx_ind1$Time <- factor(untx_ind1$Time, levels = c("tx+0", "tx+24", "tx+144"))
untx_ind3$Time <- factor(untx_ind3$Time, levels = c("tx+0", "tx+24", "tx+144"))

```

## Boxplots of Raw Data

```{r Boxplots Raw Data all Timepoints, message=FALSE, warning=FALSE}

##DOX base vs fin Ind1
#create the layout for plots to be side by side for each individiual
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - DOX Ind1 87-1", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1$RFLU_base ~ dox_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind1 Baseline"
)

#ind1 final
boxplot(dox_ind1$RFLU_fin ~ dox_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind1 Final"
)

##VEHICLE base vs fin Ind1
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - VEH Ind1 87-1", cex = 1, font = 2)

#ind1 veh baseline
boxplot(veh_ind1$RFLU_base ~ veh_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind1 Baseline"
)

#ind1 veh final
boxplot(veh_ind1$RFLU_fin ~ veh_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind1 Final"
)

##Untx base vs fin Ind1
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - UTx Ind1 87-1", cex = 1, font = 2)

#ind1 untx baseline
boxplot(untx_ind1$RFLU_base ~ untx_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind1 Baseline"
)

#ind1 untx final
boxplot(untx_ind1$RFLU_fin ~ untx_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind1 Final"
)

###individual 3
##DOX base vs fin Ind3
#create the layout for plots to be side by side for each individiual
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - DOX Ind3 75-1", cex = 1, font = 2)

#Ind3 75-1
boxplot(dox_ind3$RFLU_base ~ dox_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind3 Baseline"
)

#ind3 final
boxplot(dox_ind3$RFLU_fin ~ dox_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind3 Final"
)

##VEHICLE base vs fin Ind3
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - VEH Ind3 75-1", cex = 1, font = 2)

#ind3 veh baseline
boxplot(veh_ind3$RFLU_base ~ veh_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind3 Baseline"
)

#ind3 veh final
boxplot(veh_ind3$RFLU_fin ~ veh_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind3 Final"
)

##Untx base vs fin Ind3
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - UTx Ind3 75-1", cex = 1, font = 2)

#ind3 untx baseline
boxplot(untx_ind3$RFLU_base ~ untx_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind3 Baseline"
)

#ind1 untx final
boxplot(untx_ind3$RFLU_fin ~ untx_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind3 Final"
)

```


## Remove Background Fluorescence

```{r RFLU noBG boxplots, message=FALSE, warning=FALSE}
#now remove the background fluorescence using the PrestoBlue only wells

##Individual 1 87-1
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint noBG - DOX Ind1 87-1", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1$RFLU_base_noBG ~ dox_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind1 Baseline noBG"
)

#ind1 final
boxplot(dox_ind1$RFLU_fin_noBG ~ dox_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind1 Final noBG"
)

par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint noBG - VEH Ind1 87-1", cex = 1, font = 2)

#ind1 veh baseline
boxplot(veh_ind1$RFLU_base_noBG ~ veh_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind1 Baseline noBG"
)

#ind1 veh final
boxplot(veh_ind1$RFLU_fin_noBG ~ veh_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind1 Final noBG"
)


##untx ind1 base vs fin
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint noBG - UTx Ind1 87-1", cex = 1, font = 2)

#Ind1 87-1
boxplot(untx_ind1$RFLU_base_noBG ~ untx_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind1 Baseline noBG"
)

#ind1 final
boxplot(untx_ind1$RFLU_fin_noBG ~ untx_ind1$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind1 Final noBG"
)

###individual 3
##DOX base vs fin Ind3
#create the layout for plots to be side by side for each individiual
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint noBG - DOX Ind3 75-1", cex = 1, font = 2)

#Ind3 75-1
boxplot(dox_ind3$RFLU_base_noBG ~ dox_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind3 Baseline noBG"
)

#ind3 final
boxplot(dox_ind3$RFLU_fin_noBG ~ dox_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind3 Final noBG"
)

##VEHICLE base vs fin Ind3
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint noBG - VEH Ind3 75-1", cex = 1, font = 2)

#ind3 veh baseline
boxplot(veh_ind3$RFLU_base_noBG ~ veh_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind3 Baseline noBG"
)

#ind3 veh final
boxplot(veh_ind3$RFLU_fin_noBG ~ veh_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind3 Final noBG"
)

##untx ind3 base vs fin
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint noBG - UTx Ind3 75-1", cex = 1, font = 2)

#Ind3 87-1
boxplot(untx_ind3$RFLU_base_noBG ~ untx_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind3 Baseline noBG"
)

#ind3 final
boxplot(untx_ind3$RFLU_fin_noBG ~ untx_ind3$Time,
        ylim = c(0, 150000),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind3 Final noBG"
)


```


## Normalize by CMM-only wells

```{r subtract CMM only wells, warning=FALSE, message=FALSE}
#now, I would like to normalize my data by removing any additional background from CMM-only wells seeded with iPSCs

##Individual 1 87-1
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Time noBG normUTX - DOX Ind1", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1$RFLU_base_noBG_normUTX ~ dox_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind1 Base noBG normUTX"
)

#ind1 final
boxplot(dox_ind1$RFLU_fin_noBG_normUTX ~ dox_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind1 Fin noBG normUTX"
)

par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Time noBG normUTX - VEH Ind1", cex = 1, font = 2)

#ind1 veh baseline
boxplot(veh_ind1$RFLU_base_noBG_normUTX ~ veh_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind1 Base noBG normUTX"
)

#ind1 veh final
boxplot(veh_ind1$RFLU_fin_noBG_normUTX ~ veh_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind1 Fin noBG normUTX"
)

##untx set
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Time noBG normUTX - UTx Ind1", cex = 1, font = 2)

#Ind1 87-1
boxplot(untx_ind1$RFLU_base_noBG_normUTX ~ untx_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind1 Base noBG normUTX"
)

#ind1 final
boxplot(untx_ind1$RFLU_fin_noBG_normUTX ~ untx_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind1 Fin noBG normUTX"
)

###individual 3
##DOX base vs fin Ind3
#create the layout for plots to be side by side for each individiual
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Time noBG normUTX - DOX Ind3", cex = 1, font = 2)

#Ind3 75-1
boxplot(dox_ind3$RFLU_base_noBG_normUTX ~ dox_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind3 Base noBG normUTX"
)

#ind3 final
boxplot(dox_ind3$RFLU_fin_noBG_normUTX ~ dox_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX Ind3 Fin noBG normUTX"
)

##VEHICLE base vs fin Ind3
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Time noBG normUTX - VEH Ind3", cex = 1, font = 2)

#ind3 veh baseline
boxplot(veh_ind3$RFLU_base_noBG_normUTX ~ veh_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind3 Base noBG normUTX"
)

#ind3 veh final
boxplot(veh_ind3$RFLU_fin_noBG_normUTX ~ veh_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "VEH Ind3 Fin noBG normUTX"
)

##untx set
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Time noBG normUTX - UTx Ind3", cex = 1, font = 2)

#Ind3 75-1
boxplot(untx_ind3$RFLU_base_noBG_normUTX ~ untx_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind3 Base noBG normUTX"
)

#ind3 final
boxplot(untx_ind3$RFLU_fin_noBG_normUTX ~ untx_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "UnTx Ind3 Fin noBG normUTX"
)


```


## Divide Treated by Vehicle

```{r div by Vehicle, message=FALSE, warning=FALSE}

#now, I would like to normalize my data further by dividing my treated samples by volume-matched vehicle samples

##Individual 1 87-1
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - DOX/VEH Ind1", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1$RFLU_base_noBG_normUTX_divVEH ~ dox_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind1 Baseline"
)

#ind1 final
boxplot(dox_ind1$RFLU_fin_noBG_normUTX_divVEH ~ dox_ind1$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind1 Final"
)

###individual 3
##DOX base vs fin Ind3
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - DOX/VEH Ind3", cex = 1, font = 2)

#Ind3 75-1
boxplot(dox_ind3$RFLU_base_noBG_normUTX_divVEH ~ dox_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind3 Baseline"
)

#ind3 final
boxplot(dox_ind3$RFLU_fin_noBG_normUTX_divVEH ~ dox_ind3$Time,
        ylim = c(0, 2),
        ylab = "Cell Viability (RFLU)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind3 Final"
)

```


## Divide Final Timepoint by Baseline

```{r div final by baseline, warning=FALSE, message=FALSE}
#now, as the final step I will divide the final timepoint by the baseline measurement to get a 0-100% relative cell viability

#enforce factoring order
dox_ind1$Time <- factor(dox_ind1$Time, levels = c("tx+0", "tx+24", "tx+144"))
dox_ind3$Time <- factor(dox_ind3$Time, levels = c("tx+0", "tx+24", "tx+144"))

untx_ind1$Time <- factor(untx_ind1$Time, levels = c("tx+0", "tx+24", "tx+144"))
untx_ind3$Time <- factor(untx_ind3$Time, levels = c("tx+0", "tx+24", "tx+144"))

par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1$RFLU_ ~ dox_ind1$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind1"
)

#ind3 final
boxplot(dox_ind3$RFLU_ ~ dox_ind3$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind3"
)


#untx version
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - UnTx", cex = 1, font = 2)

#Ind1 87-1
boxplot(untx_ind1$CMM_ ~ untx_ind1$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind1 UnTx"
)

#ind3 final
boxplot(untx_ind3$CMM_ ~ untx_ind3$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind3 UnTx"
)

#untx non-norm by CMM version
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint - UnTx nn", cex = 1, font = 2)

#Ind1 87-1
boxplot(untx_ind1$CMM_nn ~ untx_ind1$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind1 UnTx"
)

#ind3 final
boxplot(untx_ind3$CMM_nn ~ untx_ind3$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind3 UnTx"
)


```


## Plot Final Viability Assay

```{r move to a ggplot object for cleaner plots, message=FALSE, warning=FALSE}
#now, plot these individuals together in one boxplot

#create dox_combined
dox_combined <- bind_rows(dox_ind1, dox_ind3)

#define colors
ind_via_cols <- c("Ind1" = "#ffa364", "Ind3" = "#a4d177")
time_col <- c(
  "tx+0"   = "#005A4C",
  "tx+24"  = "#328477",
  "tx+144" = "#8FB9B1")

#factor for ordering
dox_combined$Time <- factor(dox_combined$Time,
                             levels = c("tx+0", "tx+24", "tx+144"))

#perform paired t-test between each timepoint with BH correction
ttest_dox <- dox_combined %>%
  pairwise_t_test(RFLU_ ~ Time,
                  paired = TRUE,
                  p.adjust.method = "BH") %>%
  add_xy_position(x = "Time") %>%
  mutate(
    p.signif = case_when(
      p.adj < 0.001 ~ "***",
      p.adj < 0.01  ~ "**",
      p.adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )

#create the plot
plot_dox_viability <- ggplot(dox_combined,
                              aes(x = Time,
                                  y = RFLU_)) +
  geom_boxplot(aes(fill = Time),
               show.legend = FALSE,
               outlier.shape = NA) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 1,
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_x_discrete(limits = c("tx+0", "tx+24", "tx+144")) +
  scale_color_manual(values = ind_via_cols) +
  scale_fill_manual(values = time_col) +
  stat_pvalue_manual(ttest_dox,
                     label = "p.signif",
                     xmin = "xmin",
                     xmax = "xmax",
                     y.position = "y.position",
                     tip.length = 0.02) +
  labs(title = "DOX Cell Viability by Timepoint",
       x = "Timepoint",
       y = "Relative Cell Viability (%)") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dox_viability

# save_plot(plot_dox_viability,
#           filename = "Supp10_Viability_ABJG_Timepoints_EMP",
#           folder = output_folder)

#print a summary table with the statistics information for later
ttest_dox %>%
  dplyr::select(group1, group2, n1, n2, statistic, df, p, p.adj, p.signif) %>%
  mutate(
    p     = formatC(p,     format = "e", digits = 3),
    p.adj = formatC(p.adj, format = "e", digits = 3),
    statistic = round(statistic, 3)
  ) %>%
  dplyr::rename(
    "Group 1"         = group1,
    "Group 2"         = group2,
    "n1"              = n1,
    "n2"              = n2,
    "t"               = statistic,
    "df"              = df,
    "p (raw)"         = p,
    "p (adj, BH)"     = p.adj,
    "Sig."            = p.signif
  ) %>%
  kable(align = "c", caption = "DOX Viability Statistical Summary (Paired t-test, BH Adjusted)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

```

## Untreated Final Plot

```{r untreated final plot, warning=FALSE, message=FALSE}
#create untx_combined
untx_combined <- bind_rows(untx_ind1, untx_ind3)

#define colors
ind_via_cols <- c("Ind1" = "#ffa364", "Ind3" = "#a4d177")
time_col <- c(
  "tx+0"   = "#005A4C",
  "tx+24"  = "#328477",
  "tx+144" = "#8FB9B1")

#factor for ordering
untx_combined$Time <- factor(untx_combined$Time,
                             levels = c("tx+0", "tx+24", "tx+144"))

#perform paired t-test between each timepoint with BH correction
ttest_untx <- untx_combined %>%
  pairwise_t_test(CMM_ ~ Time,
                  paired = TRUE,
                  p.adjust.method = "BH") %>%
  add_xy_position(x = "Time") %>%
  mutate(
    p.signif = case_when(
      p.adj < 0.001 ~ "***",
      p.adj < 0.01  ~ "**",
      p.adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )

#create the plot
plot_untx_viability <- ggplot(untx_combined,
                              aes(x = Time,
                                  y = CMM_)) +
  geom_boxplot(aes(fill = Time),
               show.legend = FALSE,
               outlier.shape = NA) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 1,
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_x_discrete(limits = c("tx+0", "tx+24", "tx+144")) +
  scale_color_manual(values = ind_via_cols) +
  scale_fill_manual(values = time_col) +
  stat_pvalue_manual(ttest_dox,
                     label = "p.signif",
                     xmin = "xmin",
                     xmax = "xmax",
                     y.position = "y.position",
                     tip.length = 0.02) +
  labs(title = "UnTx Cell Viability by Timepoint",
       x = "Timepoint",
       y = "Relative Cell Viability (%)") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_untx_viability

# save_plot(plot_untx_viability,
#           filename = "UnTx_Viability_ABJG_Timepoints_EMP",
#           folder = output_folder)

#print a summary table with the statistics information for later
ttest_untx %>%
  dplyr::select(group1, group2, n1, n2, statistic, df, p, p.adj, p.signif) %>%
  mutate(
    p     = formatC(p,     format = "e", digits = 3),
    p.adj = formatC(p.adj, format = "e", digits = 3),
    statistic = round(statistic, 3)
  ) %>%
  dplyr::rename(
    "Group 1"         = group1,
    "Group 2"         = group2,
    "n1"              = n1,
    "n2"              = n2,
    "t"               = statistic,
    "df"              = df,
    "p (raw)"         = p,
    "p (adj, BH)"     = p.adj,
    "Sig."            = p.signif
  ) %>%
  kable(align = "c", caption = "UnTx Viability Statistical Summary (Paired t-test, BH Adjusted)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)


####non-normalized by UnTx version
#perform paired t-test between each timepoint with BH correction
ttest_untx_nonorm <- untx_combined %>%
  pairwise_t_test(CMM_nn ~ Time,
                  paired = TRUE,
                  p.adjust.method = "BH") %>%
  add_xy_position(x = "Time") %>%
  mutate(
    p.signif = case_when(
      p.adj < 0.001 ~ "***",
      p.adj < 0.01  ~ "**",
      p.adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )

#create the plot
plot_untx_viability_nonorm <- ggplot(untx_combined,
                              aes(x = Time,
                                  y = CMM_nn)) +
  geom_boxplot(aes(fill = Time),
               show.legend = FALSE,
               outlier.shape = NA) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 1,
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_x_discrete(limits = c("tx+0", "tx+24", "tx+144")) +
  scale_color_manual(values = ind_via_cols) +
  scale_fill_manual(values = time_col) +
  stat_pvalue_manual(ttest_dox,
                     label = "p.signif",
                     xmin = "xmin",
                     xmax = "xmax",
                     y.position = "y.position",
                     tip.length = 0.02) +
  labs(title = "UnTx Cell Viability by Timepoint",
       x = "Timepoint",
       y = "Relative Cell Viability (%)") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_untx_viability_nonorm

# save_plot(plot_untx_viability_nonorm,
#           filename = "UnTx_nonNormUnTx_Viability_ABJG_Timepoints_EMP",
#           folder = output_folder)

#print a summary table with the statistics information for later
ttest_untx_nonorm %>%
  dplyr::select(group1, group2, n1, n2, statistic, df, p, p.adj, p.signif) %>%
  mutate(
    p     = formatC(p,     format = "e", digits = 3),
    p.adj = formatC(p.adj, format = "e", digits = 3),
    statistic = round(statistic, 3)
  ) %>%
  dplyr::rename(
    "Group 1"         = group1,
    "Group 2"         = group2,
    "n1"              = n1,
    "n2"              = n2,
    "t"               = statistic,
    "df"              = df,
    "p (raw)"         = p,
    "p (adj, BH)"     = p.adj,
    "Sig."            = p.signif
  ) %>%
  kable(align = "c", caption = "UnTx Viability Statistical Summary (Paired t-test, BH Adjusted)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)


```


# Viability Assay Analysis - no normalization by CMM

```{r viablity assay noCMM set, message=FALSE, warning=FALSE}

#import the data from a different sheet on the same excel sheet
dxr_viability_2 <- read_excel("C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/DXR_DRC_Time_Combined_75JG_87AB_EMP_251114.xlsx", 
    sheet = "EMP_noBG_UTXnonorm_divVEH")

#now go ahead and follow the same workflow for these samples
#these are not normalized by the untreated wells (UnTx), also called CMM

#separate out the data
dox_via_2 <- dxr_viability_2 %>% filter(Drug=="DOX")
veh_via_2 <- dxr_viability_2 %>% filter(Drug=="VEH")
untx_via_2 <- dxr_viability_2 %>% filter(Drug=="CMM")

#now separate by individual
dox_ind1_2 <- dox_via_2 %>% filter(Individual == "Ind1") #87-1
veh_ind1_2 <- veh_via_2 %>% filter(Individual == "Ind1")
untx_ind1_2 <- untx_via_2 %>% filter(Individual == "Ind1")

dox_ind3_2 <- dox_via_2 %>% filter(Individual == "Ind3") #75-1
veh_ind3_2 <- veh_via_2 %>% filter(Individual == "Ind3")
untx_ind3_2 <- untx_via_2 %>% filter(Individual == "Ind3")

#enforce factor order for plotting 
dox_ind1_2$Time <- factor(dox_ind1_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
dox_ind3_2$Time <- factor(dox_ind3_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
veh_ind1_2$Time <- factor(veh_ind1_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
veh_ind3_2$Time <- factor(veh_ind3_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
untx_ind1_2$Time <- factor(untx_ind1_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
untx_ind3_2$Time <- factor(untx_ind3_2$Time, levels = c("tx+0", "tx+24", "tx+144"))

```


## Step 1 - Raw Data Boxplots (no change from above)

## Step 2 - remove background from presto blue-only wells (no change from above)

## Step 3 - Divide DOX/VEH (without normalizing by UnTx wells)

```{r no norm DOX div VEH, warning=FALSE, message=FALSE}
#now for this step, I have directly divided our treated wells by vehicle wells

#now, I would like to normalize my data further by dividing my treated samples by volume-matched vehicle samples

##Individual 1 87-1
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "DOX/VEH Ind1 (no UnTx norm)", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1_2$RFLU_base_noBG_divVEH ~ dox_ind1_2$Time,
        ylim = c(0, 1.5),
        ylab = "Cell Viability (Rel)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind1 Baseline"
)

#ind1 final
boxplot(dox_ind1_2$RFLU_fin_noBG_divVEH ~ dox_ind1_2$Time,
        ylim = c(0, 1.5),
        ylab = "Cell Viability (Rel)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind1 Final"
)

###individual 3
##DOX base vs fin Ind3
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "DOX/VEH Ind3 (no UnTx norm)", cex = 1, font = 2)

#Ind3 75-1
boxplot(dox_ind3_2$RFLU_base_noBG_divVEH ~ dox_ind3_2$Time,
        ylim = c(0, 1.5),
        ylab = "Cell Viability (Rel)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind3 Baseline"
)

#ind3 final
boxplot(dox_ind3_2$RFLU_fin_noBG_divVEH ~ dox_ind3_2$Time,
        ylim = c(0, 1.5),
        ylab = "Cell Viability (Rel)",
        xlab = "Timepoint",
        main = "DOX/VEH Ind3 Final"
)

```

## Step 4 - Divide Final by Baseline timepoints

```{r final div baseline DOX div VEH no UnTx normalization, message=FALSE, warning=FALSE}
#now plot the final timepoint divided by the baseline timepoint to get the final relative cell viability (%)

#enforce factoring order
dox_ind1_2$Time <- factor(dox_ind1_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
dox_ind3_2$Time <- factor(dox_ind3_2$Time, levels = c("tx+0", "tx+24", "tx+144"))

untx_ind1_2$Time <- factor(untx_ind1_2$Time, levels = c("tx+0", "tx+24", "tx+144"))
untx_ind3_2$Time <- factor(untx_ind3_2$Time, levels = c("tx+0", "tx+24", "tx+144"))

par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint: DOX/VEH & Fin/Base", cex = 1, font = 2)

#Ind1 87-1
boxplot(dox_ind1_2$RFLU_noBG_divVEH_divTime_per ~ dox_ind1_2$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind1"
)

#ind3 final
boxplot(dox_ind3_2$RFLU_noBG_divVEH_divTime_per ~ dox_ind3_2$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind3"
)


#now do statistical testing between conditions for each individual (t-tests)
run_pairwise_ttests <- function(data, metric_col, individual_label, set_label) {
  timepoints <- list(
    c("tx+0", "tx+24"),
    c("tx+0", "tx+144"),
    c("tx+24", "tx+144")
  )
  purrr::map_dfr(timepoints, function(tp) {
    sub <- data %>% filter(Time %in% tp)
    tt  <- t.test(sub[[metric_col]] ~ sub$Time)
    tibble(
      Individual = individual_label,
      Set        = set_label,
      Comparison = paste(tp[1], "vs", tp[2]),
      t          = round(tt$statistic, 3),
      df         = round(tt$parameter, 1),
      p_value    = tt$p.value
    )
  })
}

#run for each individual and bind
ttest_dox_results <- bind_rows(
  run_pairwise_ttests(dox_ind1_2, "RFLU_noBG_divVEH_divTime_per", "Ind1", "DOX"),
  run_pairwise_ttests(dox_ind3_2, "RFLU_noBG_divVEH_divTime_per", "Ind3", "DOX")
) %>%
  mutate(
    p_adj  = p.adjust(p_value, method = "BH"),
    signif = case_when(
      p_adj < 0.001 ~ "***",
      p_adj < 0.01  ~ "**",
      p_adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )

ttest_dox_results %>%
  mutate(
    p_value = formatC(p_value, format = "e", digits = 3),
    p_adj   = formatC(p_adj,   format = "e", digits = 3)
  ) %>%
  dplyr::rename(
    "Sig." = signif, 
    "p (raw)" = p_value, 
    "p (adj, BH)" = p_adj
  ) %>%
  kable(align = "c", 
        caption = "DOX Viability: Pairwise T-tests by Individual (BH Adjusted)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)


#untx version
par(mar = c(3.0, 3.0, 1.2, 1), mgp = c(2, .7, 0))
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2), heights = c(0.5, 2, 2, 2))
#layout.show(3)
plot.new()
text(0.5, 0.5, "Viability by Timepoint: Fin/Base UnTx Samples", cex = 1, font = 2)

#Ind1 87-1
boxplot(untx_ind1_2$RFLU_noBG_CMMdivTime_per ~ untx_ind1_2$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind1 UnTx"
)

#ind3 final
boxplot(untx_ind3_2$RFLU_noBG_CMMdivTime_per ~ untx_ind3_2$Time,
        ylim = c(0, 150),
        ylab = "Percent Viability (%)",
        xlab = "Timepoint",
        main = "Ind3 UnTx"
)

#now do the same stat testing between the untreated samples by individual

ttest_untx_results <- bind_rows(
  run_pairwise_ttests(untx_ind1_2, "RFLU_noBG_CMMdivTime_per", "Ind1", "UnTx"),
  run_pairwise_ttests(untx_ind3_2, "RFLU_noBG_CMMdivTime_per", "Ind3", "UnTx")
) %>%
  mutate(
    p_adj  = p.adjust(p_value, method = "BH"),
    signif = case_when(
      p_adj < 0.001 ~ "***",
      p_adj < 0.01  ~ "**",
      p_adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )

ttest_untx_results %>%
  mutate(
    p_value = formatC(p_value, format = "e", digits = 3),
    p_adj   = formatC(p_adj,   format = "e", digits = 3)
  ) %>%
  dplyr::rename(
    "Sig." = signif, 
    "p (raw)" = p_value, 
    "p (adj, BH)" = p_adj
  ) %>%
  kable(align = "c", 
        caption = "UnTx Viability: Pairwise T-tests by Individual (BH Adjusted)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)

```

# Figure 1D - DNA Damage Quantification

yH2AX Quantification (Immunofluorescence Staining).

## Fiji/ImageJ Macros

Macros used in ImageJ/Fiji to identify nuclear ROIs from yH2AX IF staining.

```{r Macros Run in Fiji/ImageJ, message=FALSE, warning=FALSE}
#these are the macros that I made for ImageJ/Fiji and are included for reference only
#they will not run in R

#DAPI ROI generator macro to split DAPI channel and map nuclei
# // === STEP 1: Define folders ===
# folder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 84-1\\rgb quant\\";
# saveRoiFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 84-1\\outline new\\";
# saveRedFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 84-1\\TxRed new\\";
# 
# list = getFileList(folder);
# 
# for (i = 0; i < list.length; i++) {
#     if (endsWith(list[i], ".tif")) {
#         print("----------------------------------------------------");
#         print("Processing file: " + list[i]);
# 
#         // Open composite image
#         open(folder + list[i]);
#         originalTitle = getTitle();
#         baseName = replace(list[i], ".tif", "");
# 
#         // === STEP 2: Duplicate and close original ===
#         newTitle = baseName + "_proc";
#         run("Duplicate...", "title=" + newTitle);
#         dupTitle = getTitle(); 
#         print("Created duplicate: " + dupTitle);
# 
#         selectImage(originalTitle);
#         close();
#         print("Closed original composite: " + originalTitle);
# 
#         // === STEP 3: Split channels of duplicate ===
#         selectImage(dupTitle);
#         run("Split Channels");
#         print("Split channels from: " + dupTitle);
# 
#         // === STEP 4: Save the red channel ===
#         redTitle = newTitle + " (red)";
#         if (isOpen(redTitle)) {
#             selectWindow(redTitle);
#             saveAs("Tiff", saveRedFolder + baseName + "_TxRed.tif");
#             print("Saved and closed red channel: " + redTitle);
#             close();
#         } else {
#             print("WARNING: Red channel not found for " + baseName);
#         }
# 
#         // === STEP 5: Close green channel ===
#         greenTitle = newTitle + " (green)";
#         if (isOpen(greenTitle)) {
#             selectWindow(greenTitle);
#             close();
#             print("Closed green channel: " + greenTitle);
#         } else {
#             print("Green channel not found for " + baseName);
#         }
# 
#         // === STEP 6: Process DAPI (blue) channel ===
#         dapiTitle = newTitle + " (blue)";
#         if (isOpen(dapiTitle)) {
#             selectWindow(dapiTitle);
#             rename(baseName + "_DAPI_proc");
#             print("Processing DAPI channel: " + baseName);
# 
#             run("8-bit");
#             run("Enhance Contrast...", "saturated=0.3 normalize");  // normalize intensity
#             setAutoThreshold("Otsu");
#             getThreshold(lower, upper);
#             minThreshold = 10;
#             if (lower < minThreshold) lower = minThreshold;
#             setThreshold(lower, 255);
#             //convert and segment nuclei
#             setOption("BlackBackground", false);
#             run("Convert to Mask");
#             run("Watershed");
# 
#             run("Analyze Particles...", 
#                 "size=21-Infinity circularity=0.30-1.00 show=[Count Masks] display exclude summarize add");
#                 
#             // === If no ROIs detected, allow manual threshold adjustment ===
#     if (roiManager("count") == 0) {
#         print("⚠️ No ROIs found for: " + baseName + " — manual adjustment required.");
#         selectWindow(baseName + "_DAPI_proc");
#         run("Threshold...");
#         waitForUser("Adjust the threshold manually until nuclei are visible, then click OK.");
#         run("Convert to Mask");
#         run("Watershed");
# 
#         // Re-run particle analysis
#         run("Analyze Particles...", 
#             "size=21-Infinity circularity=0.30-1.00 show=[Count Masks] display exclude summarize add");
#     }
# 
#             // Save ROIs
#             if (roiManager("count") > 0) {
#                 saveName = baseName + "_DAPI.zip";
#                 roiManager("Save", saveRoiFolder + saveName);
#                 print("Saved ROI zip: " + saveName);
#             } else {
#                 print("No ROIs found for: " + baseName);
#             }
# 
#             // Cleanup DAPI
#             close();
#             maskTitle = "Count Masks of " + baseName + "_DAPI_proc";
#             if (isOpen(maskTitle)) {
#                 selectWindow(maskTitle);
#                 close();
#                 print("Closed mask window: " + maskTitle);
#             }
#             roiManager("Deselect");
#             roiManager("Reset");
#         } else {
#             print("DAPI channel not found for " + baseName);
#         }
# 
#         // === STEP 7: Close duplicate container window if still open ===
#         if (isOpen(dupTitle)) {
#             selectWindow(dupTitle);
#             close();
#             print("Closed duplicate window: " + dupTitle);
#         }
# 
#         print("Finished processing file: " + baseName);
#     }
# }

#each time I ran this I just changed the cell line name in the top

#Red Positivity Macro for identifying red pixels within ROIs and accounting for area
# // ==== USER SETTINGS ====
# roiFolder  = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 84-1\\outline new\\";   
# redImageFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 84-1\\TxRed new\\"
# outputFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 84-1\\outline new\\analyzed\\";
# imageExt  = ".tif";                                       
# thresholdLow = 20;
# thresholdHigh = 250;
# // ========================
# 
# // Clear results table for combined per-ROI data
# run("Clear Results");
# 
# // Initialize arrays for summary
# roiAreas = newArray();
# maxInts = newArray();
# meanInts = newArray();
# medianInts = newArray();
# percentReds = newArray();
# nROIsWithRed = 0;
# 
# // Get list of ROI ZIP files
# roiFiles = getFileList(roiFolder);
# 
# for (f = 0; f < roiFiles.length; f++) {
#     roiFile = roiFiles[f];
#     if (!endsWith(roiFile, "_DAPI.zip")) continue;
# 
#     // Build corresponding red image filename
#     base = replace(roiFile, "_DAPI.zip", "");
#     redImageName = base + "_TxRed.tif";
#     redImagePath = redImageFolder + redImageName;
# 
#     print("Trying to open red image: " + redImagePath);
# 
#     if (!File.exists(redImagePath)) {
#         print("Red image not found for " + roiFile + ", skipping.");
#         continue;
#     }
# 
#     roiZipPath = roiFolder + roiFile;
# 
#     // Open red image
#     open(redImagePath);
#     redTitle = getTitle();
#     selectWindow(redTitle);
# 
#     // Load ROIs
#     roiManager("Reset");
#     roiManager("Open", roiZipPath);
# 
#     nROIs = roiManager("Count");
#     for (r = 0; r < nROIs; r++) {
#         roiManager("Select", r);
#         selectWindow(redTitle);
# 
#         // Declare measurement variables
#         areaInt = 0;
#         meanInt = 0;
#         minInt = 0;
#         maxInt = 0;
#         stdInt = 0;
#         histogram = newArray(256);
# 
#         getStatistics(areaInt, meanInt, minInt, maxInt, stdInt, histogram);
# 
#         // Count red-positive pixels
#         redCount = 0;
#         for (p = thresholdLow; p <= thresholdHigh; p++) redCount += histogram[p];
#         if (redCount == 0) continue; // skip ROIs with no red
# 
#         percentRed = (redCount / areaInt) * 100;
#         medianIntVal = getMedianFromHistogram(histogram, areaInt);
# 
#         // Save per-ROI results
#         row = nResults;
#         setResult("ROI_File", row, roiFile);
#         setResult("ROI_Area", row, areaInt);
#         setResult("Max_Intensity", row, maxInt);
#         setResult("Mean_Intensity", row, meanInt);
#         setResult("Median_Intensity", row, medianIntVal);
#         setResult("Red_Positive_Pixels", row, redCount);
#         setResult("Percent_Red_Positive", row, percentRed);
#         updateResults();
# 
#         // Store for summary
#         roiAreas = arrayAdd(roiAreas, areaInt);
#         maxInts = arrayAdd(maxInts, maxInt);
#         meanInts = arrayAdd(meanInts, meanInt);
#         medianInts = arrayAdd(medianInts, medianIntVal);
#         percentReds = arrayAdd(percentReds, percentRed);
#         nROIsWithRed++;
#     }
# 
#     close(redTitle);
# }
# 
# // Save combined per-ROI CSV
# saveAs("Results", outputFolder + "All_ROIs_Combined.csv");
# print("Combined per-ROI results saved.");
# 
# // Build summary table
# run("Clear Results");
# 
# totalNuclearArea = sumArray(roiAreas);
# 
# setResult("Experiment", 0, roiFolder);
# setResult("N_Nuclei_With_Red", 0, nROIsWithRed);
# setResult("Mean_ROI_Area", 0, mean(roiAreas));
# setResult("SD_ROI_Area", 0, sd(roiAreas));
# setResult("Mean_Max_Intensity", 0, mean(maxInts));
# setResult("SD_Max_Intensity", 0, sd(maxInts));
# setResult("Mean_Mean_Intensity", 0, mean(meanInts));
# setResult("SD_Mean_Intensity", 0, sd(meanInts));
# setResult("Mean_Median_Intensity", 0, mean(medianInts));
# setResult("SD_Median_Intensity", 0, sd(medianInts));
# setResult("Mean_Percent_Red_Positive", 0, mean(percentReds));
# setResult("SD_Percent_Red_Positive", 0, sd(percentReds));
# setResult("Total_Nuclear_Area", 0, totalNuclearArea);
# updateResults();
# 
# // Save summary CSV
# saveAs("Results", outputFolder + "Experiment_Summary.csv");
# print("Summary saved.");
# 
# // ==========================
# // Helper functions
# function getMedianFromHistogram(hist, totalPixels) {
#     sum = 0;
#     for (i = 0; i < lengthOf(hist); i++) {
#         sum += hist[i];
#         if (sum >= totalPixels / 2) return i;
#     }
#     return NaN;
# }
# 
# function arrayAdd(arr, value) {
#     if (lengthOf(arr) == 0) {
#         newArr = newArray(value);
#     } else {
#         newArr = newArray();
#         for (i = 0; i < lengthOf(arr); i++) newArr[i] = arr[i];
#         newArr[lengthOf(newArr)] = value;
#     }
#     return newArr;
# }
# 
# function mean(arr) {
#     if (lengthOf(arr) == 0) return NaN;
#     s = 0;
#     for (i = 0; i < lengthOf(arr); i++) s += arr[i];
#     return s / lengthOf(arr);
# }
# 
# function sd(arr) {
#     if (lengthOf(arr) <= 1) return NaN;
#     m = mean(arr);
#     s = 0;
#     for (i = 0; i < lengthOf(arr); i++) s += (arr[i] - m) * (arr[i] - m);
#     return sqrt(s / (lengthOf(arr) - 1));
# }
# 
# function sumArray(arr) {
#     s = 0;
#     for (i = 0; i < lengthOf(arr); i++) s += arr[i];
#     return s;
# }

#these two macros gave the desired output of a csv file with all of the parameters I needed

```

## Read in ROI Summaries

```{r Read in ROI summary from IF macros, message=FALSE, warning=FALSE}
#read in the data from the macros that I ran for each individual
#I am going to focus on 20X magnification images, remove any non-20X samples 
#ensure that you label each with respective individual number for ease later on

####Ind 1 - 87-1####
# ind1_roi <- read.csv("D:/DXR Project/IF yH2AX/if yh2ax dxr 87-1/outline new/analyzed/All_ROIs_Combined_87-1_EMP_250825.csv") %>%
#   mutate(Individual = "Ind1",
#          Line = "87-1",
#          Magnification = case_when(
#            grepl("20X", ROI_File, ignore.case = TRUE) ~ "20X",
#            grepl("10X", ROI_File, ignore.case = TRUE) ~ "10X",
#            grepl("40X", ROI_File, ignore.case = TRUE) ~ "40X",
#            TRUE ~ "Other"
#          )) %>%
#   dplyr::select(-"X")

# saveRDS(ind1_roi, "data/Fig1/IF_quant/All_ROIs_Combined_87-1_Ind1.RDS")

####Ind 3 - 75-1####
# ind3_roi <- read.csv("D:/DXR Project/IF yH2AX/if yh2ax dxr 75-1/outline new/analyzed/All_ROIs_Combined_75-1_EMP_250825.csv") %>%
#     mutate(Individual = "Ind3",
#            Line = "75-1",
#          Magnification = case_when(
#            grepl("20X", ROI_File, ignore.case = TRUE) ~ "20X",
#            grepl("10X", ROI_File, ignore.case = TRUE) ~ "10X",
#            grepl("40X", ROI_File, ignore.case = TRUE) ~ "40X",
#            TRUE ~ "Other"
#          )) %>%
#   dplyr::select(-"X")

# saveRDS(ind3_roi, "data/Fig1/IF_quant/All_ROIs_Combined_75-1_Ind3.RDS")

####Ind 4 - 84-1####
# ind4_roi <- read.csv("D:/DXR Project/IF yH2AX/if yh2ax dxr 84-1/outline new/analyzed/All_ROIs_Combined_84-1_EMP_250916.csv") %>%
#   mutate(Individual = "Ind4") %>%
#   mutate(Line = "84-1") %>%
#   mutate(Magnification = "20X") #specify since all were 20X so not in labels

# saveRDS(ind4_roi, "data/Fig1/IF_quant/All_ROIs_Combined_84-1_Ind4.RDS")

#keep all of this for posterity and read them in for plots
ind1_roi <- readRDS("data/Fig1/IF_quant/All_ROIs_Combined_87-1_Ind1.RDS")

ind3_roi <- readRDS("data/Fig1/IF_quant/All_ROIs_Combined_75-1_Ind3.RDS")

ind4_roi <- readRDS("data/Fig1/IF_quant/All_ROIs_Combined_84-1_Ind4.RDS")


```

## Create Dataframes

```{r Create Summary Sheet all Ind, message=FALSE, warning=FALSE}
#I want to combine all of these csv files into one dataframe

#ensure that I am removing secondary only controls and any other magnifications (ie 10X and 40X images)
 
# ind_roi_all <- bind_rows(ind1_roi, ind3_roi, ind4_roi) %>%
#   mutate(
#     Treatment = case_when(
#       grepl("DOX", ROI_File) ~ "DOX",
#       grepl("DMSO", ROI_File) ~ "VEH",
#       TRUE ~ "Other"
#     ),
#     Timepoint = case_when(
#       grepl("24T", ROI_File) ~ "24T",
#       grepl("24R", ROI_File) ~ "24R",
#       grepl("144R", ROI_File) ~ "144R",
#       TRUE ~ "Other"
#     ),
#     Condition = paste0(Treatment, "_", Timepoint)
#   ) %>%
#   dplyr::filter(Treatment != "Other") %>%
#   dplyr::filter(Magnification == "20X") %>%
#   dplyr::filter(!grepl("^2nd", ROI_File, ignore.case = TRUE)) %>%
#   dplyr::select(-"X")

# saveRDS(ind_roi_all, "data/Fig1/IF_quant/individual_rois_all.RDS")

#now I have successfully made sure this only contains 20X and has all of the three individuals

```

## IF Quant Step 1 

Pull all ROIs from each image to make a combined dataframe.

```{r Step 1 - ROI Level All Individuals, message=FALSE, warning=FALSE}

#make sure to normalize by area to get positive red for the next step
# ind_roi_all <- ind_roi_all %>%
#   mutate(
#     Percent_Red_Positive_Area = (Red_Positive_Pixels / ROI_Area) * 100
#   )

#save as an R Object
# saveRDS(ind_roi_all, "data/Fig1/IF_quant/individual_rois_all.RDS")

ind_roi_all <- readRDS("data/Fig1/IF_quant/individual_rois_all.RDS") 

ind_cols_if <- c("Ind1" = "#FFA364",
              "Ind3" = "#A4D177",
              "Ind4" = "#EAD050")

#previous timepoint names prior to tx+0 (24T), tx+24 (24R), tx+144 (144R)
condition_levels <- c("DOX_24T", "VEH_24T", 
                      "DOX_24R", "VEH_24R", 
                      "DOX_144R", "VEH_144R")

#add in colors for boxplots
txtime_col_old <- list(
  "DOX_24T" = "#263CA5",
  "VEH_24T" = "#444343",
  "DOX_24R" = "#5B8FD1",
  "VEH_24R" = "#757575",
  "DOX_144R" = "#89C5E5",
  "VEH_144R" = "#AAAAAA"
)

plot_step1 <- ind_roi_all %>% 
  mutate(Condition = factor(Condition, levels = c(condition_levels))) %>% 
  ggplot(aes(x = Condition, y = Percent_Red_Positive_Area)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 1.5,
             alpha = 0.6, #for visibility of all of these points
             position = position_identity()) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 1: ROI Level % Red Positive",
       x = "Condition",
       y = "Percent Red Positive Pixels") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_step1

```

## IF Quant Step 2

Collapse ROIs into an image level summary.

```{r Image Level Summary Dataframe, message=FALSE, warning=FALSE, include=FALSE}

#calculate the area for each individual level roi before passing through function
# ind1_roi <- ind1_roi %>% mutate(Percent_Red_Positive_Area = (Red_Positive_Pixels / ROI_Area) * 100)
# ind3_roi <- ind3_roi %>% mutate(Percent_Red_Positive_Area = (Red_Positive_Pixels / ROI_Area) * 100)
# ind4_roi <- ind4_roi %>% mutate(Percent_Red_Positive_Area = (Red_Positive_Pixels / ROI_Area) * 100)

#now I will collapse these to image level using the below function
# collapse_to_image <- function(roi_df) {
#   roi_df %>%
#     mutate(Image = gsub("_ROI.*", "", ROI_File),
#            Image = gsub("\\.zip$", "", Image),
#            Image = trimws(Image)
#            ) %>%
#     group_by(Image, Individual) %>%
#     summarise(
#       Mean_of_Mean   = mean(Mean_Intensity, na.rm = TRUE),
#       Mean_of_Median = mean(Median_Intensity, na.rm = TRUE),
#       Mean_of_Max    = mean(Max_Intensity, na.rm = TRUE),
#       Mean_Percent_Red_Positive = mean(Percent_Red_Positive, na.rm = TRUE),
#       Mean_Red_Positive_Pixels  = mean(Red_Positive_Pixels, na.rm = TRUE),
#       Mean_Norm_Percent_Red_Positive =
#         mean(Percent_Red_Positive_Area, na.rm = TRUE),
#       .groups = "drop"
#     )
# }

#use function to create a summary for each individual
# ind1_img_summary <- collapse_to_image(ind1_roi) %>%
#   mutate(Individual = "Ind1",
#          Magnification = case_when(
#            grepl("20X", Image, ignore.case = TRUE) ~ "20X",
#            grepl("10X", Image, ignore.case = TRUE) ~ "10X",
#            grepl("40X", Image, ignore.case = TRUE) ~ "40X",
#            TRUE ~ "Other"
#            ))
# 
# ind3_img_summary <- collapse_to_image(ind3_roi) %>%
#   mutate(Individual = "Ind3",
#          Magnification = case_when(
#            grepl("20X", Image, ignore.case = TRUE) ~ "20X",
#             grepl("10X", Image, ignore.case = TRUE) ~ "10X",
#            grepl("40X", Image, ignore.case = TRUE) ~ "40X",
#             TRUE ~ "Other"
#            ))
# 
# ind4_img_summary <- collapse_to_image(ind4_roi) %>%
#   mutate(Individual = "Ind4",
#          Magnification = "20X") #only has 20X so add this label

#now, combine all individuals together with their image level summaries
# roi_image_level_summary <- bind_rows(ind1_img_summary,
#                                      ind3_img_summary,
#                                      ind4_img_summary)

#ensure that the secondary only ctrls are removed from this one too
# roi_image_level_summary <- roi_image_level_summary %>%
#   dplyr::filter(!grepl("^2nd", Image, ignore.case = TRUE)) %>%
#    dplyr::filter(Magnification == "20X")

# saveRDS(roi_image_level_summary, "data/Fig1/IF_quant/roi_image_level_summary_all.RDS")

```

## Read in Image Level Summary

```{r ROI Summary Dataframe, message=FALSE, warning=FALSE}
#now that I've created the image level summary table for all individuals above, read it in so I don't have to create it multiple times

roi_image_level_summary <- readRDS("data/Fig1/IF_quant/roi_image_level_summary_all.RDS")

```

## Make a Metadata Table

```{r Metadata Table Image Level, warning=FALSE, include=FALSE, message=FALSE}
#now I want to make a metadata table with all of the relevant info 
#these files come from my DAPI Counts new folder from each individual

# roi_folders <- list(
#   Ind1 = "D:/DXR Project/IF yH2AX/if yh2ax dxr 87-1/DAPI counts new",
#   Ind3 = "D:/DXR Project/IF yH2AX/if yh2ax dxr 75-1/DAPI counts new",
#   Ind4 = "D:/DXR Project/IF yH2AX/if yh2ax dxr 84-1/DAPI Counts new" #be careful since this one has a slightly different name
# )

#read the csvs that come from the folders and loop over each individual
# read_csvs <- function(folder, individual_label) {
#   files <- list.files(folder, pattern = "^Cell_intensity",
#                       full.names = TRUE,
#                       ignore.case = TRUE)
# 
#   if(length(files) == 0) {
#     message("No csv files found in folder: ", folder)
#     return(NULL)
#   }
# 
#   lapply(files, function(file) {
#     data <- read.csv(file)
#     tibble(
#       Individual = individual_label,
#       Image = tools::file_path_sans_ext(basename(file)),
#       Nuclei_Count = nrow(data)
#     )
#   }) %>% bind_rows()
# }

#now loop over each individual using the above function
# ind1_metadata <- read_csvs(roi_folders$Ind1, "Ind1") %>%
#   mutate(Magnification = case_when(
#     grepl("20X", Image, ignore.case = TRUE) ~ "20X",
#     grepl("10X", Image, ignore.case = TRUE) ~ "10X",
#     grepl("40X", Image, ignore.case = TRUE) ~ "40X",
#     TRUE ~ "Other"
#     )) %>%
#   dplyr::filter(Magnification == "20X")
# 
# ind3_metadata <- read_csvs(roi_folders$Ind3, "Ind3") %>%
#   mutate(Magnification = case_when(
#           grepl("20X", Image, ignore.case = TRUE) ~ "20X",
#            grepl("10X", Image, ignore.case = TRUE) ~ "10X",
#            grepl("40X", Image, ignore.case = TRUE) ~ "40X",
#           TRUE ~ "Other"
#           )) %>%
#   dplyr::filter(Magnification == "20X")
# 
# ind4_metadata <- read_csvs(roi_folders$Ind4, "Ind4") %>%
#   mutate(Magnification = "20X")
# 
# image_level_metadata <- bind_rows(ind1_metadata,
#                                   ind3_metadata,
#                                   ind4_metadata)


#now parse out the treatments and remove the prefix cell_intensity so it matches my other summary file
#add Line in so I can make sure that the individuals are correct after renum
# image_level_metadata <- image_level_metadata %>%
#   mutate(
#     Image = str_remove(Image, "^Cell_intensity_"),
#     Treatment = ifelse(str_detect(Image, "DOX"), "DOX",
#                        ifelse(str_detect(Image, "DMSO"), "VEH", NA_character_)),
#     Timepoint = str_extract(Image, "24T|24R|144R"),
#     Line = case_when(
#       Individual == "Ind1" ~ "87-1",
#       Individual == "Ind3" ~ "75-1",
#       Individual == "Ind4" ~ "84-1",
#       TRUE ~ NA_character_
#     )
#   ) %>%
#   dplyr::filter(!is.na(Treatment)) %>%
#   dplyr::filter(!str_detect(Image, "test")) %>%
#   dplyr::filter(!str_detect(Image, "edit")) %>%
#   dplyr::select(Individual, Line, Treatment, Timepoint, Image, Nuclei_Count) %>%
#   mutate(
#     Condition = paste0(Treatment, "_", Timepoint),
#     Magnification = "20X"
#   )

#save this metadata dataframe as an R object and as a csv for later
# saveRDS(image_level_metadata, "data/Fig1/IF_quant/image_level_metadata_all.RDS")

# write.csv(image_level_metadata, "data/Fig1/IF_quant/image_level_metadata_all.csv")

```

## Combine Metadata with Image Level Summary

```{r Summary and Metadata, include = FALSE, message=FALSE, warning=FALSE}
#Now that I've made my image_level_metadata sheet match to my roi_image_level_summary, let's combine them

#I commented out the top for posterity so I don't have to run it multiple times
#thus I will read in my two sheets to combine them

#original summary sheet roi_image_level_summary needed

#metadata sheet
image_level_metadata <- readRDS("data/Fig1/IF_quant/image_level_metadata_all.RDS")

#create the merged image summary for all individuals
# merged_image_summary <-
#   image_level_metadata %>%
#   dplyr::select(Individual, Line, Treatment, Timepoint, Condition,
#                 Magnification, Image, Nuclei_Count) %>%
#   dplyr::left_join(roi_image_level_summary, by = c("Image", "Individual", "Magnification"))

# merged_image_summary <- merged_image_summary %>%
#   mutate(Condition = factor(Condition,
#                             levels = c(condition_levels)))

# saveRDS(merged_image_summary, "data/Fig1/IF_quant/merged_image_summary_all.RDS")

# write.csv(merged_image_summary, "data/Fig1/IF_quant/merged_image_summary_all.csv")

#read back in my merged image summary so I don't have to run it multiple times

merged_image_summary <- readRDS("data/Fig1/IF_quant/merged_image_summary_all.RDS")

```

## IF Quant Step 2

Collapse ROIs into their respective images.

```{r Step 2 - Collapse by Image, message=FALSE, warning=FALSE}
#now I'll plot the outcome of the second step where I collapsed by image and merged the metadata with the measurements

plot_step2 <- ggplot(merged_image_summary, 
         aes(x = Condition,
             y = Mean_Red_Positive_Pixels)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             alpha = 0.8, #for visibility as I collapse data down
             size = 3,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 2: Image-Level Mean Red Positive Pixels",
       x = "Condition", 
       y = "Mean Red Positive Pixels") +
  theme_custom() +
  theme(legend.position = "bottom")


plot_step2

```

## IF Quant Step 3 

Select 3 images with consistent numbers of nuclei per individual.

```{r Step 3 - Images by Consistent Nuclei Count, message=FALSE, warning=FALSE}
#now I will be selecting images based on their nuclei count

#first let's find the median number of nuclei per individual to get an idea of how many are there per experiment
#all experiments were performed separately but with the same protocol, so each individual is separate

# median_nuclei <- merged_image_summary %>%
#   group_by(Individual, Condition) %>%
#   summarize(
#     Median_Nuclei = median(Nuclei_Count, na.rm = TRUE),
#     Mean_Positive = mean(Mean_Red_Positive_Pixels),
#     Image_Count = n(),
#     .groups = "drop"
#   ) %>%
#   arrange(Individual, Condition)

#let's pick out the three images per condition that have consistent nuclei counts

#first I am going to remove some outlier images from my Ind1 since I had some dim/blurry ones that should not be included since they weren't properly detected
#I chose the images that were in focus and had consistent numbers of nuclei based on summary from above 

# Define the good images for Ind4 (IDs = the 4-digit part before "_DAPI")
# images_Ind4 <- tibble::tribble(
#   ~Condition,   ~ID,
#   "DOX_24T",  "0012",
#   "DOX_24T",  "0013",
#   "DOX_24T",  "0017",
#   "VEH_24T",  "0006",
#   "VEH_24T",  "0004",
#   "VEH_24T",  "0003",
#   "DOX_24R",  "0006",
#   "DOX_24R",  "0004",
#   "DOX_24R",  "0003",
#   "VEH_24R",  "0004",
#   "VEH_24R",  "0005",
#   "VEH_24R",  "0003",
#   "DOX_144R", "0002",
#   "DOX_144R", "0003",
#   "DOX_144R", "0005",
#   "VEH_144R", "0003",
#   "VEH_144R", "0004",
#   "VEH_144R", "0012"
# )

#filter my merged_image_summary for Ind1 using IDs
# ind4_filtered <- merged_image_summary %>%
#   filter(Individual == "Ind4") %>%
#   mutate(
#     Image_ID  = str_extract(Image, "_\\d{4}_") %>% str_remove_all("_")
#   ) %>%
#   inner_join(images_Ind1, by = c("Condition", "Image_ID" = "ID"))

#now join this back into my pipeline because now I have the images I need based on nucleus count?

# merged_image_summary_filt <- merged_image_summary %>%
#   dplyr::filter(Individual != "Ind4") %>%
#   bind_rows(ind1_filtered)

#now that I've put Ind1 back in, let's select images based on nuclei for Ind1 and Ind3

# select_top3_images <- function(df, individual) {
#   df %>%
#     filter(Individual == individual) %>%
#     group_by(Condition) %>%
#     mutate(
#       med_nuclei = median(Nuclei_Count, na.rm = TRUE),
#       med_pos    = median(Mean_Norm_Percent_Red_Positive, na.rm = TRUE),
#       score = abs(Nuclei_Count - med_nuclei) + abs(Mean_Norm_Percent_Red_Positive - med_pos)
#     ) %>%
#     arrange(score) %>%
#     slice_head(n = 3) %>%
#     ungroup()
# }

# ind1_filtered <- select_top3_images(merged_image_summary, "Ind1")
# ind3_filtered <- select_top3_images(merged_image_summary, "Ind3")

#filter Ind 4 to get the same scores at the end
# ind4_filtered <- ind4_filtered %>%
#     group_by(Condition) %>%
#     mutate(
#         med_nuclei = median(Nuclei_Count, na.rm = TRUE),
#         med_pos    = median(Mean_Percent_Red_Positive, na.rm = TRUE),
#         score = abs(Nuclei_Count - med_nuclei) +
#             abs(Mean_Percent_Red_Positive - med_pos)
#     ) %>%
#     ungroup()

#merge these summaries together to have all individuals represented in one
# merged_image_summary_final <- bind_rows(
#   merged_image_summary %>% filter(!Individual %in% c("Ind1", "Ind3", "Ind4")),
#   ind1_filtered,
#   ind3_filtered,
#   ind4_filtered
# ) %>%
#   dplyr::select(-("Image_ID"))

#save the finalized version as an R object and read it back in, keep code for posterity
# saveRDS(merged_image_summary_final, "data/Fig1/IF_quant/merged_image_summary_final.RDS")

merged_image_summary_final <- readRDS("data/Fig1/IF_quant/merged_image_summary_final.RDS")

#plot these in a boxplot now that they're filtered down to the top 3 images per individual based on consistent nuclei counts

plot_step3 <- ggplot(merged_image_summary_final, 
                     aes(x = Condition, 
                         y = Mean_Norm_Percent_Red_Positive)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 2, 
             alpha = 1, #now fewer points so no need to be transparent
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 3: Select 3 Images Per Individual with Consistent Nuclei",
       x = "Condition",
       y = "Mean Percent Red Positive") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_step3

#do stats for step 3 boxplot between DOX and VEH at each timepoint for comparison
ttests_step3 <- merged_image_summary_final %>%
  group_by(Timepoint) %>%
  t_test(Mean_Norm_Percent_Red_Positive ~ Treatment) %>%
  mutate(
    y.position = max(merged_image_summary_final$Mean_Norm_Percent_Red_Positive, na.rm = TRUE) * 1.05
  )

sig_table_step3 <- merged_image_summary_final %>%
  group_by(Timepoint) %>%
  summarise(
    p = t.test(Mean_Norm_Percent_Red_Positive ~ Treatment)$p.value,
    y.position = max(Mean_Norm_Percent_Red_Positive, na.rm = TRUE) * 1.05,
    .groups = "drop"
  ) %>%
  mutate(
    signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint)
  )
  
ttests_step3

#plot with t test statistics included above plot
plot_step3_test <- ggplot(merged_image_summary_final, 
                     aes(x = Condition,
                         y = Mean_Norm_Percent_Red_Positive)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 2,
             alpha = 1,
             position = position_identity()) +
  stat_pvalue_manual(sig_table_step3, 
                     label = "signif", 
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 3: Select 3 Images Per Individual with Consistent Nuclei",
       x = "Condition", 
       y = "Mean Percent Red Positive") +
  theme_custom() +
  theme(legend.position = "bottom")


plot_step3_test


```

## IF Quant Step 4

Normalize DOX/VEH before combining images by individual.


```{r Step 4 - Normalization by Max Intensity, message=FALSE, warning=FALSE}

#Now I want to normalize these images by max intensity

merged_image_summary_norm <- merged_image_summary_final %>%
  group_by(Magnification) %>%
  mutate(Norm_Percent_Red_Positive = (Mean_Percent_Red_Positive / max(Mean_Percent_Red_Positive, na.rm = TRUE)) * 100) %>%
  ungroup() %>%
  mutate(Condition = paste0(Treatment, "_", Timepoint))

#now these are scaled 0-100

individual_summary_norm <- merged_image_summary_norm %>%
  group_by(Individual, Treatment, Timepoint) %>%
  summarise(
    Mean_Norm = mean(Norm_Percent_Red_Positive, na.rm = TRUE),
    Median_Norm = median(Norm_Percent_Red_Positive, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )

#alter the t test table from before to plot the same comparisons
ttest_table_step4 <- ttests_step3 %>%
  filter(Timepoint %in% c("24T", "24R", "144R")) %>%
  mutate(
    group1 = paste0("DOX_", Timepoint),
    group2 = paste0("VEH_", Timepoint),
    y.position = max(individual_summary_norm$Mean_Norm, na.rm = TRUE) * 1.05,
    
    #add in significance stars
    p.signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ "ns"
    )
  )

#plot step 4
plot_step4 <- ggplot(merged_image_summary_norm, 
                     aes(x = Condition,
                         y = Norm_Percent_Red_Positive)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 2, 
             alpha = 1, 
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 105)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_step4,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  labs(title = "Step 4: Normalized by Max Intensity (0–100%)",
       x = "Condition", 
       y = "Percent Red Positive") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_step4

```

## IF Quant Step 5

Collapse images by individual.


```{r Step 5 - Collapse by Individual}
#now I want to collapse these three images per condition into their respective individual after normalization
#I now only have three points, one per individual, on each boxplot
#I also performed a t-test between DOX and VEH at each timepoint to assess differences between treated and untreated within each timepoint

individual_summary_step5 <- merged_image_summary_norm %>%
  group_by(Individual, Treatment, Timepoint, Magnification) %>%
  summarise(
    Mean_Norm = mean(Norm_Percent_Red_Positive, na.rm = TRUE),
    Median_Norm = median(Norm_Percent_Red_Positive, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )

#utilize the same t tests performed earlier since it is just a collapsed version of the same data
#changed the name to reduce confusion
ttest_table_step5 <- ttests_step3 %>%
  filter(Timepoint %in% c("24T", "24R", "144R")) %>%
  mutate(
    group1 = paste0("DOX_", Timepoint),
    group2 = paste0("VEH_", Timepoint),
    y.position = max(individual_summary_step5$Mean_Norm, na.rm = TRUE) * 1.05,
    
    #add in significance stars
    p.signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01  ~ "**",
      p < 0.05  ~ "*",
      TRUE      ~ "ns"
    )
  )


#plot the collapsed image-level data 0-100
plot_step5 <- ggplot(individual_summary_step5, 
                     aes(x = Condition, y = Mean_Norm)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1, 
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_step5,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  labs(title = "Step 5: Individual-Level Percent Red Positive",
       x = "Condition", 
       y = "Normalized Percent Positive") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_step5

# save_plot(
#   plot = plot_step5,
#   filename = "IFQuant_Step5_Final_EMP",
#   folder = output_folder,
#   height = plot_height,
#   width = plot_width
# )

```


## IF Quant Step 6

Perform significance testing between treated timepoints.

```{r Step 6 - Significance Between Timepoints DOX}
#Above, I compared DOX to VEH for each timepoint
#I want to now make comparisons between DOX timepoints to see if there are any differences between each timepoint after treatment and recovery time
#I will then add these to the plot in Illustrator to show both comparisons together since they don't want to plot correctly

individual_summary_step6_dox <- merged_image_summary_norm %>%
  group_by(Individual, Treatment, Timepoint, Magnification) %>%
  summarise(
    Mean_Norm = mean(Norm_Percent_Red_Positive, na.rm = TRUE),
    Median_Norm = median(Norm_Percent_Red_Positive, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )

#set condition levels for x axis to plot all boxes
condition_levels <- c("DOX_24T", "VEH_24T",
                      "DOX_24R", "VEH_24R",
                      "DOX_144R", "VEH_144R")

dox_conditions <- c("DOX_24T", "DOX_24R", "DOX_144R")

individual_summary_step6_dox <- individual_summary_step6_dox %>%
  mutate(Condition = factor(Condition, levels = condition_levels))

#perform paired t-tests between DOX timepoints
timepoint_pairs <- list(
  c("24T","24R"),
  c("24R","144R"),
  c("24T","144R")
)

#t test table
ttest_table_dox_time <- purrr::map_df(seq_along(timepoint_pairs), function(i){
  tp <- timepoint_pairs[[i]]
  
  df_wide <- individual_summary_step6_dox %>%
    filter(Timepoint %in% tp) %>%
    group_by(Individual, Timepoint) %>%
    summarise(Mean_Norm = mean(Mean_Norm, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Timepoint, values_from = Mean_Norm) %>%
    drop_na()
  
  base_y <- max(individual_summary_step6_dox$Mean_Norm, 
                na.rm = TRUE) * 0.9
  
  y_increment <- max(individual_summary_step6_dox$Mean_Norm, 
                     na.rm = TRUE) * 0.05
  
  y_pos <- base_y + y_increment * i 
  
  if(nrow(df_wide) < 2){
    tibble(
      group1 = paste0("DOX_", tp[1]),
      group2 = paste0("DOX_", tp[2]),
      p.value = NA_real_,
      p.signif = "NA",
      y.position = y_pos
    )
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1 = paste0("DOX_", tp[1]),
      group2 = paste0("DOX_", tp[2]),
      p.value = test$p.value,
      p.signif = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01  ~ "**",
        test$p.value < 0.05  ~ "*",
        TRUE                 ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to numeric positions for stat_pvalue_manual
ttest_table_dox_time <- ttest_table_dox_time %>%
  mutate(
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

#manually set y positions for each comparison (so I can see all of it)
max_val <- max(individual_summary_step6_dox$Mean_Norm, na.rm = TRUE)

ttest_table_dox_time <- ttest_table_dox_time %>%
  mutate(
    y.position = case_when(
      group1 == "DOX_24T" & group2 == "DOX_24R" ~ max_val * 1.05,  #middle
      group1 == "DOX_24T" & group2 == "DOX_144R" ~ max_val * 1.10, #top
      group1 == "DOX_24R" & group2 == "DOX_144R" ~ max_val * 1.00  #bottom
    ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

#plot DOX timepoint comparisons
plot_step6_dox_time <- ggplot(individual_summary_step6_dox, 
                              aes(x = Condition, 
                                  y = Mean_Norm)) +
  geom_boxplot(aes(fill = Condition), 
               show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1,
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 105)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_dox_time, 
                     label = "p.signif", 
                     xmin = "group1_num", 
                     xmax = "group2_num", 
                     y.position = "y.position",
                     tip.length = 0.02) +
  labs(title = "Step 6: DOX Timepoint Comparisons Norm Percent Positive",
       x = "Condition",
       y = "Normalized Percent Positive") +
  theme_custom() +
  theme(legend.position = "bottom")

#print the plot
plot_step6_dox_time

# save_plot(
#   plot = plot_step6_dox_time,
#   filename = "IFQuant_Step6_DOX_TimeComp_EMP",
#   folder = output_folder
# )

```
# IF Hoechst (DAPI) Quant - Macros for FIJI/ImageJ

```{r imagej quant macros, warning=FALSE, message=FALSE}

# // DAPI Intensity Only — using existing ROI zips from above
# folder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 75-1\\rgb quant\\";
# roiFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 75-1\\outline new\\";
# saveDAPIFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 75-1\\DAPI intensity\\";
# 
# list = getFileList(folder);
# 
# for (i = 0; i < list.length; i++) {
#     if (endsWith(list[i], ".tif")) {
#         print("----------------------------------------------------");
#         print("Processing file: " + list[i]);
# 
#         baseName = replace(list[i], ".tif", "");
# 
#         // === Check ROI zip exists ===
#         roiPath = roiFolder + baseName + "_DAPI.zip";
#         if (!File.exists(roiPath)) {
#             print("WARNING: No ROI zip found for " + baseName + " — skipping.");
#             continue;
#         }
# 
#         // === Open composite and extract DAPI (blue) channel ===
#         open(folder + list[i]);
#         originalTitle = getTitle();
# 
#         run("Duplicate...", "title=" + baseName + "_proc");
#         selectImage(originalTitle);
#         close();
# 
#         selectWindow(baseName + "_proc");
#         run("Split Channels");
# 
#         // Close red and green, keep blue
#         redTitle = baseName + "_proc (red)";
#         greenTitle = baseName + "_proc (green)";
#         blueTitle = baseName + "_proc (blue)";
# 
#         if (isOpen(redTitle)) { selectWindow(redTitle); close(); }
#         if (isOpen(greenTitle)) { selectWindow(greenTitle); close(); }
# 
#         if (!isOpen(blueTitle)) {
#             print("WARNING: Blue channel not found for " + baseName + " — skipping.");
#             continue;
#         }
# 
#         // === Load ROIs and measure DAPI intensity ===
#         selectWindow(blueTitle);
#         roiManager("Reset");
#         roiManager("Open", roiPath);
#         print("Loaded " + roiManager("count") + " ROIs for " + baseName);
# 
#         run("Set Measurements...", "mean min max integrated area redirect=None decimal=3");
#         roiManager("Deselect");
#         roiManager("Measure");
# 
#         saveAs("Results", saveDAPIFolder + baseName + "_DAPI_intensity.csv");
#         print("Saved: " + baseName + "_DAPI_intensity.csv");
#         run("Clear Results");
# 
#         // === Cleanup ===
#         selectWindow(blueTitle);
#         close();
#         roiManager("Deselect");
#         roiManager("Reset");
# 
#         print("Finished: " + baseName);
#     }
# }

```

# IF DAPI Quant - Step 1

## Read in DAPI csvs from ImageJ

```{r dapi quantification step 1, warning=FALSE, message=FALSE}
#for the first step, I want to collect all of the information from the macros I ran to get mean intensity from the same images used above for the yH2AX quantification

#create a function to read all per-image DAPI CSVs for each cell line and input magnification
# read_dapi_intensity <- function(cell_line, individual_label, default_magnification = NULL) {
#   dapi_folder <- paste0("D:/DXR Project/IF yH2AX/if yh2ax dxr ", cell_line, "/DAPI intensity/")
#   files <- list.files(dapi_folder, pattern = "_DAPI_intensity\\.csv$", full.names = TRUE)
#   
#   bind_rows(lapply(files, function(f) {
#     read.csv(f) %>%
#       mutate(
#         ROI_File = gsub("_DAPI_intensity\\.csv$", "", basename(f)),
#         Individual = individual_label,
#         Line = cell_line,
#         Magnification = if (!is.null(default_magnification)) {
#           default_magnification
#         } else {
#           case_when(
#             grepl("20X", ROI_File, ignore.case = TRUE) ~ "20X",
#             grepl("10X", ROI_File, ignore.case = TRUE) ~ "10X",
#             grepl("40X", ROI_File, ignore.case = TRUE) ~ "40X",
#             TRUE ~ "Other"
#           )
#         }
#       )
#   }))
# }

# ind1_dapi <- read_dapi_intensity("87-1", "Ind1") %>% dplyr::select(-X)
# ind3_dapi <- read_dapi_intensity("75-1", "Ind3") %>% dplyr::select(-X)
# ind4_dapi <- read_dapi_intensity("84-1", "Ind4", default_magnification = "20X") %>% dplyr::select(-X)

# saveRDS(ind1_dapi, "data/Fig1/ind1_dapi_rois.RDS")
# saveRDS(ind3_dapi, "data/Fig1/ind3_dapi_rois.RDS")
# saveRDS(ind4_dapi, "data/Fig1/ind4_dapi_rois.RDS")

#read in the finalized files
ind1_dapi <- readRDS("data/Fig1/ind1_dapi_rois.RDS")
ind3_dapi <- readRDS("data/Fig1/ind3_dapi_rois.RDS")
ind4_dapi <- readRDS("data/Fig1/ind4_dapi_rois.RDS")

```

## Create DAPI Dataframe

```{r Create DAPI Summary matching existing pipeline, message=FALSE, warning=FALSE}

#combine and apply identical filtering/annotation as above for yH2AX
# ind_dapi_all <- bind_rows(ind1_dapi, ind3_dapi, ind4_dapi) %>%
#   mutate(
#     Treatment = case_when(
#       grepl("DOX", ROI_File) ~ "DOX",
#       grepl("DMSO", ROI_File) ~ "VEH",
#       TRUE ~ "Other"
#     ),
#     Timepoint = case_when(
#       grepl("24T", ROI_File) ~ "24T",
#       grepl("24R", ROI_File) ~ "24R",
#       grepl("144R", ROI_File) ~ "144R",
#       TRUE ~ "Other"
#     ),
#     Condition = paste0(Treatment, "_", Timepoint)
#   ) %>%
#   dplyr::filter(Treatment != "Other") %>%
#   dplyr::filter(Magnification == "20X") %>%
#   dplyr::filter(!grepl("^2nd", ROI_File, ignore.case = TRUE))

# saveRDS(ind_dapi_all, "data/Fig1/IF_quant/individual_dapi_all.RDS")
ind_dapi_all <- readRDS("data/Fig1/IF_quant/individual_dapi_all.RDS")

```

## DAPI Intensity Plot

```{r DAPI Intensity - ROI Level, message=FALSE, warning=FALSE}

#plot all of these intensity measurements for Hoechst (DAPI) nuclear staining

plot_dapi_1 <- ind_dapi_all %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  ggplot(aes(x = Condition, y = Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 1.5,
             alpha = 1,
             position = position_identity()) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  ylim(0, 250) +
  labs(title = "Step 1: ROI Level Mean Hoechst Intensity",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_1

# save_plot(plot_dapi_1,
#           filename = "HoechstInt_Step1_ROILevel_Boxplot_EMP",
#           folder = output_folder)

```

## Create an Image Level Summary

```{r collapse DAPI to image level and plot, message=FALSE, warning=FALSE}
#create an image-level summary of all of the relevant ROIs (nuclei)

# dapi_image_level_summary <- ind_dapi_all %>%
#   mutate(Image = trimws(ROI_File)) %>%
#   group_by(Image, Individual, Line, Magnification, Treatment, Timepoint, Condition) %>%
#   dplyr::select(Image, Individual, Line, Magnification, Treatment, Timepoint, Condition, Mean) %>% 
#   summarize(
#     Mean_of_Mean = mean(Mean, na.rm = TRUE),
#     .groups = "drop"
#   )

#this results in a dataframe of the 146 20X images I took total across three cell lines

# saveRDS(dapi_image_level_summary, "data/Fig1/IF_quant/dapi_image_level_summary_all.RDS")
dapi_image_level_summary <- readRDS("data/Fig1/IF_quant/dapi_image_level_summary_all.RDS")

plot_dapi_image <- dapi_image_level_summary %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  ggplot(aes(x = Condition, y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 0.8,
             position = position_identity()) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  ylim(0, 105) +
  labs(title = "Step 2: Image-Level Mean Hoechst Intensity",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_image

# save_plot(plot_dapi_image,
#          filename = "HoechstInt_Step2_ImgLevel_Boxplot",
#          folder = output_folder)

```

## Merge with Metadata

```{r Merge Image-Level Summary with Nuclei Counts, message=FALSE, warning=FALSE}
#now I would like to merge my image_level_metadata made above (on the same samples) with these

image_level_metadata <- readRDS("data/Fig1/IF_quant/image_level_metadata_all.RDS")

# image_merged_dapi_summary <- 
#   image_level_metadata %>% 
#   mutate(
#     Image = str_remove(Image, "_DAPI") #remove the ending that doesn't match here
#   ) %>% 
#   dplyr::select(Individual, Line, Treatment, Timepoint, Condition, Magnification,
#                 Image, Nuclei_Count) %>% 
#   dplyr::left_join(dapi_image_level_summary, by = c("Image", "Individual", "Line", "Magnification",
#                                                     "Treatment", "Timepoint", "Condition"))

# saveRDS(image_merged_dapi_summary, "data/Fig1/image_merged_dapi_summary.RDS")
image_merged_dapi_summary <- readRDS("data/Fig1/image_merged_dapi_summary.RDS")

```

## Assess the number of nuclei per image

In these experiments, Ind1 was very confluent at the time of IF staining. 

Ind4 had fewer cells and appeared sparser across all wells (despite seeding at the same density).

Ind3 was also less confluent than Ind1.

```{r Assess Images for Quality, message=FALSE, warning=FALSE}
#first look at the distribution of nuclei counts to inform a threshold
image_merged_dapi_summary %>%
  ggplot(aes(x = Nuclei_Count)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  facet_wrap(~ Individual) +
  labs(title = "Distribution of Nuclei Counts per Image",
       x = "Nuclei Count", y = "Count") +
  theme_custom()

#also look by condition to see if low counts cluster anywhere
image_merged_dapi_summary %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  ggplot(aes(x = Condition, y = Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), size = 2, alpha = 0.7) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Nuclei Count per Image by Condition",
       x = "Condition", y = "Nuclei Count") +
  theme_custom()

#look at distribution per individual
#84-1 was less confluent and Ind1 was very confluent at the time of staining
image_merged_dapi_summary %>%
  ggplot(aes(x = Nuclei_Count)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  facet_wrap(~ Individual, scales = "free") +  # free scales important here
  labs(title = "Nuclei Count Distribution by Individual",
       x = "Nuclei Count", y = "Count") +
  theme_custom()
```

## Hoechst Quant Step 3

For this step, I will test two scenarios. For the first scenario, I will use the exact same images used in the previous yH2AX quantification (3x per condition)

In the second scenario, I will keep all of the images so as not to artificially select those images with the same numbers of nuclei across conditions, which may mask differences across conditions (if any).

```{r step 3 dapi quant 3 images, warning=FALSE, message=FALSE}
#now, I would like to use the same 3 images per condition to plot Hoechst intensity

pipeline_images <- merged_image_summary_final %>%
  distinct(Image, Individual, Line) %>%
  mutate(Image = gsub("_DAPI", "", Image)) %>%
  dplyr::select(Image, Individual, Line)

#now that I have the set of 54 images from the previous experiment, let's make my final set for dapi quantification like merged_image_summary_final above

image_merged_dapi_summary_final <- image_merged_dapi_summary %>%
  semi_join(pipeline_images, by = c("Image", "Individual", "Line"))

# saveRDS(image_merged_dapi_summary_final, "data/Fig1/image_merged_dapi_summary_final.RDS")

#now plot these in a boxplot with the three images per individual

plot_dapi_3 <- ggplot(image_merged_dapi_summary_final,
                      aes(x = Condition,
                          y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  ylim(0, 105) +
  labs(title = "Step 3: Select 3 Images Per Individual With Consistent Nuclei",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_3

#save plot
# save_plot(plot_dapi_3,
#           filename = "HoechstInt_Step3_3Img_Boxplot_EMP",
#           folder = output_folder)

```

## Step 3 Hoechst Quantification - no image selection

```{r step 3 no image selection, message=FALSE, warning=FALSE}
#now, perform these steps in the same fashion as above on all images processed for this experiment. At the individual level, these numbers of nuclei may differ due to experimental factors including cell confluence (where 84-1 Ind4 is less confluent and therefore has fewer nuclei in all plates).


plot_dapi_3.5 <- ggplot(image_merged_dapi_summary,
                      aes(x = Condition,
                          y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  ylim(0, 105) +
  labs(title = "Step 3: All Images",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_3.5

#save the plots

# save_plot(plot_dapi_3.5,
#           filename = "HoechstInt_Step3.5_ImgLevel_Boxplot_EMP",
#           folder = output_folder)

```

## Step 4 - Use Weighted Mean Intensity to Normalize

```{r weighted mean norm, message=FALSE, warning=FALSE}
#As I visually checked images that had a high mean of mean intensity, I noted that they have much fewer nuclei
#when doing this analysis, I want to check whether high mean intensity is correlated to lower number of nuclei - since calculating the mean of the mean intensity across all images will treat each image as if they have the same number of nuclei

#checks:

ggplot(image_merged_dapi_summary, 
       aes(x = Nuclei_Count, y = Mean_of_Mean, color = Condition)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_x_log10() +  #look at log scale to see relationship easily
  labs(title = "Nuclei Count vs Mean Intensity per Image") +
  theme_custom()

#make a summary table to see
image_merged_dapi_summary %>%
  group_by(Condition) %>%
  summarise(
    r = cor(log1p(Nuclei_Count), Mean_of_Mean, use = "complete.obs"),
    n_images = n(),
    total_nuclei = sum(Nuclei_Count, na.rm = TRUE),
    median_nuclei_per_image = median(Nuclei_Count, na.rm = TRUE)
  )

#these data indicate that higher median nuclei per image are correlated with higher hoechst intensity (which makes sense)


#some conditions also have a lot more images at 20X
#i took a lot of images of Ind3 as I had many regions with few to no cells due to confluency and uneven plating - the number of images being higher could skew results

image_merged_dapi_summary %>%
  group_by(Individual, Condition) %>%
  summarise(
    n_images = n(),
    total_nuclei = sum(Nuclei_Count, na.rm = TRUE),
    median_nuclei_per_image = median(Nuclei_Count, na.rm = TRUE),
    .groups = "drop"
    ) %>%
  arrange(Condition, Individual)

#as I noted earlier, Ind4 was much less confluent - this shows that when I'm taking the mean it's inflating the images from Ind4 with fewer nuclei to give them equal weight to those with hundreds of nuclei per image

#now, I want to use a weighted mean intensity to normalize the results


```

## Step 4 - Collapse to Individual Level with Weighted Mean

```{r step 4 Ind level collapse weight mean, warning=FALSE, message=FALSE}

#collapse to individual level with weighted mean
ind_summary_dapi_4 <- image_merged_dapi_summary %>%
  dplyr::filter(Nuclei_Count >= 10) %>% #exclude imgs with fewer than 10 nuclei
  group_by(Individual, Treatment, Timepoint, Condition) %>%
  summarise(
    n_images = n(),
    total_nuclei = sum(Nuclei_Count, na.rm = TRUE),
    Weight_Mean = weighted.mean(Mean_of_Mean, w = Nuclei_Count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Condition = factor(Condition, levels = condition_levels))

#paired t-test at individual level with BH correction
sig_table_dapi_4 <- ind_summary_dapi_4 %>%
  dplyr::select(Individual, Timepoint, Treatment, Weight_Mean) %>%  # drop Condition
  group_by(Timepoint) %>%
  summarise(
    p = {
      wide <- pick(everything()) %>%
        pivot_wider(names_from = Treatment, values_from = Weight_Mean) %>%
        drop_na(DOX, VEH)
      t.test(wide$DOX, wide$VEH, paired = TRUE)$p.value
    },
    .groups = "drop"
  ) %>%
  mutate(
    p.adj = p.adjust(p, method = "BH"),
    signif = case_when(
      p.adj < 0.001 ~ "***",
      p.adj < 0.01  ~ "**",
      p.adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint),
    y.position = max(ind_summary_dapi_4$Weight_Mean, na.rm = TRUE) * 1.05
  )

#plot at individual level
plot_dapi_4_stats <- ggplot(ind_summary_dapi_4,
                               aes(x = Condition, y = Weight_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), size = 3, alpha = 1) +
  stat_pvalue_manual(sig_table_dapi_4,
                     label = "signif",
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  ylim(0, 105) +
  labs(title = "Step 4: All Images - Individual Level Weighted Mean",
       x = "Condition",
       y = "Weighted Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

#save this plot of all images

plot_dapi_4_stats

# save_plot(plot_dapi_4_stats,
#           filename = "HoechstInt_Step4_AllImg_IndLevel_WeightMean_EMP",
#           folder = output_folder)


```

## Step 4.5 - 3 Images Per Condition

```{r step 4.5 weighted mean 3 images, warning=FALSE, message=FALSE}
#now make another version of the above plot with the three images used in my earlier experiment per timepoint
#these are the same three images

ind_summary_dapi_4_filt <- image_merged_dapi_summary %>%
  dplyr::filter(Nuclei_Count >= 10) %>% 
  semi_join(pipeline_images, by = c("Image", "Individual", "Line")) %>% 
  group_by(Individual, Treatment, Timepoint, Condition) %>%
  summarise(
    n_images = n(),
    total_nuclei = sum(Nuclei_Count, na.rm = TRUE),
    Weight_Mean = weighted.mean(Mean_of_Mean, w = Nuclei_Count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(Condition = factor(Condition, levels = condition_levels))

sig_table_dapi_4_filt <- ind_summary_dapi_4_filt %>%
  dplyr::select(Individual, Timepoint, Treatment, Weight_Mean) %>%
  group_by(Timepoint) %>%
  summarise(
    p = {
      wide <- pick(everything()) %>%
        pivot_wider(names_from = Treatment, values_from = Weight_Mean) %>%
        drop_na(DOX, VEH)
      t.test(wide$DOX, wide$VEH, paired = TRUE)$p.value
    },
    .groups = "drop"
  ) %>%
  mutate(
    p.adj = p.adjust(p, method = "BH"),
    signif = case_when(
      p.adj < 0.001 ~ "***",
      p.adj < 0.01  ~ "**",
      p.adj < 0.05  ~ "*",
      TRUE          ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint),
    y.position = max(ind_summary_dapi_4_filt$Weight_Mean, na.rm = TRUE) * 1.05
  )

plot_dapi_4_filt_stats <- ggplot(ind_summary_dapi_4_filt,
                                  aes(x = Condition, y = Weight_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), size = 3, alpha = 1) +
  stat_pvalue_manual(sig_table_dapi_4_filt,
                     label = "signif",
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  ylim(0, 105) +
  labs(title = "Step 4.5: 3 Images - Individual Level Weighted Mean",
       x = "Condition",
       y = "Weighted Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_4_filt_stats

sig_table_dapi_4_filt

# save_plot(plot_dapi_4_filt_stats,
#           filename = "HoechstInt_Step4_3Img_IndLevel_WeightMean_EMP",
#           folder = output_folder)


#should I be testing each individual separately? then I only have one point though so I'd have to use statistics from image-level comparisons 

```


## Step 5 - Treatment-Level Comparisons

```{r tx level comparisons hoechst, message=FALSE, warning=FALSE}
#now, I want to take these comparisons and add another layer where I'm comparing between DOX treated at different timepoints

ind_summary_dapi_step5_dox <- ind_summary_dapi_4_filt %>% 
  group_by(Individual, Treatment, Timepoint) %>% 
  summarise(
    Weight_Mean = Weight_Mean,
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )

#set the condition levels
condition_levels <- c("DOX_24T", "VEH_24T",
                      "DOX_24R", "VEH_24R",
                      "DOX_144R", "VEH_144R")

dox_conditions <- c("DOX_24T", "DOX_24R", "DOX_144R")

ind_summary_dapi_step5_dox <- ind_summary_dapi_step5_dox %>% 
  mutate(Condition = factor(Condition, levels = condition_levels))

#perform paired t-tests between DOX timepoints
timepoint_pairs <- list(
  c("24T","24R"),
  c("24R","144R"),
  c("24T","144R")
)

ttest_table_dox_time_dapi <- purrr::map_df(seq_along(timepoint_pairs), function(i) {
  tp <- timepoint_pairs[[i]]
  
  df_wide <- ind_summary_dapi_step5_dox %>% 
    filter(Treatment == "DOX", Timepoint %in% tp) %>% 
    group_by(Individual, Timepoint) %>% 
    summarise(Weight_Mean = mean(Weight_Mean, na.rm = TRUE), .groups = "drop") %>% 
    pivot_wider(names_from = Timepoint, values_from = Weight_Mean) %>% 
    drop_na()
  
  base_y <- max(ind_summary_dapi_step5_dox$Weight_Mean,
                na.rm = TRUE) * 0.9
  
  y_increment <- max(ind_summary_dapi_step5_dox$Weight_Mean,
                     na.rm = TRUE) * 0.05
  
  y_pos <- base_y + y_increment * i
  
  if(nrow(df_wide) < 2) {
    tibble(
      group1 = paste0("DOX_", tp[1]),
      group2 = paste0("DOX_", tp[2]),
      p.value = NA_real_,
      p.signif = "NA",
      y.position = y_pos
    )
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1 = paste0("DOX_", tp[1]),
      group2 = paste0("DOX_", tp[2]),
      p.value = test$p.value,
      p.signif = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01 ~ "**",
        test$p.value < 0.05 ~ "*",
        TRUE ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to numeric positions for stat_pvalue_manual + add adj. p value (BH corrected)
ttest_table_dox_time_dapi <- ttest_table_dox_time_dapi %>% 
  mutate(
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

#manually set y positions for each comparison (so it is visible like above)
max_val <- max(ind_summary_dapi_step5_dox$Weight_Mean, na.rm = TRUE)

ttest_table_dox_time_dapi <- ttest_table_dox_time_dapi %>% 
  mutate(
    y.position = case_when(
      group1 == "DOX_24T" & group2 == "DOX_24R" ~ max_val * 1.15, #middle
      group1 == "DOX_24T" & group2 == "DOX_144R" ~ max_val * 1.25,  #top
      group1 == "DOX_24R" & group2 == "DOX_144R" ~ max_val * 1.05 #bottom
      ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )


#now plot the DOX timepoint comparisons
plot_step5_dox_time_dapi <- ggplot(ind_summary_dapi_step5_dox,
                                   aes(x = Condition,
                                       y = Weight_Mean)) +
  geom_boxplot(aes(fill = Condition),
               show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 1,
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 105)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_dox_time_dapi,
                     label = "p.signif",
                     xmin = "group1_num",
                     xmax = "group2_num",
                     y.position = "y.position",
                     tip.length = 0.02) +
  labs(title = "Step 5: DOX Timepoint Comparisons Hoechst Intensity (3 Images)",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_step5_dox_time_dapi

#print the stats table as well for visibility
ttest_table_dox_time_dapi

#save the plot
# save_plot(plot_step5_dox_time_dapi,
#           filename = "HoechstInt_Step5_IndLev_DOXComp_EMP",
#           folder = output_folder)

```

## Step 5 - Vehicle-Level Comparisons

```{r compare between vehicle timepoints too, warning=FALSE, message=FALSE}

#now I want to do the same as above but test between vehicle timepoints

ind_summary_dapi_step5_veh <- ind_summary_dapi_4_filt %>% 
  group_by(Individual, Treatment, Timepoint) %>% 
  summarise(
    Weight_Mean = Weight_Mean,
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )

#set the condition levels
condition_levels <- c("DOX_24T", "VEH_24T",
                      "DOX_24R", "VEH_24R",
                      "DOX_144R", "VEH_144R")

veh_conditions <- c("VEH_24T", "VEH_24R", "VEH_144R")

ind_summary_dapi_step5_veh <- ind_summary_dapi_step5_veh %>% 
  mutate(Condition = factor(Condition, levels = condition_levels))

#perform paired t-tests between VEH timepoints
timepoint_pairs <- list(
  c("24T","24R"),
  c("24R","144R"),
  c("24T","144R")
)

ttest_table_veh_time_dapi <- purrr::map_df(seq_along(timepoint_pairs), function(i) {
  tp <- timepoint_pairs[[i]]
  
  df_wide <- ind_summary_dapi_step5_veh %>% 
    filter(Treatment == "VEH", Timepoint %in% tp) %>% 
    group_by(Individual, Timepoint) %>% 
    summarise(Weight_Mean = mean(Weight_Mean, na.rm = TRUE), .groups = "drop") %>% 
    pivot_wider(names_from = Timepoint, values_from = Weight_Mean) %>% 
    drop_na()
  
  base_y <- max(ind_summary_dapi_step5_veh$Weight_Mean,
                na.rm = TRUE) * 0.9
  
  y_increment <- max(ind_summary_dapi_step5_veh$Weight_Mean,
                     na.rm = TRUE) * 0.05
  
  y_pos <- base_y + y_increment * i
  
  if(nrow(df_wide) < 2) {
    tibble(
      group1 = paste0("VEH_", tp[1]),
      group2 = paste0("VEH_", tp[2]),
      p.value = NA_real_,
      p.signif = "NA",
      y.position = y_pos
    )
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1 = paste0("VEH_", tp[1]),
      group2 = paste0("VEH_", tp[2]),
      p.value = test$p.value,
      p.signif = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01 ~ "**",
        test$p.value < 0.05 ~ "*",
        TRUE ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to numeric positions for stat_pvalue_manual + add adj. p value (BH corrected)
ttest_table_veh_time_dapi <- ttest_table_veh_time_dapi %>% 
  mutate(
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

#manually set y positions for each comparison (so it is visible like above)
max_val <- max(ind_summary_dapi_step5_veh$Weight_Mean, na.rm = TRUE)

ttest_table_veh_time_dapi <- ttest_table_veh_time_dapi %>% 
  mutate(
    y.position = case_when(
      group1 == "VEH_24T" & group2 == "VEH_24R" ~ max_val * 1.15, #middle
      group1 == "VEH_24T" & group2 == "VEH_144R" ~ max_val * 1.25,  #top
      group1 == "VEH_24R" & group2 == "VEH_144R" ~ max_val * 1.05 #bottom
      ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )


#now plot the VEH timepoint comparisons
plot_step5_veh_time_dapi <- ggplot(ind_summary_dapi_step5_veh,
                                   aes(x = Condition,
                                       y = Weight_Mean)) +
  geom_boxplot(aes(fill = Condition),
               show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 1,
             position = position_identity()) +
  scale_y_continuous(limits = c(0, 105)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_veh_time_dapi,
                     label = "p.signif",
                     xmin = "group1_num",
                     xmax = "group2_num",
                     y.position = "y.position",
                     tip.length = 0.02) +
  labs(title = "Step 5: VEH Timepoint Comparisons Hoechst Intensity (3 Images)",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_step5_veh_time_dapi

#print the stats table as well for visibility
ttest_table_veh_time_dapi

#save the plot
# save_plot(plot_step5_veh_time_dapi,
#           filename = "HoechstInt_Step5_IndLev_VEHComp_EMP",
#           folder = output_folder)

```



# Assess Number of Nuclei Per Condition

I want to understand whether there are differences in nuclei numbers across conditions.

```{r Nuclei Counts per Image, message=FALSE, warning=FALSE}

read_nuclei_counts <- function(cell_line, individual_label) {
  summary_path <- paste0("D:/DXR Project/IF yH2AX/if yh2ax dxr ", cell_line, 
                         "/DAPI intensity/nuclei_counts_summary.csv")
  read.csv(summary_path) %>%
    mutate(Individual = individual_label,
           Line = cell_line)
}

nuclei_counts_imgs <- bind_rows(
  read_nuclei_counts("87-1", "Ind1"),
  read_nuclei_counts("75-1", "Ind3"),
  read_nuclei_counts("84-1", "Ind4")
) %>%
  mutate(
    Magnification = case_when(
      grepl("20X", ImageName, ignore.case = TRUE) ~ "20X",
      grepl("10X", ImageName, ignore.case = TRUE) ~ "10X",
      grepl("40X", ImageName, ignore.case = TRUE) ~ "40X",
      TRUE ~ "20X" #Ind4 is all 20X and doesn't have it in the name
    ),
    Treatment = case_when(
      grepl("DOX", ImageName) ~ "DOX",
      grepl("DMSO", ImageName) ~ "VEH",
      TRUE ~ "Other"
    ),
    Timepoint = case_when(
      grepl("24T", ImageName) ~ "24T",
      grepl("24R", ImageName) ~ "24R",
      grepl("144R", ImageName) ~ "144R",
      TRUE ~ "Other"
    ),
    Condition = paste0(Treatment, "_", Timepoint)
  ) %>%
  dplyr::filter(Magnification == "20X") %>%
  dplyr::filter(Treatment != "Other") %>%
  dplyr::filter(!grepl("^2nd", ImageName, ignore.case = TRUE)) 

#check the number of images per condition (all raw images)
image_counts <- nuclei_counts_imgs %>%
  group_by(Individual, Condition) %>%
  summarise(n_images = n(), .groups = "drop")

#now standardize to the lowest max across conditions with random sampling
min_max_n <- image_counts %>%
  pull(n_images) %>%
  min()

#print this output
cat("Images per Individual x Condition (before standardization):\n")
print(image_counts)
cat(sprintf("\nStandardizing to %d images per Individual x Condition.\n", min_max_n))

#now randomly sample these images to get an equal number
set.seed(9196)  #set for reproducibility
nuclei_counts_imgs_std <- nuclei_counts_imgs %>%
  group_by(Individual, Condition) %>%
  slice_sample(n = min_max_n) %>%
  ungroup()

#non-standardized overview table of all images — one row per image
nuclei_counts_imgs %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  arrange(Individual, Condition, ImageName) %>%
  dplyr::select(Individual, Line, Condition, ImageName, NucleiCount) %>%
  knitr::kable(caption = "Nuclei Count per Image") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

#standardized output table with sampled images
nuclei_counts_imgs_std %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  arrange(Individual, Condition, ImageName) %>%
  dplyr::select(Individual, Line, Condition, ImageName, NucleiCount) %>%
  knitr::kable(caption = sprintf("Nuclei Count per Image (standardized to %d images per group)", min_max_n)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)

#check on which combinations exist in the data and ensure that each individual and condition are present with 3 images each
cat("Individuals present in standardized data:\n")
print(unique(nuclei_counts_imgs_std$Individual))

cat("\nImage counts per Individual x Condition (standardized):\n")
nuclei_counts_imgs_std %>%
  group_by(Individual, Condition) %>%
  summarise(n = n(), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = Condition, values_from = n, values_fill = 0) %>%
  print()

```

```{r nuclei counts pipeline images, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# nuclei_counts_imgs <- bind_rows(
#   read_nuclei_counts("87-1", "Ind1"),
#   read_nuclei_counts("75-1", "Ind3"),
#   read_nuclei_counts("84-1", "Ind4")
# ) %>%
#   mutate(
#     Magnification = case_when(
#       grepl("20X", ImageName, ignore.case = TRUE) ~ "20X",
#       grepl("10X", ImageName, ignore.case = TRUE) ~ "10X",
#       grepl("40X", ImageName, ignore.case = TRUE) ~ "40X",
#       TRUE ~ "Other"
#     ),
#     Treatment = case_when(
#       grepl("DOX", ImageName) ~ "DOX",
#       grepl("DMSO", ImageName) ~ "VEH",
#       TRUE ~ "Other"
#     ),
#     Timepoint = case_when(
#       grepl("24T", ImageName) ~ "24T",
#       grepl("24R", ImageName) ~ "24R",
#       grepl("144R", ImageName) ~ "144R",
#       TRUE ~ "Other"
#     ),
#     Condition = paste0(Treatment, "_", Timepoint)
#   ) %>%
#   dplyr::filter(Magnification == "20X") %>%
#   dplyr::filter(Treatment != "Other") %>%
#   dplyr::filter(!grepl("^2nd", ImageName, ignore.case = TRUE)) %>% 
#   semi_join(pipeline_images, by = c("ImageName" = "ROI_File", "Individual", "Line"))

# saveRDS(nuclei_counts_imgs, file = "data/Fig1/nuclei_counts_imgs.RDS")
# nuclei_counts_imgs <- readRDS("data/Fig1/nuclei_counts_imgs.RDS")

```

## Nuclei Count Assessment - Step 1: Image-Level Comparisons

```{r assess nuclei count from raw images across all cond, warning=FALSE, message=FALSE}
#assess whether nucleus count decreases over time from raw images

plot_nuc_count_std <- ggplot(nuclei_counts_imgs_std,
                      aes(x = Condition,
                          y = NucleiCount)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Assess # of Nuclei per Condition, Image Level (3 images)",
       x = "Condition",
       y = "Number of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_nuc_count_std

#now do stats for these to see if there are any differences over time
ttests_nuc_count_std <- nuclei_counts_imgs_std %>% 
  group_by(Timepoint) %>% 
  t_test(NucleiCount ~ Treatment) %>% 
  mutate(
    y.position = max(nuclei_counts_imgs_std$NucleiCount, na.rm = TRUE) * 1
  )

sig_table_nuc_count_std <- nuclei_counts_imgs_std %>% 
  group_by(Timepoint) %>% 
  summarise(
    p = t.test(NucleiCount ~ Treatment)$p.value,
    y.position = max(NucleiCount, na.rm = TRUE) * 1.05,
    .groups = "drop"
  ) %>% 
  mutate(
    signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint)
  )

ttests_nuc_count_std

#now plot with these ttest statistics above the plot
plot_nuc_count_stats_std <- ggplot(nuclei_counts_imgs_std,
                            aes(x = Condition,
                                y = NucleiCount)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  stat_pvalue_manual(sig_table_nuc_count_std,
                     label = "signif",
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Assess # of Nuclei by Condition, Image-Level (3 images)",
       x = "Condition",
       y = "Number of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_nuc_count_stats_std

# save_plot(plot_nuc_count_stats_std,
#           filename = "NucleiCount_3Imgs_Boxplot_ImgLevel_EMP",
#           folder = output_folder)

```

## Nuclei Count Assessment - Step 2: Individual-Level Comparisons


```{r ind level nuclei count quant, warning=FALSE, message=FALSE}
#now I would like to assess these numbers of nuclei at an individual level

#first, collapse the data by individual
ind_summary_nuc_count_std <- nuclei_counts_imgs_std %>% 
  group_by(Individual, Treatment, Timepoint, Magnification) %>% 
  summarise(
    Median_Nuclei_Count = median(NucleiCount, na.rm = TRUE),
    Mean_Nuclei_count = mean(NucleiCount, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )


#assess the mean and the median, but first do stats
ttests_nuc_count_ind_std <- ttests_nuc_count_std %>% 
  filter(Timepoint %in% c("24T", "24R", "144R")) %>% 
  mutate(
    group1 = paste0("DOX_", Timepoint),
    group2 = paste0("VEH_", Timepoint),
    y.position = max(nuclei_counts_imgs_std$NucleiCount, na.rm = TRUE) * 1.05,
    
    #add significance stars as defined earlier
    p.signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

#now plot these individual-level measurements with stats

#first plot mean number of nuclei

plot_nuc_counts_ind_mean_std <- ggplot(ind_summary_nuc_count_std, 
                     aes(x = Condition, y = Mean_Nuclei_count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1, 
             position = position_identity()) +
  # scale_y_continuous(limits = c(0, 105)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttests_nuc_count_ind_std,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  ylim(0, 400) +
  labs(title = "Assess # of Nuclei per Condition, Ind Level (3 Images)",
       x = "Condition", 
       y = "Mean # of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_nuc_counts_ind_mean_std


#next, plot median number of nuclei

plot_nuc_counts_ind_med_std <- ggplot(ind_summary_nuc_count_std, 
                     aes(x = Condition, y = Median_Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1, 
             position = position_identity()) +
  # scale_y_continuous(limits = c(0, 105)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttests_nuc_count_ind_std,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  ylim(0, 400) +
  labs(title = "Assess # of Nuclei per Condition, Ind Level (3 Images)",
       x = "Condition", 
       y = "Median # of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_nuc_counts_ind_med_std

#save plots

# save_plot(plot_nuc_counts_ind_mean_std,
#           filename = "NucleiCount_IndLevel_Mean_3Imgs_EMP",
#           folder = output_folder)
# 
# save_plot(plot_nuc_counts_ind_med_std,
#           filename = "NucleiCount_IndLevel_Median_3Imgs_EMP",
#           folder = output_folder)

```

## Nuclei Count Assessment - Testing between timepoints within DOX or VEH

### Mean Nuclei Counts - DOX timepoints tests

```{r nuclei counts testing between timepoints within condition mean dox, warning=FALSE, message=FALSE}
#define the condition levels
condition_levels <- c("DOX_24T", "VEH_24T",
                      "DOX_24R", "VEH_24R",
                      "DOX_144R", "VEH_144R")

dox_conditions <- c("DOX_24T", "DOX_24R", "DOX_144R")
veh_conditions <- c("VEH_24T", "VEH_24R", "VEH_144R")

timepoint_pairs <- list(
  c("24T", "24R"),
  c("24R", "144R"),
  c("24T", "144R")
)

ind_summary_nuc_count_std <- ind_summary_nuc_count_std %>% 
  mutate(Condition = factor(Condition, levels = condition_levels))

max_val_nuc_mean <- max(ind_summary_nuc_count_std$Mean_Nuclei_count, na.rm = TRUE)

#perform t-tests between DOX conditions for the Mean nuclei count
ttest_table_dox_time_nuc_mean <- purrr::map_df(seq_along(timepoint_pairs), function(i) {
  tp <- timepoint_pairs[[i]]
  
  df_wide <- ind_summary_nuc_count_std %>% 
    filter(Treatment == "DOX", Timepoint %in% tp) %>% 
    group_by(Individual, Timepoint) %>% 
    summarise(Mean_Nuclei_count = mean(Mean_Nuclei_count, na.rm = TRUE), .groups = "drop") %>% 
    pivot_wider(names_from = Timepoint, values_from = Mean_Nuclei_count) %>% 
    drop_na()
  
  base_y     <- max_val_nuc_mean * 0.9
  y_increment <- max_val_nuc_mean * 0.05
  y_pos      <- base_y + y_increment * i
  
  if (nrow(df_wide) < 2) {
    tibble(group1 = paste0("DOX_", tp[1]), group2 = paste0("DOX_", tp[2]),
           p.value = NA_real_, p.signif = "NA", y.position = y_pos)
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1     = paste0("DOX_", tp[1]),
      group2     = paste0("DOX_", tp[2]),
      p.value    = test$p.value,
      p.signif   = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01  ~ "**",
        test$p.value < 0.05  ~ "*",
        TRUE                 ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to positions on boxplot, add adj. p value, and manually set y positions
ttest_table_dox_time_nuc_mean <- ttest_table_dox_time_nuc_mean %>% 
  mutate(
    y.position = case_when(
      group1 == "DOX_24T" & group2 == "DOX_24R"   ~ max_val_nuc_mean * 1.15,
      group1 == "DOX_24T" & group2 == "DOX_144R"  ~ max_val_nuc_mean * 1.25,
      group1 == "DOX_24R" & group2 == "DOX_144R"  ~ max_val_nuc_mean * 1.05
    ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

ttest_table_dox_time_nuc_mean

#now plot this onto a boxplot as before with these within-condition stats
plot_dox_time_nuc_mean <- ggplot(ind_summary_nuc_count_std,
                            aes(x = Condition, y = Mean_Nuclei_count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3, alpha = 1, position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_dox_time_nuc_mean,
                     label = "p.signif",
                     xmin = "group1_num",
                     xmax = "group2_num",
                     y.position = "y.position",
                     tip.length = 0.02) +
  ylim(0, max_val_nuc_mean * 1.35) +
  labs(title = "DOX Timepoint Comparisons Mean Nuclei Count (3 Images)",
       x = "Condition",
       y = "Mean # of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dox_time_nuc_mean

# save_plot(plot_dox_time_nuc_mean,
#           filename = "NucleiCount_Mean_btwDOX_EMP",
#           folder = output_folder)

```


### Mean Nuclei Counts - VEH timepoints tests

```{r nuclei counts testing between timepoints within condition veh mean, warning=FALSE, message=FALSE}

max_val_nuc_mean <- max(ind_summary_nuc_count_std$Mean_Nuclei_count, na.rm = TRUE)

#perform t-tests between VEH conditions for the Mean nuclei count
ttest_table_veh_time_nuc_mean <- purrr::map_df(seq_along(timepoint_pairs), function(i) {
  tp <- timepoint_pairs[[i]]
  
  df_wide <- ind_summary_nuc_count_std %>% 
    filter(Treatment == "VEH", Timepoint %in% tp) %>% 
    group_by(Individual, Timepoint) %>% 
    summarise(Mean_Nuclei_count = mean(Mean_Nuclei_count, na.rm = TRUE), .groups = "drop") %>% 
    pivot_wider(names_from = Timepoint, values_from = Mean_Nuclei_count) %>% 
    drop_na()
  
  base_y     <- max_val_nuc_mean * 0.9
  y_increment <- max_val_nuc_mean * 0.05
  y_pos      <- base_y + y_increment * i
  
  if (nrow(df_wide) < 2) {
    tibble(group1 = paste0("VEH_", tp[1]), 
           group2 = paste0("VEH_", tp[2]),
           p.value = NA_real_, p.signif = "NA", y.position = y_pos)
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1     = paste0("VEH_", tp[1]),
      group2     = paste0("VEH_", tp[2]),
      p.value    = test$p.value,
      p.signif   = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01  ~ "**",
        test$p.value < 0.05  ~ "*",
        TRUE                 ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to positions on boxplot, add adj. p value, and manually set y positions
ttest_table_veh_time_nuc_mean <- ttest_table_veh_time_nuc_mean %>% 
  mutate(
    y.position = case_when(
      group1 == "VEH_24T" & group2 == "VEH_24R"   ~ max_val_nuc_mean * 1.15,
      group1 == "VEH_24T" & group2 == "VEH_144R"  ~ max_val_nuc_mean * 1.25,
      group1 == "VEH_24R" & group2 == "VEH_144R"  ~ max_val_nuc_mean * 1.05
    ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

ttest_table_veh_time_nuc_mean

#now plot this onto a boxplot as before with these within-condition stats
plot_veh_time_nuc_mean <- ggplot(ind_summary_nuc_count_std,
                            aes(x = Condition, y = Mean_Nuclei_count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3, alpha = 1, position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_veh_time_nuc_mean,
                     label = "p.signif",
                     xmin = "group1_num",
                     xmax = "group2_num",
                     y.position = "y.position",
                     tip.length = 0.02) +
  ylim(0, max_val_nuc_mean * 1.35) +
  labs(title = "VEH Timepoint Comparisons Mean Nuclei Count (3 Images)",
       x = "Condition",
       y = "Mean # of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_veh_time_nuc_mean


# save_plot(plot_veh_time_nuc_mean,
#           filename = "NucleiCount_Mean_btwVEH_EMP",
#           folder = output_folder)
```


### Median Nuclei Counts - DOX timepoints tests

```{r nuclei counts testing between timepoints within condition median dox, warning=FALSE, message=FALSE}

max_val_nuc_med <- max(ind_summary_nuc_count_std$Median_Nuclei_Count, na.rm = TRUE)

#perform t-tests between DOX conditions for the Median nuclei count
ttest_table_dox_time_nuc_med <- purrr::map_df(seq_along(timepoint_pairs), function(i) {
  tp <- timepoint_pairs[[i]]
  
  df_wide <- ind_summary_nuc_count_std %>% 
    filter(Treatment == "DOX", Timepoint %in% tp) %>% 
    group_by(Individual, Timepoint) %>% 
    summarise(Median_Nuclei_Count = median(Median_Nuclei_Count, na.rm = TRUE), .groups = "drop") %>% 
    pivot_wider(names_from = Timepoint, values_from = Median_Nuclei_Count) %>% 
    drop_na()
  
  base_y     <- max_val_nuc_med * 0.9
  y_increment <- max_val_nuc_med * 0.05
  y_pos      <- base_y + y_increment * i
  
  if (nrow(df_wide) < 2) {
    tibble(group1 = paste0("DOX_", tp[1]), group2 = paste0("DOX_", tp[2]),
           p.value = NA_real_, p.signif = "NA", y.position = y_pos)
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1     = paste0("DOX_", tp[1]),
      group2     = paste0("DOX_", tp[2]),
      p.value    = test$p.value,
      p.signif   = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01  ~ "**",
        test$p.value < 0.05  ~ "*",
        TRUE                 ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to positions on boxplot, add adj. p value, and manually set y positions
ttest_table_dox_time_nuc_med <- ttest_table_dox_time_nuc_med %>% 
  mutate(
    y.position = case_when(
      group1 == "DOX_24T" & group2 == "DOX_24R"   ~ max_val_nuc_med * 1.15,
      group1 == "DOX_24T" & group2 == "DOX_144R"  ~ max_val_nuc_med * 1.25,
      group1 == "DOX_24R" & group2 == "DOX_144R"  ~ max_val_nuc_med * 1.05
    ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

ttest_table_dox_time_nuc_med

#now plot this onto a boxplot as before with these within-condition stats
plot_dox_time_nuc_med <- ggplot(ind_summary_nuc_count_std,
                            aes(x = Condition, y = Median_Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3, alpha = 1, position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_dox_time_nuc_med,
                     label = "p.signif",
                     xmin = "group1_num",
                     xmax = "group2_num",
                     y.position = "y.position",
                     tip.length = 0.02) +
  ylim(0, max_val_nuc_mean * 1.35) +
  labs(title = "DOX Timepoint Comparisons Median Nuclei Count (3 Images)",
       x = "Condition",
       y = "Median # of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dox_time_nuc_med

# save_plot(plot_dox_time_nuc_med,
#           filename = "NucleiCount_Median_btwDOX_EMP",
#           folder = output_folder)

```


### Median Nuclei Counts - VEH timepoints tests

```{r nuclei counts testing between timepoints within condition median veh, warning=FALSE, message=FALSE}

max_val_nuc_med <- max(ind_summary_nuc_count_std$Median_Nuclei_Count, na.rm = TRUE)

#perform t-tests between VEH conditions for the Median nuclei count
ttest_table_veh_time_nuc_med <- purrr::map_df(seq_along(timepoint_pairs), function(i) {
  tp <- timepoint_pairs[[i]]
  
  df_wide <- ind_summary_nuc_count_std %>% 
    filter(Treatment == "VEH", Timepoint %in% tp) %>% 
    group_by(Individual, Timepoint) %>% 
    summarise(Median_Nuclei_Count = median(Median_Nuclei_Count, na.rm = TRUE), .groups = "drop") %>% 
    pivot_wider(names_from = Timepoint, values_from = Median_Nuclei_Count) %>% 
    drop_na()
  
  base_y     <- max_val_nuc_med * 0.9
  y_increment <- max_val_nuc_med * 0.05
  y_pos      <- base_y + y_increment * i
  
  if (nrow(df_wide) < 2) {
    tibble(group1 = paste0("VEH_", tp[1]), 
           group2 = paste0("VEH_", tp[2]),
           p.value = NA_real_, p.signif = "NA", y.position = y_pos)
  } else {
    test <- t.test(df_wide[[tp[1]]], df_wide[[tp[2]]], paired = TRUE)
    tibble(
      group1     = paste0("VEH_", tp[1]),
      group2     = paste0("VEH_", tp[2]),
      p.value    = test$p.value,
      p.signif   = case_when(
        test$p.value < 0.001 ~ "***",
        test$p.value < 0.01  ~ "**",
        test$p.value < 0.05  ~ "*",
        TRUE                 ~ "ns"
      ),
      y.position = y_pos
    )
  }
})

#map to positions on boxplot, add adj. p value, and manually set y positions
ttest_table_veh_time_nuc_med <- ttest_table_veh_time_nuc_med %>% 
  mutate(
    y.position = case_when(
      group1 == "VEH_24T" & group2 == "VEH_24R"   ~ max_val_nuc_med * 1.15,
      group1 == "VEH_24T" & group2 == "VEH_144R"  ~ max_val_nuc_med * 1.25,
      group1 == "VEH_24R" & group2 == "VEH_144R"  ~ max_val_nuc_med * 1.05
    ),
    group1_num = as.numeric(factor(group1, levels = condition_levels)),
    group2_num = as.numeric(factor(group2, levels = condition_levels))
  )

ttest_table_veh_time_nuc_med

#now plot this onto a boxplot as before with these within-condition stats
plot_veh_time_nuc_med <- ggplot(ind_summary_nuc_count_std,
                            aes(x = Condition, y = Median_Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3, alpha = 1, position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttest_table_dox_time_nuc_med,
                     label = "p.signif",
                     xmin = "group1_num",
                     xmax = "group2_num",
                     y.position = "y.position",
                     tip.length = 0.02) +
  ylim(0, max_val_nuc_mean * 1.35) +
  labs(title = "VEH Timepoint Comparisons Median Nuclei Count (3 Images)",
       x = "Condition",
       y = "Median # of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_veh_time_nuc_med

# save_plot(plot_veh_time_nuc_med,
#           filename = "NucleiCount_Median_btwVEH_EMP",
#           folder = output_folder)
```


# Figure 1E - yH2AX IF Images (no code)

Images from individual 87-1 (Ind1).
