---
title: "Figure 4 Analysis"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{css, echo=FALSE}
pre {
  max-height: 400px;
  overflow-y: auto;
}

pre[class] {
  max-height: 200px;
  
}
```


# Libraries

Load necessary libraries for analysis.

```{r Necessary Libraries, message=FALSE, warning=FALSE, echo=TRUE}
library(tidyverse)
library(readxl)
library(readr)
```

Define my custom plot theme

```{r Custom plot theme, echo=TRUE, message=FALSE, warning=FALSE}

# Define the custom theme
# plot_theme_custom <- function() {
#   theme_minimal() +
#     theme(
#       #line for x and y axis
#       axis.line = element_line(linewidth = 1,
#                                color = "black"),
# 
#       #axis ticks only on x and y, length standard
#       axis.ticks.x = element_line(color = "black",
#                                 linewidth = 1),
#       axis.ticks.y = element_line(color = "black",
#                                 linewidth = 1),
#       axis.ticks.length = unit(0.05, "in"),
# 
#       #text and font
#       axis.text = element_text(color = "black",
#                                family = "Arial",
#                                size = 7),
#       axis.title = element_text(color = "black",
#                                 family = "Arial",
#                                 size = 9),
#       legend.text = element_text(color = "black",
#                                  family = "Arial",
#                                  size = 7),
#       legend.title = element_text(color = "black",
#                                   family = "Arial",
#                                   size = 9),
#       plot.title = element_text(color = "black",
#                                 family = "Arial",
#                                 size = 10),
# 
#       #blank background and border
#       panel.background = element_blank(),
#       panel.border = element_blank(),
# 
#       #gridlines for alignment
#       panel.grid.major = element_line(color = "grey80", linewidth = 0.5),  #grey major grid for align in illus
#       panel.grid.minor = element_line(color = "grey90", linewidth = 0.5) #grey minor grid for align in illus
#     )
# }

# saveRDS(plot_theme_custom, "data/plot_theme_custom.RDS")

# theme_custom <- readRDS("data/plot_theme_custom.RDS")
theme_custom <- readRDS("data/plot_theme_custom_update.RDS")
```


Define a function to save plots as .pdf and .png

```{r pdf png saving function, echo=TRUE, message=FALSE, warning=FALSE}

save_plot <- function(plot, filename, 
                      folder = ".", 
                      width = 8, 
                      height = 6, 
                      units = "in", 
                      dpi = 300, 
                      add_date = TRUE) {
  if (missing(filename)) stop("Please provide a filename (without extension) for the plot.")

  date_str <- if (add_date) paste0("_", format(Sys.Date(), "%y%m%d")) else ""
  pdf_file <- file.path(folder, paste0(filename, date_str, ".pdf"))
  png_file <- file.path(folder, paste0(filename, date_str, ".png"))
  ggsave(filename = pdf_file, plot = plot, device = cairo_pdf, width = width, height = height, units = units, bg = "transparent")
  ggsave(filename = png_file, plot = plot, device = "png", width = width, height = height, units = units, dpi = dpi, bg = "transparent")
  message("Saved plot as Cairo PDF: ", pdf_file)
  message("Saved plot as PNG: ", png_file)
}

output_folder <- "C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/Fig pdfs"


#save plot function created
#to use: just define the plot name, filename_base, width, height
```

# Compare Response Clusters to Known TOP2i Treatments

## Read in TOP2i DEGs from Matthews et al., Paul et al.

Matthews ER, Johnson OD, Horn KJ, Gutiérrez JA, Powell SR, et al. (2024) Anthracyclines induce cardiotoxicity through a shared gene expression response signature. PLOS Genetics 20(2): e1011164. https://doi.org/10.1371/journal.pgen.1011164


Topoisomerase II inhibitors CX-5461 and Doxorubicin differ in their cardiotoxicity profiles
Sayan Paul, José A. Gutiérrez, Alyssa R. Bogar, Michelle C. Ward. bioRxiv 2025.07.20.665626; doi: https://doi.org/10.1101/2025.07.20.665626


```{r TOP2i DEG Lists, message=FALSE, warning=FALSE}

#These gene lists were obtained from:
#Matthews ER, Johnson OD, Horn KJ, Gutiérrez JA, Powell SR, et al. (2024) Anthracyclines induce cardiotoxicity through a shared gene expression response signature. PLOS Genetics 20(2): e1011164. https://doi.org/10.1371/journal.pgen.1011164

DEGs_RM_list <- readRDS("data/Fig4/DEGs_TOP2i_RM.RDS")

#only keep the conditions that I want to compare to
keep_conditions <- c("sigVDA24", "sigVDX24", "sigVEP24", "sigVMT24")

DEGs_RM_24 <- DEGs_RM_list[keep_conditions]

#change the first column to Entrez_ID for all sets
DEGs_RM_24 <- lapply(DEGs_RM_24, function(df) {
  # Rename the first column to Entrez_ID, regardless of current name
  colnames(df)[1] <- "Entrez_ID"
  return(df)
})

#DOX
DOX_RM_DEGs <- DEGs_RM_24$sigVDX24 %>% 
  as.data.frame()

dim(DOX_RM_DEGs)
#6645 genes

#DNR
DNR_RM_DEGs <- DEGs_RM_24$sigVDA24 %>% 
  as.data.frame()

dim(DNR_RM_DEGs)
#7017 genes

#EPI
EPI_RM_DEGs <- DEGs_RM_24$sigVEP24 %>% 
  as.data.frame()

dim(EPI_RM_DEGs)
#6328 genes

#MTX
MTX_RM_DEGs <- DEGs_RM_24$sigVMT24 %>% 
  as.data.frame()

dim(MTX_RM_DEGs)
#1115 genes

#now combine these all into one dataframe for later

DEGs_RM_24_df <- dplyr::bind_rows(
  # DOX_RM_DEGs %>% mutate(Drug = "DOX_RM"),
  DNR_RM_DEGs %>% mutate(Drug = "DAUN_RM"),
  EPI_RM_DEGs %>% mutate(Drug = "EPI_RM"),
  MTX_RM_DEGs %>% mutate(Drug = "MITO_RM"),
)


#These gene lists were obtained from: 
#Topoisomerase II inhibitors CX-5461 and Doxorubicin differ in their cardiotoxicity profiles
#Sayan Paul, José A. Gutiérrez, Alyssa R. Bogar, Michelle C. Ward
#bioRxiv 2025.07.20.665626; doi: https://doi.org/10.1101/2025.07.20.665626


#CX 0.5 24hr
CX_SP_24 <- read.csv("data/Fig4/Toptable_CX_0.5_24.csv")

dim(CX_SP_24)
#14279 genes, make sure to now filter this for significance

#DOX 0.5 24hr
DOX_SP_24 <- read.csv("data/Fig4/Toptable_DOX_0.5_24.csv")
dim(DOX_SP_24)
#14279 genes, filter for significance

#filter both above for adj.P.Val < 0.05

DEGs_SP_sig <- list(
  "CX_SP" = CX_SP_24$Entrez_ID[CX_SP_24$adj.P.Val < 0.05],
  "DOX_SP" = DOX_SP_24$Entrez_ID[DOX_SP_24$adj.P.Val < 0.05]
)

#now make dataframes and place them with RM's data to complete my Top2i comparative dataframe

CX_SP_DEGs <- CX_SP_24 %>%
  filter(adj.P.Val < 0.05) %>%
  as.data.frame() 

dim(CX_SP_DEGs)
#278 genes

DOX_SP_DEGs <- DOX_SP_24 %>%
  filter(adj.P.Val < 0.05) %>%
  as.data.frame()

dim(DOX_SP_DEGs)
#9828 genes

#since SP's data has Entrez_ID as an integer rather than a character, coerce these to character in order for bind_rows to work properly

# CX_SP_DEGs$Entrez_ID <- as.character(CX_SP_DEGs$Entrez_ID)
# DOX_SP_DEGs$Entrez_ID <- as.character(DOX_SP_DEGs$Entrez_ID)

#these two also don't have a SYMBOL column so let's add that in

#CX df
# CX_SP_DEGs <- CX_SP_DEGs %>%
#   mutate(
#     SYMBOL = mapIds(
#       org.Hs.eg.db,
#       keys = Entrez_ID,
#       column = "SYMBOL",
#       keytype = "ENTREZID",
#       multiVals = "first"
#     )
#   ) %>%
#   dplyr::select(Entrez_ID, SYMBOL, everything())  # move SYMBOL to second column

# Same for df6
# DOX_SP_DEGs <- DOX_SP_DEGs %>%
#   mutate(
#     SYMBOL = mapIds(
#       org.Hs.eg.db,
#       keys = Entrez_ID,
#       column = "SYMBOL",
#       keytype = "ENTREZID",
#       multiVals = "first"
#     )
#   ) %>%
#   dplyr::select(Entrez_ID, SYMBOL, everything())


# DEGs_Top2i_24_df <- dplyr::bind_rows(
#   DOX_RM_DEGs %>% mutate(Treatment = "DOX_RM"),
#   DNR_RM_DEGs %>% mutate(Treatment = "DAUN_RM"),
#   EPI_RM_DEGs %>% mutate(Treatment = "EPI_RM"),
#   MTX_RM_DEGs %>% mutate(Treatment = "MITO_RM"),
#   CX_SP_DEGs %>% mutate(Treatment = "CX_SP"),
#   DOX_SP_DEGs %>% mutate(Treatment = "DOX_SP")
# ) 
  
#now this table has all columns consistently named and in the same formats

# saveRDS(DEGs_Top2i_24_df, "data/Fig4/DEGlist_Top2i_SP_RM.RDS")

DEGs_Top2i_24_df <- readRDS("data/Fig4/DEGlist_Top2i_SP_RM.RDS")

#also make a list version just in case

# DEGs_Top2i_24_list <- list(
#   "DOX_RM" = DOX_RM_DEGs, #6645 DEGs
#   "DNR_RM" = DNR_RM_DEGs, #7017 DEGs
#   "EPI_RM" = EPI_RM_DEGs, #6328 DEGs
#   "MTX_RM" = MTX_RM_DEGs, #1115 DEGs
#   "CX_SP" = CX_SP_DEGs, #278 DEGs
#   "DOX_SP" = DOX_SP_DEGs #9828 DEGs
# )

#total 10,106 DEGs from SP
#total 21,105 DEGs from RM

# saveRDS(DEGs_Top2i_24_list, "data/Fig4/DEGs_Top2i_RM_SP_list.RDS")

DEGs_Top2i_24_list <- readRDS("data/Fig4/DEGs_Top2i_RM_SP_list.RDS")

```


# Figure 4A - Percentage of TOP2i DEGs in tx+0 DEGs

Proportion barplots with permutation testing to ascertain whether DEGs shared between DOX and other TOP2i treatments are more likely than chance to be in both groups.

```{r Fig4 Barplots Permutation Testing, message=FALSE, warning=FALSE}

#read in my DEGs
DOX_24T_R <- read.csv("data/DE/Toptable_RUV_24T_final.csv") %>% dplyr::select(-X)
DOX_24R_R <- read.csv("data/DE/Toptable_RUV_24R_final.csv") %>% dplyr::select(-X)
DOX_144R_R <- read.csv("data/DE/Toptable_RUV_144R_final.csv") %>% dplyr::select(-X)

deg_list <- list(
  "t0"   = DOX_24T_R$Entrez_ID[DOX_24T_R$adj.P.Val < 0.05],
  "t24"   = DOX_24R_R$Entrez_ID[DOX_24R_R$adj.P.Val < 0.05],
  "t144"  = DOX_144R_R$Entrez_ID[DOX_144R_R$adj.P.Val < 0.05]
)

#reference set - t0
t0_set <- unique(na.omit(as.character(deg_list$"t0")))
cat("Total t0 DEGs:", length(t0_set), "\n")

#background set - all genes tested
all_genes <- readRDS("data/DE/all_genes.RDS")
cat("Total background genes:", length(all_genes), "\n")


#define TOP2i treatment DEG sets from literature
comparison_DEGs <- list(
  DOX_RM = as.character(DOX_RM_DEGs$Entrez_ID),
  DNR_RM = as.character(DNR_RM_DEGs$Entrez_ID),
  EPI_RM = as.character(EPI_RM_DEGs$Entrez_ID),
  MTX_RM = as.character(MTX_RM_DEGs$Entrez_ID),
  CX_SP  = as.character(CX_SP_DEGs$Entrez_ID),
  DOX_SP = as.character(DOX_SP_DEGs$Entrez_ID)
)

#build data for a stacked barplot with percentages

plot_df <- purrr::map_dfr(names(comparison_DEGs), function(name) {
  comp_DEGs <- unique(comparison_DEGs[[name]])
  
  overlap <- sum(t0_set %in% comp_DEGs)
  non_overlap <- length(t0_set) - overlap
  
  tibble(
    Treatment = name,
    Category = c("DEG", "nonDEG"),
    Count = c(overlap, non_overlap),
    Total = length(t0_set),
    Fraction = c(overlap, non_overlap)/length(t0_set)
  )
})

plot_df <- plot_df %>%
  mutate(
    Treatment = factor(Treatment, levels = names(comparison_DEGs)),
    Category = factor(Category, levels = c("DEG", "nonDEG"))
  )

#legend labels (with n counts)
legend_labels <- plot_df %>%
  filter(Category == "DEG Status") %>%
  mutate(Label = paste0(Category, " (n=", Count, ")")) %>%
  dplyr::select(Treatment, Label) %>%
  deframe()

#perform permutation testing for enrichment
perm_test_overlap <- function(gene_set, t0_genes, gene_universe, n_perm = 10000) {
  observed <- length(intersect(t0_genes, gene_set))
  
  perm_overlaps <- replicate(n_perm, {
    random_genes <- sample(gene_universe, length(t0_genes))
    length(intersect(random_genes, gene_set))
  })
  
  p_val <- mean(perm_overlaps >= observed)
  enrichment <- observed / mean(perm_overlaps)
  
  return(list(observed = observed,
              mean_perm = mean(perm_overlaps),
              enrichment = enrichment,
              p_value = p_val))
}

perm_results <- lapply(comparison_DEGs, 
                       perm_test_overlap,
                       t0_genes = t0_set,
                       gene_universe = all_genes,
                       n_perm = 10000)

#summary stats table
perm_summary_df <- map_dfr(names(perm_results), function(name) {
  res <- perm_results[[name]]
  tibble(
    Treatment = name,
    Observed = res$observed,
    MeanPerm = res$mean_perm,
    Enrichment = res$enrichment,
    PValue = res$p_value
  )
}) %>%
  mutate(FDR = p.adjust(PValue, method = "BH"),
         Significant = case_when(
           FDR < 0.001 ~ "***",
           FDR < 0.01  ~ "**",
           FDR < 0.05  ~ "*",
           TRUE ~ "ns"
         ))

#show summary stats table
print(perm_summary_df)

#label positions
#inside overlap bars: percentages
overlap_labels <- plot_df %>%
  filter(Category == "Overlap") %>%
  left_join(perm_summary_df %>% dplyr::select(Treatment, Significant), by = "Treatment") %>%
  mutate(
    Label = paste0(round(Fraction*100, 1), ""),
    y_pos = 0.5   #fixed position at 50%
  )

#above bars: significance stars
label_positions <- plot_df %>%
  group_by(Treatment) %>%
  summarise(y_pos = 1.0, .groups = "drop") %>%
  left_join(perm_summary_df %>% dplyr::select(Treatment, Significant), by = "Treatment")


#plot final barplot
TOP2i_Combined <- ggplot(plot_df, aes(x = Treatment, y = Fraction, fill = Category)) +
  geom_bar(stat = "identity", position = "stack", width = 0.6) +
  
  #percentages inside overlap bars (fixed position)
  geom_text(
    data = overlap_labels,
    aes(x = Treatment, y = y_pos, label = Label),
    inherit.aes = FALSE,
    color = "white",
    fontface = "bold",
    size = 4
  ) +
  
  #significance stars above bars
  geom_text(
    data = label_positions,
    aes(x = Treatment, y = y_pos, label = Significant),
    inherit.aes = FALSE,
    size = 4,
    fontface = "bold",
    color = "black"
  ) +
  
  scale_y_continuous(labels = c(0, 100)) +
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8"),
                    labels = legend_labels) +
  labs(
    title = "Proportion of TOP2i DEGs in tx+0 DEGs",
    x = "TOP2i Treatment",
    y = "Percentage of DOX tx+0 DEGs (%)",
    fill = "TOP2i DEGs"
  ) +
  theme_custom()

# save_plot(
#   plot = TOP2i_Combined,
#   folder = output_folder,
#   filename = "TOP2i_Barplot_Combined_Fig4_EMP",
#   height = 6,
#   width = 8
# )

```


# Supplementary Figure 7 - TOP2i DEGs Acute vs Chronic Clusters

```{r Sup Fig 7 Barplots, message=FALSE, warning=FALSE}

#read in my chronic and acute cluster gene lists
final_genes_1_RUV <- readRDS("data/Fig3/final_genes_1_RUV.RDS")
final_genes_2_RUV <- readRDS("data/Fig3/final_genes_2_RUV.RDS")

acute_set <- unique(na.omit(final_genes_1_RUV))
chronic_set <- unique(na.omit(final_genes_2_RUV))

all_genes

comparison_DEGs <- list(
  DOX_RM = as.character(DOX_RM_DEGs$Entrez_ID),
  DNR_RM = as.character(DNR_RM_DEGs$Entrez_ID),
  EPI_RM = as.character(EPI_RM_DEGs$Entrez_ID),
  MTX_RM = as.character(MTX_RM_DEGs$Entrez_ID),
  CX_SP  = as.character(CX_SP_DEGs$Entrez_ID),
  DOX_SP = as.character(DOX_SP_DEGs$Entrez_ID)
)

#fisher's exact test per treatment
fisher_test_overlap <- function(acute, chronic, degs) {
  a <- sum(acute %in% degs)
  b <- length(acute) - a
  c <- sum(chronic %in% degs)
  d <- length(chronic) - c
  
  contingency <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  fisher_res <- fisher.test(contingency)
  
  list(
    contingency = contingency,
    p_value = fisher_res$p.value,
    significant = case_when(
      fisher_res$p.value < 0.001 ~ "***",
      fisher_res$p.value < 0.01  ~ "**",
      fisher_res$p.value < 0.05  ~ "*",
      TRUE ~ ""
    )
  )
}

# Prepare summary table for plotting
plot_summary <- map_dfr(names(comparison_DEGs), function(treat) {
  degs <- comparison_DEGs[[treat]]
  
  # Fisher test
  fisher_res <- fisher_test_overlap(acute_set, chronic_set, degs)
  
  # Counts for each motif
  acute_overlap <- sum(acute_set %in% degs)
  acute_nonoverlap <- length(acute_set) - acute_overlap
  chronic_overlap <- sum(chronic_set %in% degs)
  chronic_nonoverlap <- length(chronic_set) - chronic_overlap
  
  tibble(
    Treatment = treat,
    Motif = rep(c("Acute", "Chronic"), each = 2),
    Category = rep(c("DEG", "nonDEG"), 2),
    Count = c(acute_overlap, acute_nonoverlap, chronic_overlap, chronic_nonoverlap),
    Fraction = c(acute_overlap, acute_nonoverlap, chronic_overlap, chronic_nonoverlap) /
               c(length(acute_set), length(acute_set), length(chronic_set), length(chronic_set)),
    Significant = rep(c(fisher_res$significant, ""), 2)  #only on DEG bars
  )
})

#ensure correct order
plot_summary$Category <- factor(plot_summary$Category, levels = c("DEG", "nonDEG"))

#position stars at y = 1
label_positions <- plot_summary %>%
  filter(Category == "DEG") %>%
  mutate(y_pos = 1)

#plot
top2i_barplot_motifs_supp7 <- ggplot(plot_summary, aes(x = Treatment, y = Fraction, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = label_positions %>% filter(Significant != ""),
            aes(x = Treatment, y = y_pos, label = Significant),
            inherit.aes = FALSE, size = 6, fontface = "bold", color = "black") +
  geom_text(aes(label = paste0(round(Fraction*100,1), "")),
            position = position_stack(vjust = 0.5), size = 5, color = "white") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_fill_manual(values = c("DEG" = "blue", "nonDEG" = "red")) +
  facet_wrap(~Motif) +
  labs(
    title = "Percentage of TOP2i DEGs (Matthews et al., Paul et al.) in Acute and Chronic response clusters",
    x = "TOP2i Treatment",
    y = "Overlap Percentage (%)",
    fill = ""
  ) +
  theme_custom()

print(top2i_barplot_motifs_supp7)

# save_plot(
#   plot = top2i_barplot_motifs_supp7,
#   filename = "Supp7_TOP2i_Barplots_EMP",
#   folder = output_folder,
#   height = 5,
#   width = 8.5
# )

#print percentages and p-values

for(treat in names(comparison_DEGs)) {
  degs <- comparison_DEGs[[treat]]
  acute_overlap <- sum(acute_set %in% degs)
  chronic_overlap <- sum(chronic_set %in% degs)
  prop_ac <- acute_overlap / length(acute_set)
  prop_ch <- chronic_overlap / length(chronic_set)
  fisher_res <- fisher_test_overlap(acute_set, chronic_set, degs)
  
  cat(sprintf("%s - Acute: %.2f%% (%d/%d), Chronic: %.2f%% (%d/%d), Fisher p = %.5f\n",
              treat, 100*prop_ac, acute_overlap, length(acute_set),
              100*prop_ch, chronic_overlap, length(chronic_set),
              fisher_res$p_value))
}

#significance stars are for likelihood of being in the Chronic cluster over the Acute cluster - stars are the same on both sides but only really count for Chronic

```

# Figure 4B - Enrichment of TOP2i DEGs in Chronic cluster relative to Acute

```{r Fig 4b, message=FALSE, warning=FALSE}

#use the workflow above to now plot forest plots of odds ratios for these comparisons
#I am comparing sets of DEGs from other TOP2i against my motifs here
#I want to use my fisher's exact test data

#same inputs as before to preserve them
# final_genes_1_RUV
# final_genes_2_RUV

# acute_set 
# chronic_set

#remove DOX and preserve order DNR EPI MTX CX
comparison_DEGs_noDOX <- list(
  DNR_RM = as.character(DNR_RM_DEGs$Entrez_ID),
  EPI_RM = as.character(EPI_RM_DEGs$Entrez_ID),
  MTX_RM = as.character(MTX_RM_DEGs$Entrez_ID),
  CX_SP  = as.character(CX_SP_DEGs$Entrez_ID)
)

treatment_order <- c("DNR_RM", "EPI_RM", "MTX_RM", "CX_SP")

fisher_test_overlap <- function(acute, chronic, degs) {
  a <- sum(acute %in% degs)
  b <- length(acute) - a
  c <- sum(chronic %in% degs)
  d <- length(chronic) - c
  
  contingency <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  fisher_res <- fisher.test(contingency)
  
  tibble(
    OR = unname(fisher_res$estimate),
    CI_low = fisher_res$conf.int[1],
    CI_high = fisher_res$conf.int[2],
    p_value = fisher_res$p.value,
    Significant = case_when(
      fisher_res$p.value < 0.001 ~ "***",
      fisher_res$p.value < 0.01 ~ "**",
      fisher_res$p.value < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )
}


#build a summary table for the results of fishers exact for the forest plots
forest_summary <- map_dfr(names(comparison_DEGs_noDOX), function(treat) {
  res <- fisher_test_overlap(chronic_set, acute_set, comparison_DEGs_noDOX[[treat]])
  res %>% mutate(Treatment = treat)
}) %>% 
  mutate(Treatment = factor(Treatment, levels = rev(treatment_order)), #define order
         Group = ifelse(grepl("_RM$", Treatment), "RM", "SP"))

#create my combined plot
forest_TOP2i <- ggplot(forest_summary, 
                       aes(y = Treatment, x = OR, 
                           xmin = CI_low, xmax = CI_high)) + 
  geom_point(size = 4) +
  geom_errorbarh(height = 0.2, linewidth = 0.5, color = "black") +
  geom_text(aes(label = Significant), position = position_nudge(y = 0.1), 
            color = "black", fontface = "bold", size = 5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  scale_x_log10(limits = c(0.1, 1000),
                breaks = c(0.1, 1, 10, 100, 1000)) +
  scale_color_manual(values = c("_" = "grey", "_" = "grey")) + #remove the groups this way even though it's not the best method
  labs(
    title = "Enrichment of TOP2i DEGs in Chronic response cluster relative to Acute response cluster",
    x = "log odds ratio (95% confidence interval)",
    y = ""
  ) +
  theme_custom()

print(forest_TOP2i)

#make a summary table

forest_summary_TOP2i <- forest_summary

View(forest_summary_TOP2i)



```


# Comparing response to DOX across cell types

We obtained three gene sets for this comparison from various proliferating cell types to compare with our non-proliferating cells. 

Set 1: Distefano, R., Ilieva, M., Madsen, J. H., Rennie, S., & Uchida, S. (2023). DoxoDB: A Database for the Expression Analysis of Doxorubicin-Induced lncRNA Genes. Non-Coding RNA, 9(4), 39. https://doi.org/10.3390/ncrna9040039 

Set 2: DePianto DJ, Heiden JAV, Morshead KB, Sun KH, Modrusan Z, Teng G, Wolters PJ, Arron JR. Molecular mapping of interstitial lung disease reveals a phenotypically distinct senescent basal epithelial cell population. JCI Insight. 2021 Apr 22;6(8):e143626. doi: 10.1172/jci.insight.143626. PMID: 33705361; PMCID: PMC8119199.


```{r cell type comparisons as in 4B, warning=FALSE, message=FALSE}

filcpm_mat_genes <- readRDS("data/QC/filcpm_matrix_genes.RDS")
all_symb_genes <- filcpm_mat_genes$SYMBOL

distefano_s8_ids <- read_excel("C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/Gene Tables/TableS8_Distefano_2023_SenescenceGenes.xlsx", 
    range = "A3:F2910")

#check how many of these genes (2907) are expressed in our dataset

cat("distefano_s8_ids:", sum(distefano_s8_ids$Gene_symbol %in% all_symb_genes), "of", length(distefano_s8_ids$Gene_symbol), "\n")
#2,053 of 2,907 are expressed in my dataset

depianto_s1_ids <- c(
  "ABCA1", "CLDN4", "FANK1", "IL1RN", "MYLK", "REEP", "SP6", "TRAF3IP3",
  "ACKR3", "COL5A1", "FAT2", "IL32", "NAGK", "RHBDL2", "SPRR1A", "TREM2",
  "ACSF2", "CPA4", "FBN1", "IL33", "NCCRP1", "RHCG", "SPRR1B", "TRIM16",
  "ADAMTS7", "CPPED1", "FER1L4", "IL36G", "NDST1", "RHOV", "SPRR2A", "TUBA1A",
  "ADAMTSL4", "CRABP2", "FGF11", "IRF6", "NRCAM", "RNF212B", "SRPX2", "TUBB2A",
  "AIM1L", "CRYAB", "FHDC1", "IVL", "NTN1", "RRAD", "SSC4D", "UCP2",
  "ALOX15B", "CST6", "FLG", "KIAA0922", "NYNRIN", "S100A4", "STARD5", "ULBP2",
  "ANGPTL4", "CTNNBIP1", "FLG-AS1", "KLRC4", "OTUB2", "S100A8", "STC1", "UPK3B",
  "ANKRD22", "CUX2", "FRMD8", "KLRK1", "OVOL1", "S100A9", "SUGCT", "VGLL3",
  "ANXA4", "CXCL1", "GALNT5", "KRT13", "PAM", "SAA1", "SYNPO", "WDR63",
  "ARHGEF4", "CXCL17", "GAS6-AS1", "KRT16", "PAMR1", "SCD5", "SYTL4", "XG",
  "ATL1", "CYB5R1", "GBP3", "KRT23", "PAPLN", "SCN4B", "TACSTD2", "ZNF185",
  "B4GALNT4", "DAPK1", "GCNT4", "KRT6A", "PAQR7", "SEC14L2", "TAGLN", "ZNF429",
  "C1orf74", "DHRS1", "GGT1", "KRT6B", "PCBP4", "SEPT5-GP1BB", "TCEA2", "ZNF488",
  "C1QTNF1", "DHRS12", "GJB4", "LAYN", "PDE9A", "SERPINB13", "TENM2", "ZNF702P",
  "C1QTNF6", "DIXDC1", "GPX3", "LDLRAD2", "PDLIM1", "SERPINB3", "THSD4", "ZNF704",
  "C3", "DKK3", "GRHL3", "LOC100862671", "PGBD5", "SERPINB7", "TLL2", "ZNF750",
  "CALML3", "DOCK8", "HEG1", "LOC101928100", "PIK3IP1", "SERPINB9", "TMBIM1",
  "CALML3-AS1", "DPYSL4", "HEPHL1", "LOC101928281", "PLA2G4F", "SH3BGRL",
  "TMEM132A", "CAPN3", "DTNA", "HES2", "LOC101930241", "PLK2", "SH3TC2",
  "TMEM217", "CCND1", "DUOXA1", "HIST1H1C", "LOC103021296", "PLOD2", "SLC15A3",
  "TMEM229B", "CCND2", "ECM1", "HIST1H2AC", "LRP1", "PORCN", "SLC16A2", "TMEM40",
  "CCRN4L", "EDN1", "HIST1H2BD", "LY6D", "PPP2R2C", "SLC1A1", "TMEM63C", "CD24",
  "EHD3", "HIST1H2BK", "LYPD3", "PROM2", "SLC31A2", "TMEM79",
  "CD82", "ELFN2", "HIST2H2BE", "MAGED1", "PRR5L", "SLC39A11", "TMPRSS13",
  "CEACAM6", "ENC1", "HSPG2", "MATN2", "PRSS22", "SLC44A2", "TMPRSS4",
  "CHST2", "ENTPD3", "ID3", "MFI2", "PRSS23", "SLC46A3", "TNFSF10",
  "CKB", "EPHA4", "IGFBP2", "MMP28", "PTGES", "SLC6A11", "TOM1L2",
  "CLCA2", "EPHB3", "IKZF2", "MXRA5", "RAB7B", "SLC9A3", "TP53I3",
  "CLDN1", "ESPN", "IL12RB2", "MYH14", "REC", "SNCG", "TPM1"
)

depianto_s2_ids <- depianto_s2_ids <- c(
  "ABCA3", "CYB5D2", "LYNX1", "RRM2B",
  "ABCA7", "DHRS3", "MANSC1", "SCG50",
  "ABCA8", "DPP4", "MGARP", "SECTM1",
  "AC007388.1", "DRAM1", "MGP", "SERPING1",
  "ACSS3", "DYNC2LI1", "MGST2", "SLC1A1",
  "ADAMTS5", "FAM162A", "MORN4", "SLC40A1",
  "ADGRL1", "FAM198B", "MXI1", "SOD2",
  "AMPD3", "FAM43A", "NALCN", "SPATA18",
  "AMZ2P1", "FAXDC2", "NCOA7", "ST3GAL5",
  "ANK2", "FBXO32", "NEAT1", "SULF1",
  "AP001972.5", "FER1L4", "NTN4", "SULF2",
  "ARHGEF37", "FIBIN", "ORAI3", "SVEP1",
  "ARRB1", "GFRA1", "PAPPA", "TCEA2",
  "BAMBI", "GPR155", "PCMTD1", "TLCD2",
  "BMP4", "GPRC5B", "PCSK5", "TM7SF2",
  "BTG2", "HIST1H1C", "PDGFD", "TMEM170B",
  "C1QTNF1", "HIST1H2AC", "PIK3IP1", "TMEM176B",
  "C1S", "HIST1H2BD", "PLD1", "TMEM178B",
  "C3", "HIST1H3E", "PLXNC1", "TP53INP1",
  "CAMK1D", "HIST1H4H", "PODN", "TXNIP",
  "CBLN3", "HIST1H4I", "PPP1R3C", "VSIR",
  "CCND2", "HNMT", "PRRG1", "WDR63",
  "CCPG1", "IFI6", "PTCHD4", "WFDC21P",
  "COL4A5", "IGDCC4", "QPRT", "YPEL2",
  "CPE", "JAM2", "RARRES2", "ZMAT3",
  "CPZ", "JMY", "RDH10", "ZNF219",
  "CREBRF", "KCNJ2", "RGCC", "ZNF667",
  "CRELD1", "KLHL24", "RHBDL1",
  "CTSF", "LACC1", "RILP",
  "CTSO", "LRRK2", "RPS6KA2"
)

#check overlap with my all_genes excluding lowly-expressed genes
cat("depianto_s1_ids:", sum(depianto_s1_ids %in% all_symb_genes), "of", length(depianto_s1_ids), "\n")

cat("depianto_s2_ids:", sum(depianto_s2_ids %in% all_symb_genes), "of", length(depianto_s2_ids), "\n")


#now translate these genes into Entrez_ID so I can make accurate comparisons
to_entrez <- function(symbols, annotation_df) {
  annotation_df %>%
    dplyr::filter(SYMBOL %in% symbols) %>%
    dplyr::pull(Entrez_ID) %>%
    as.character() %>%
    unique() %>%
    na.omit()
}

PDL1pos_HCA2_entrez <- to_entrez(distefano_s8_ids$Gene_symbol, filcpm_mat_genes)
Lung_Epi_entrez   <- to_entrez(depianto_s1_ids, filcpm_mat_genes)
Fibro_entrez      <- to_entrez(depianto_s2_ids, filcpm_mat_genes)

cat("PDL1+_HCA2:", length(PDL1pos_HCA2_entrez), "of", length(unique(distefano_s8_ids$Gene_symbol)), "\n")
cat("Lung_Epi:",   length(Lung_Epi_entrez),   "of", length(depianto_s1_ids), "\n")
cat("Fibro:",      length(Fibro_entrez),       "of", length(depianto_s2_ids), "\n")

```

# Test for enrichment

```{r enrichment between chronic and acute revisions, warning=FALSE, message=FALSE}

comp_celltype_genes <- list(
  PDL1pos_HCA2 = PDL1pos_HCA2_entrez,
  Lung_Epi   = Lung_Epi_entrez,
  Fibro      = Fibro_entrez
)

celltype_order <- c("PDL1pos_HCA2", "Lung_Epi", "Fibro")

#Fisher's exact test function
fisher_test_overlap <- function(acute, chronic, degs) {
  a <- sum(acute %in% degs)
  b <- length(acute) - a
  c <- sum(chronic %in% degs)
  d <- length(chronic) - c
  
  contingency <- matrix(c(a, b, c, d), nrow = 2, byrow = TRUE)
  fisher_res <- fisher.test(contingency)
  
  tibble(
    OR = unname(fisher_res$estimate),
    CI_low = fisher_res$conf.int[1],
    CI_high = fisher_res$conf.int[2],
    p_value = fisher_res$p.value,
    Significant = case_when(
      fisher_res$p.value < 0.001 ~ "***",
      fisher_res$p.value < 0.01  ~ "**",
      fisher_res$p.value < 0.05  ~ "*",
      TRUE ~ "ns"
    )
  )
}

#build forest summary
forest_summary_celltype <- map_dfr(names(comp_celltype_genes), function(treat) {
  res <- fisher_test_overlap(chronic_set, acute_set, comp_celltype_genes[[treat]])
  res %>% mutate(Gene_Set = treat)
}) %>%
  mutate(Gene_Set = factor(Gene_Set, levels = rev(celltype_order)))

#create a forest plot
forest_celltype <- ggplot(forest_summary_celltype,
                          aes(y = Gene_Set, x = OR,
                              xmin = CI_low, xmax = CI_high)) +
  geom_point(size = 4) +
  geom_errorbarh(height = 0.2, linewidth = 0.5, color = "black") +
  geom_text(aes(label = Significant), position = position_nudge(y = 0.3),
            color = "black", fontface = "bold", size = 5) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  scale_x_log10(limits = c(0.1, 1000),
                breaks = c(0.1, 1, 10, 100, 1000)) +
  labs(
    title = "Enrichment of Cell-type Senescence Genes in Chronic vs Acute",
    x = "log odds ratio (95% confidence interval)",
    y = ""
  ) +
  theme_custom()

print(forest_celltype)

# Save summary table
forest_summary_celltype_table <- forest_summary_celltype
View(forest_summary_celltype_table)


```

