---
title: "test"
author: "Emma M Pfortmiller"
date: "2026-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# DOX Proteomics Comparisons

For this comparison, we utilized Supplementary Table 2 from Johnson, O. D., Paul, S., Gutiérrez, J. A., Russell, W. K., & Ward, M. C. (2025). DNA-damage-associated protein co-expression network in cardiomyocytes informs on tolerance to genetic variation and disease. iScience, 28(5). https://doi.org/10.1016/j.isci.2025.112474 

I have taken the differentially abundant proteins (DAPs) identified after DOX treatment for 24 hours, as defined by an adj. p value < 0.05

```{r read in DAPs and nonDAPs from ODJ SP proteomics paper, warning=FALSE, message=FALSE}
#I have the excel sheet downloaded from the paper mentioned above, which I will read in here and then filter the proteins for DAPs vs nonDAPs

doxprot_supp2 <- read_excel("C:/Users/emmap/OneDrive/Desktop/DXR Manuscript Materials/proteomics_supptable2_ODJ_SP_MCW_downloaded_260408.xlsx", 
    range = "B2:AA4180")

#now go ahead and make two lists, one with DAPs and one with nonDAPs
dox_daps <- doxprot_supp2 %>% 
  filter(P.Value.adj < 0.05) %>% 
  pull(Protein)
#according to the paper, should be 319 DAPs
length(dox_daps)

dox_nondaps <- doxprot_supp2 %>%
  filter(P.Value.adj >= 0.05) %>% 
  pull(Protein)
#according to the paper, should be 3,859 DAPs
length(dox_nondaps)

#now combine these into one list
doxprot_list <- list(
  DAPs = dox_daps,
  nonDAPs = dox_nondaps
)

#now I'll have to convert these protein names (UniProt IDs) so I can compare with my Entrez_ID genes
#I will use bitr from the clusterProfiler package as a first attempt to translate between ids

dox_daps_entrez <- 

```










