---
title: "test"
author: "Emma M Pfortmiller"
date: "2026-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# IF DAPI Quant - Macros for FIJI/ImageJ

```{r imagej quant macros, warning=FALSE, message=FALSE}

# // DAPI Intensity Only — using existing ROI zips from above
# folder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 75-1\\rgb quant\\";
# roiFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 75-1\\outline new\\";
# saveDAPIFolder = "D:\\DXR Project\\IF yH2AX\\if yh2ax dxr 75-1\\DAPI intensity\\";
# 
# list = getFileList(folder);
# 
# for (i = 0; i < list.length; i++) {
#     if (endsWith(list[i], ".tif")) {
#         print("----------------------------------------------------");
#         print("Processing file: " + list[i]);
# 
#         baseName = replace(list[i], ".tif", "");
# 
#         // === Check ROI zip exists ===
#         roiPath = roiFolder + baseName + "_DAPI.zip";
#         if (!File.exists(roiPath)) {
#             print("WARNING: No ROI zip found for " + baseName + " — skipping.");
#             continue;
#         }
# 
#         // === Open composite and extract DAPI (blue) channel ===
#         open(folder + list[i]);
#         originalTitle = getTitle();
# 
#         run("Duplicate...", "title=" + baseName + "_proc");
#         selectImage(originalTitle);
#         close();
# 
#         selectWindow(baseName + "_proc");
#         run("Split Channels");
# 
#         // Close red and green, keep blue
#         redTitle = baseName + "_proc (red)";
#         greenTitle = baseName + "_proc (green)";
#         blueTitle = baseName + "_proc (blue)";
# 
#         if (isOpen(redTitle)) { selectWindow(redTitle); close(); }
#         if (isOpen(greenTitle)) { selectWindow(greenTitle); close(); }
# 
#         if (!isOpen(blueTitle)) {
#             print("WARNING: Blue channel not found for " + baseName + " — skipping.");
#             continue;
#         }
# 
#         // === Load ROIs and measure DAPI intensity ===
#         selectWindow(blueTitle);
#         roiManager("Reset");
#         roiManager("Open", roiPath);
#         print("Loaded " + roiManager("count") + " ROIs for " + baseName);
# 
#         run("Set Measurements...", "mean min max integrated area redirect=None decimal=3");
#         roiManager("Deselect");
#         roiManager("Measure");
# 
#         saveAs("Results", saveDAPIFolder + baseName + "_DAPI_intensity.csv");
#         print("Saved: " + baseName + "_DAPI_intensity.csv");
#         run("Clear Results");
# 
#         // === Cleanup ===
#         selectWindow(blueTitle);
#         close();
#         roiManager("Deselect");
#         roiManager("Reset");
# 
#         print("Finished: " + baseName);
#     }
# }

```

# IF DAPI Quant - Step 1

## Read in DAPI csvs from ImageJ

```{r dapi quantification step 1, warning=FALSE, message=FALSE}
#for the first step, I want to collect all of the information from the macros I ran to get mean intensity from the same images used above for the yH2AX quantification

#create a function to read all per-image DAPI CSVs for each cell line and input magnification
# read_dapi_intensity <- function(cell_line, individual_label, default_magnification = NULL) {
#   dapi_folder <- paste0("D:/DXR Project/IF yH2AX/if yh2ax dxr ", cell_line, "/DAPI intensity/")
#   files <- list.files(dapi_folder, pattern = "_DAPI_intensity\\.csv$", full.names = TRUE)
#   
#   bind_rows(lapply(files, function(f) {
#     read.csv(f) %>%
#       mutate(
#         ROI_File = gsub("_DAPI_intensity\\.csv$", "", basename(f)),
#         Individual = individual_label,
#         Line = cell_line,
#         Magnification = if (!is.null(default_magnification)) {
#           default_magnification
#         } else {
#           case_when(
#             grepl("20X", ROI_File, ignore.case = TRUE) ~ "20X",
#             grepl("10X", ROI_File, ignore.case = TRUE) ~ "10X",
#             grepl("40X", ROI_File, ignore.case = TRUE) ~ "40X",
#             TRUE ~ "Other"
#           )
#         }
#       )
#   }))
# }

# ind1_dapi <- read_dapi_intensity("87-1", "Ind1") %>% dplyr::select(-X)
# ind3_dapi <- read_dapi_intensity("75-1", "Ind3") %>% dplyr::select(-X)
# ind4_dapi <- read_dapi_intensity("84-1", "Ind4", default_magnification = "20X") %>% dplyr::select(-X)

# saveRDS(ind1_dapi, "data/Fig1/ind1_dapi_rois.RDS")
# saveRDS(ind3_dapi, "data/Fig1/ind3_dapi_rois.RDS")
# saveRDS(ind4_dapi, "data/Fig1/ind4_dapi_rois.RDS")

#read in the finalized files
ind1_dapi <- readRDS("data/Fig1/ind1_dapi_rois.RDS")
ind3_dapi <- readRDS("data/Fig1/ind3_dapi_rois.RDS")
ind4_dapi <- readRDS("data/Fig1/ind4_dapi_rois.RDS")

```

## Create DAPI Dataframe

```{r Create DAPI Summary matching existing pipeline, message=FALSE, warning=FALSE}

#combine and apply identical filtering/annotation as above for yH2AX
# ind_dapi_all <- bind_rows(ind1_dapi, ind3_dapi, ind4_dapi) %>%
#   mutate(
#     Treatment = case_when(
#       grepl("DOX", ROI_File) ~ "DOX",
#       grepl("DMSO", ROI_File) ~ "VEH",
#       TRUE ~ "Other"
#     ),
#     Timepoint = case_when(
#       grepl("24T", ROI_File) ~ "24T",
#       grepl("24R", ROI_File) ~ "24R",
#       grepl("144R", ROI_File) ~ "144R",
#       TRUE ~ "Other"
#     ),
#     Condition = paste0(Treatment, "_", Timepoint)
#   ) %>%
#   dplyr::filter(Treatment != "Other") %>%
#   dplyr::filter(Magnification == "20X") %>%
#   dplyr::filter(!grepl("^2nd", ROI_File, ignore.case = TRUE))

# saveRDS(ind_dapi_all, "data/Fig1/IF_quant/individual_dapi_all.RDS")
ind_dapi_all <- readRDS("data/Fig1/IF_quant/individual_dapi_all.RDS")

```

## DAPI Intensity Plot

```{r DAPI Intensity - ROI Level, message=FALSE, warning=FALSE}

#plot all of these intensity measurements for Hoechst (DAPI) nuclear staining

plot_dapi_1 <- ind_dapi_all %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  ggplot(aes(x = Condition, y = Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 1.5,
             alpha = 1,
             position = position_identity()) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 1: ROI Level Mean Hoechst Intensity",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_1

```

## Create an Image Level Summary

```{r collapse DAPI to image level and plot, message=FALSE, warning=FALSE}
#create an image-level summary of all of the relevant ROIs (nuclei)

# dapi_image_level_summary <- ind_dapi_all %>%
#   mutate(Image = trimws(ROI_File)) %>%
#   group_by(Image, Individual, Line, Magnification, Treatment, Timepoint, Condition) %>%
#   dplyr::select(Image, Individual, Line, Magnification, Treatment, Timepoint, Condition, Mean) %>% 
#   summarize(
#     Mean_of_Mean = mean(Mean, na.rm = TRUE),
#     .groups = "drop"
#   )

#this results in a dataframe of the 146 20X images I took total across three cell lines

# saveRDS(dapi_image_level_summary, "data/Fig1/IF_quant/dapi_image_level_summary_all.RDS")
dapi_image_level_summary <- readRDS("data/Fig1/IF_quant/dapi_image_level_summary_all.RDS")

plot_dapi_image <- dapi_image_level_summary %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  ggplot(aes(x = Condition, y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 3,
             alpha = 0.8,
             position = position_identity()) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 2: Image-Level Mean Hoechst Intensity",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_image



```

## Merge with Metadata

```{r Merge Image-Level Summary with Nuclei Counts, message=FALSE, warning=FALSE}
#now I would like to merge my image_level_metadata made above (on the same samples) with these

image_level_metadata <- readRDS("data/Fig1/IF_quant/image_level_metadata_all.RDS")

# image_merged_dapi_summary <- 
#   image_level_metadata %>% 
#   mutate(
#     Image = str_remove(Image, "_DAPI") #remove the ending that doesn't match here
#   ) %>% 
#   dplyr::select(Individual, Line, Treatment, Timepoint, Condition, Magnification,
#                 Image, Nuclei_Count) %>% 
#   dplyr::left_join(dapi_image_level_summary, by = c("Image", "Individual", "Line", "Magnification",
#                                                     "Treatment", "Timepoint", "Condition"))

# saveRDS(image_merged_dapi_summary, "data/Fig1/image_merged_dapi_summary.RDS")
image_merged_dapi_summary <- readRDS("data/Fig1/image_merged_dapi_summary.RDS")

```

## Assess the number of nuclei per image

In these experiments, Ind1 was very confluent at the time of IF staining. 

Ind4 had fewer cells and appeared sparser across all wells (despite seeding at the same density).

Ind3 was also less confluent than Ind1.

```{r Assess Images for Quality, message=FALSE, warning=FALSE}
#first look at the distribution of nuclei counts to inform a threshold
image_merged_dapi_summary %>%
  ggplot(aes(x = Nuclei_Count)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  facet_wrap(~ Individual) +
  labs(title = "Distribution of Nuclei Counts per Image",
       x = "Nuclei Count", y = "Count") +
  theme_custom()

#also look by condition to see if low counts cluster anywhere
image_merged_dapi_summary %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  ggplot(aes(x = Condition, y = Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), size = 2, alpha = 0.7) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Nuclei Count per Image by Condition",
       x = "Condition", y = "Nuclei Count") +
  theme_custom()

#look at distribution per individual
#84-1 was less confluent and Ind1 was very confluent at the time of staining
image_merged_dapi_summary %>%
  ggplot(aes(x = Nuclei_Count)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
  facet_wrap(~ Individual, scales = "free") +  # free scales important here
  labs(title = "Nuclei Count Distribution by Individual",
       x = "Nuclei Count", y = "Count") +
  theme_custom()
```



## Hoechst Quant Step 3

For this step, I will test two scenarios. For the first scenario, I will use the exact same images used in the previous yH2AX quantification (3x per condition)

In the second scenario, I will keep all of the images so as not to artificially select those images with the same numbers of nuclei across conditions, which may mask differences across conditions (if any).

```{r step 3 dapi quant 3 images, warning=FALSE, message=FALSE}
#now, I would like to use the same 3 images per condition to plot Hoechst intensity

pipeline_images <- merged_image_summary_final %>%
  distinct(Image, Individual, Line) %>%
  mutate(Image = gsub("_DAPI", "", Image)) %>%
  dplyr::select(Image, Individual, Line)

#now that I have the set of 54 images from the previous experiment, let's make my final set for dapi quantification like merged_image_summary_final above

image_merged_dapi_summary_final <- image_merged_dapi_summary %>%
  semi_join(pipeline_images, by = c("Image", "Individual", "Line"))

# saveRDS(image_merged_dapi_summary_final, "data/Fig1/image_merged_dapi_summary_final.RDS")

#now plot these in a boxplot with the three images per individual

plot_dapi_3 <- ggplot(image_merged_dapi_summary_final,
                      aes(x = Condition,
                          y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 3: Select 3 Images Per Individual With Consistent Nuclei",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_3

#now do stats for these to see if there are any differences over time
ttests_step3_dapi <- image_merged_dapi_summary_final %>% 
  group_by(Timepoint) %>% 
  t_test(Mean_of_Mean ~ Treatment) %>% 
  mutate(
    y.position = max(image_merged_dapi_summary_final$Mean_of_Mean, na.rm = TRUE) * 1
  )

sig_table_dapi_3 <- image_merged_dapi_summary_final %>% 
  group_by(Timepoint) %>% 
  summarise(
    p = t.test(Mean_of_Mean ~ Treatment)$p.value,
    y.position = max(Mean_of_Mean, na.rm = TRUE) * 1.05,
    .groups = "drop"
  ) %>% 
  mutate(
    signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint)
  )

ttests_step3_dapi

#now plot with these ttest statistics above the plot
plot_dapi_3_stats <- ggplot(image_merged_dapi_summary_final,
                            aes(x = Condition,
                                y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  stat_pvalue_manual(sig_table_dapi_3,
                     label = "signif",
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 3: Select 3 Images Per Individual with Consistent Nuclei")

```

## Step 3 Hoechst Quantification - no image selection

```{r step 3 no image selection, message=FALSE, warning=FALSE}
#now, perform these steps in the same fashion as above on all images processed for this experiment. At the individual level, these numbers of nuclei may differ due to experimental factors including cell confluence (where 84-1 Ind4 is less confluent and therefore has fewer nuclei in all plates).


plot_dapi_3.5 <- ggplot(image_merged_dapi_summary,
                      aes(x = Condition,
                          y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 3: All Images",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_3.5

#now do stats for these to see if there are any differences over time
ttests_step3.5_dapi <- image_merged_dapi_summary %>% 
  group_by(Timepoint) %>% 
  t_test(Mean_of_Mean ~ Treatment) %>% 
  mutate(
    y.position = max(image_merged_dapi_summary$Mean_of_Mean, na.rm = TRUE) * 1
  )

sig_table_dapi_3.5 <- image_merged_dapi_summary %>% 
  group_by(Timepoint) %>% 
  summarise(
    p = t.test(Mean_of_Mean ~ Treatment)$p.value,
    y.position = max(Mean_of_Mean, na.rm = TRUE) * 1.05,
    .groups = "drop"
  ) %>% 
  mutate(
    signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint)
  )

ttests_step3.5_dapi

#now plot with these ttest statistics above the plot
plot_dapi_3.5_stats <- ggplot(image_merged_dapi_summary,
                            aes(x = Condition,
                                y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  stat_pvalue_manual(sig_table_dapi_3.5,
                     label = "signif",
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Step 3: All Images",
       x = "Condition",
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_3.5_stats

#there is some difference between VEH and DOX treated iPSC-CMs at the later timepoints

```

##Step 4 - Collapse images by individual

```{r images by ind step 4 hoechst, message=FALSE, warning=FALSE}
#now I will collapse each of these images into individual-level data
#there is some variance between individuals so it will not be perfect

#first, try the data with each image included from the entire set of images
ind_summary_dapi_step4 <- image_merged_dapi_summary %>% 
  group_by(Individual, Treatment, Timepoint, Magnification) %>% 
  summarise(
    Mean_of_Mean = mean(Mean_of_Mean, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )


ttests_step4_dapi <- ttests_step3.5_dapi %>% 
  filter(Timepoint %in% c("24T", "24R", "144R")) %>% 
  mutate(
    group1 = paste0("DOX_", Timepoint),
    group2 = paste0("VEH_", Timepoint),
    y.position = max(image_merged_dapi_summary$Mean_of_Mean, na.rm = TRUE) * 1.05,
    
    #add significance stars as defined earlier
    p.signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

#now plot these individual-level measurements with stats

plot_dapi_4 <- ggplot(ind_summary_dapi_step4, 
                     aes(x = Condition, y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1, 
             position = position_identity()) +
  # scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttests_step4_dapi,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  labs(title = "Step 4: Individual-Level Hoechst Intensity, All Images",
       x = "Condition", 
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_4


#next, try the data with only three images per condition
ind_summary_dapi_step4_filt <- image_merged_dapi_summary_final %>% 
  group_by(Individual, Treatment, Timepoint, Magnification) %>% 
  summarise(
    Mean_of_Mean = mean(Mean_of_Mean, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )


ttests_step4_dapi_filt <- ttests_step3_dapi %>% 
  filter(Timepoint %in% c("24T", "24R", "144R")) %>% 
  mutate(
    group1 = paste0("DOX_", Timepoint),
    group2 = paste0("VEH_", Timepoint),
    y.position = max(image_merged_dapi_summary_final$Mean_of_Mean, na.rm = TRUE) * 1.05,
    
    #add significance stars as defined earlier
    p.signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

#now plot these individual-level measurements with stats on the filtered image set

plot_dapi_4_filt <- ggplot(ind_summary_dapi_step4_filt, 
                     aes(x = Condition, y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1, 
             position = position_identity()) +
  # scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttests_step4_dapi_filt,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  labs(title = "Step 4: Individual-Level Hoechst Intensity, 3 Images",
       x = "Condition", 
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_4_filt

```

## Step 5 - Treatment-Level Comparisons

```{r tx level comparisons hoechst, message=FALSE, warning=FALSE}
#now, I want to take these comparisons and add another layer where I'm comparing between DOX treated at different timepoints

ind_summary_dapi_step4_dox <- image_merged_dapi_summary %>% 
  group_by(Individual, Treatment, Timepoint, Magnification) %>% 
  summarise(
    Mean_Nuclei_Count = mean(Nuclei_Count, na.rm = TRUE),
    Median_Nuclei_Count = median(Nuclei_Count, na.rm= TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )

#set the condition levels
condition_levels <- c("DOX_24T", "VEH_24T",
                      "DOX_24R", "VEH_24R",
                      "DOX_144R", "VEH_144R")

dox_conditions <- c("DOX_24T", "DOX_24R", "DOX_144R")

ind_summary_dapi_step4_dox <- ind_summary_dapi_step4_dox %>% 
  mutate(Condition = factor(Condition, levels = condition_levels))

#perform paired t-tests between DOX timepoints
timepoint_pairs <- list(
  
)

```


## Nuclei Counts Overview

```{r Nuclei Counts per Image, message=FALSE, warning=FALSE}

read_nuclei_counts <- function(cell_line, individual_label) {
  summary_path <- paste0("D:/DXR Project/IF yH2AX/if yh2ax dxr ", cell_line, 
                         "/DAPI intensity/nuclei_counts_summary.csv")
  read.csv(summary_path) %>%
    mutate(Individual = individual_label,
           Line = cell_line)
}

nuclei_counts_imgs <- bind_rows(
  read_nuclei_counts("87-1", "Ind1"),
  read_nuclei_counts("75-1", "Ind3"),
  read_nuclei_counts("84-1", "Ind4")
) %>%
  mutate(
    Magnification = case_when(
      grepl("20X", ImageName, ignore.case = TRUE) ~ "20X",
      grepl("10X", ImageName, ignore.case = TRUE) ~ "10X",
      grepl("40X", ImageName, ignore.case = TRUE) ~ "40X",
      TRUE ~ "Other"
    ),
    Treatment = case_when(
      grepl("DOX", ImageName) ~ "DOX",
      grepl("DMSO", ImageName) ~ "VEH",
      TRUE ~ "Other"
    ),
    Timepoint = case_when(
      grepl("24T", ImageName) ~ "24T",
      grepl("24R", ImageName) ~ "24R",
      grepl("144R", ImageName) ~ "144R",
      TRUE ~ "Other"
    ),
    Condition = paste0(Treatment, "_", Timepoint)
  ) %>%
  dplyr::filter(Magnification == "20X") %>%
  dplyr::filter(Treatment != "Other") %>%
  dplyr::filter(!grepl("^2nd", ImageName, ignore.case = TRUE)) 

#overview table — one row per image
nuclei_counts %>%
  mutate(Condition = factor(Condition, levels = condition_levels)) %>%
  arrange(Individual, Condition, ImageName) %>%
  dplyr::select(Individual, Line, Condition, ImageName, NucleiCount) %>%
  knitr::kable(caption = "Nuclei Count per Image") %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```



```{r nuclei counts pipeline images, message=FALSE, warning=FALSE}

nuclei_counts_imgs <- bind_rows(
  read_nuclei_counts("87-1", "Ind1"),
  read_nuclei_counts("75-1", "Ind3"),
  read_nuclei_counts("84-1", "Ind4")
) %>%
  mutate(
    Magnification = case_when(
      grepl("20X", ImageName, ignore.case = TRUE) ~ "20X",
      grepl("10X", ImageName, ignore.case = TRUE) ~ "10X",
      grepl("40X", ImageName, ignore.case = TRUE) ~ "40X",
      TRUE ~ "Other"
    ),
    Treatment = case_when(
      grepl("DOX", ImageName) ~ "DOX",
      grepl("DMSO", ImageName) ~ "VEH",
      TRUE ~ "Other"
    ),
    Timepoint = case_when(
      grepl("24T", ImageName) ~ "24T",
      grepl("24R", ImageName) ~ "24R",
      grepl("144R", ImageName) ~ "144R",
      TRUE ~ "Other"
    ),
    Condition = paste0(Treatment, "_", Timepoint)
  ) %>%
  dplyr::filter(Magnification == "20X") %>%
  dplyr::filter(Treatment != "Other") %>%
  dplyr::filter(!grepl("^2nd", ImageName, ignore.case = TRUE)) %>% 
  semi_join(pipeline_images, by = c("ImageName" = "ROI_File", "Individual", "Line"))

```



## Nuclei Count Assessment

```{r assess nuclei count from raw images across all cond, warning=FALSE, message=FALSE}
#assess whether nucleus count decreases over time from raw images

plot_nuc_count <- ggplot(image_merged_dapi_summary,
                      aes(x = Condition,
                          y = Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Assess # of Nuclei per Condition, Image Level",
       x = "Condition",
       y = "Number of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_nuc_count

#now do stats for these to see if there are any differences over time
ttests_nuc_count <- image_merged_dapi_summary %>% 
  group_by(Timepoint) %>% 
  t_test(Nuclei_Count ~ Treatment) %>% 
  mutate(
    y.position = max(image_merged_dapi_summary$Nuclei_Count, na.rm = TRUE) * 1
  )

sig_table_nuc_count <- image_merged_dapi_summary %>% 
  group_by(Timepoint) %>% 
  summarise(
    p = t.test(Nuclei_Count ~ Treatment)$p.value,
    y.position = max(Nuclei_Count, na.rm = TRUE) * 1.05,
    .groups = "drop"
  ) %>% 
  mutate(
    signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    group1 = paste0("VEH_", Timepoint),
    group2 = paste0("DOX_", Timepoint)
  )

ttests_nuc_count

#now plot with these ttest statistics above the plot
plot_nuc_count_stats <- ggplot(image_merged_dapi_summary,
                            aes(x = Condition,
                                y = Nuclei_Count)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual),
             size = 2,
             alpha = 1,
             position = position_identity()) +
  stat_pvalue_manual(sig_table_nuc_count,
                     label = "signif",
                     tip.length = 0.02) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = c(ind_cols_if)) +
  scale_fill_manual(values = txtime_col_old) +
  labs(title = "Assess # of Nuclei by Condition, Image-Level",
       x = "Condition",
       y = "Number of Nuclei") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_nuc_count_stats

```

## Assess Nuclei Count by Individual Level

```{r ind level nuclei count quant, warning=FALSE, message=FALSE}
#now I would like to assess these numbers of nuclei at an individual level

#first, collapse the data by individual
ind_summary_nuc_count <- image_merged_dapi_summary %>% 
  group_by(Individual, Treatment, Timepoint, Magnification) %>% 
  summarise(
    Median_Nuclei_Count = median(Nuclei_Count, na.rm = TRUE),
    Mean_Nuclei_count = mean(Nuclei_Count, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(
    Condition = paste0(Treatment, "_", Timepoint)
  )


#assess the mean and the median, but first do stats
ttests_nuc_count_ind <- ttests_nuc_count %>% 
  filter(Timepoint %in% c("24T", "24R", "144R")) %>% 
  mutate(
    group1 = paste0("DOX_", Timepoint),
    group2 = paste0("VEH_", Timepoint),
    y.position = max(image_merged_dapi_summary$Nuclei_Count, na.rm = TRUE) * 1.05,
    
    #add significance stars as defined earlier
    p.signif = case_when(
      p < 0.001 ~ "***",
      p < 0.01 ~ "**",
      p < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

#now plot these individual-level measurements with stats

plot_dapi_4 <- ggplot(ind_summary_dapi_step4, 
                     aes(x = Condition, y = Mean_of_Mean)) +
  geom_boxplot(aes(fill = Condition), show.legend = FALSE) +
  geom_point(aes(color = Individual), 
             size = 3, 
             alpha = 1, 
             position = position_identity()) +
  # scale_y_continuous(limits = c(0, 100)) +
  scale_x_discrete(limits = condition_levels) +
  scale_color_manual(values = ind_cols_if) +
  scale_fill_manual(values = txtime_col_old) +
  stat_pvalue_manual(ttests_step4_dapi,
                     label = "p.signif", 
                     xmin = "group1",
                     xmax = "group2", 
                     tip.length = 0.02) +
  labs(title = "Step 4: Individual-Level Hoechst Intensity, All Images",
       x = "Condition", 
       y = "Mean Intensity") +
  theme_custom() +
  theme(legend.position = "bottom")

plot_dapi_4


```

